\input{common/hyp-en}

\begin{document}
%\setlength{\emergencystretch}{1ex}
\raggedbottom

\begin{center}
\textbf{\huge\textsf{Chicken Little}}

\medskip
Cory Doctorow

\end{center}

\bigskip

\begin{flushleft}
This story is part of Cory Doctorow’s short story collection
“With a Little Help” published by himself. It is licensed under a
\href{http://creativecommons.org/licenses/by-nc-sa/}
{Creative Commons Attribution-NonCommercial-ShareAlike 3.0} license.

\bigskip

The whole volume is available at:
\texttt{http://craphound.com/walh/}

\medskip

The volume has been split into individual stories for the purpose of the
\href{http://ccbib.org}{Creative Commons Bibliothek.}
The introduction and similar accompanying texts are available under the 
title:
\end{flushleft}
\begin{center}
With a Little Help -- Extra Stuff
\end{center}

\newpage

\section{Chicken Little}

The first lesson Leon learned at the ad agency was: Nobody is your 
friend at the ad agency.

Take today: Brautigan was going to see an actual vat, at an actual 
clinic, which housed an actual target consumer, and he wasn't taking 
Leon.

“Don't sulk, it's unbecoming,” Brautigan said, giving him one of 
those tight-lipped smiles where he barely got his mouth over those big, 
horsey, comical teeth of his. They were disarming, those pearly whites. 
“It's out of the question. Getting clearance to visit a vat in 
person, that's a one month, two month process. Background checks. 
Biometrics. Interviews with their psych staff. The physicals: they have 
to take a census of your microbial nation. It takes time, Leon. You 
might be a mayfly in a mayfly hurry, but the man in the vat, he's got a 
lot of time on his hands. No skin off his dick if you get held up for a 
month or two.”

“Bullshit,” Leon said. “It's all a show. They've got a brick wall 
a hundred miles high around the front, and a sliding door around the 
back. There's always an exception in these protocols. There has to 
be.”

“When you're 180 years old and confined to a vat, you don't make 
exceptions. Not if you want to go on to 181.”

“You're telling me that if the old monster suddenly developed a rare, 
fast-moving liver cancer and there was only one oncologist in the whole 
goddamned world who could make it better, you're telling me that guy 
would be sent home to France or whatever, No thanks, we're OK, you 
don't have clearance to see the patient?”

“I'm telling you the monster \emph{doesn't have a liver}. What that 
man has, he has \emph{machines} and \emph{nutrients} and 
\emph{systems}.”

“And if a machine breaks down?”

“The man who invented that machine works for the monster. He lives on 
the monster's private estate, with his family. \emph{Their} microbial 
nations are identical to the monster's. He is not only the emperor of 
their lives, he is the emperor of the lives of their intestinal flora. 
If the machine that man invented stopped working, he would be standing 
by the vat in less than two minutes, with his staff, all in disposable, 
sterile bunny suits, murmuring reassuring noises as he calmly, expertly 
fitted one of the ten replacements he has standing by, the ten 
replacements he checks, \emph{personally}, every single day, to make 
sure that they are working.”

Leon opened his mouth, closed it. He couldn't help himself, he snorted 
a laugh. “Really?”

Brautigan nodded.

“And what if none of the machines worked?”

“If that man couldn't do it, then his rival, who \emph{also} lives on 
the monster's estate, who has developed the second-most-exciting liver 
replacement technology in the history of the world, who burns to try it 
on the man in the vat -- \emph{that} man would be there in ten minutes, 
and the first man, and his family --”

“Executed?”

Brautigan made a disappointed noise. “Come on, he's a 
quadrillionaire, not a Bond villain. No, that man would be demoted to 
nearly nothing, but given one tiny chance to redeem himself: invent a 
technology better than the one that's currently running in place of the 
vat-man's liver, and you will be restored to your fine place with your 
fine clothes and your wealth and your privilege.”

“And if he fails?”

Brautigan shrugged. “Then the man in the vat is out an unmeasurably 
minuscule fraction of his personal fortune. He takes the loss, applies 
for a research tax-credit for it, and deducts it from the pittance he 
deigns to send to the IRS every year.”

“Shit.”

Brautigan slapped his hands together. “It's wicked, isn't it? All 
that money and power and money and money?”

Leon tried to remember that Brautigan wasn't his friend. It was those 
teeth, they were so \emph{disarming}. Who could be suspicious of a man 
who was so horsey you wanted to feed him sugar cubes? “It's something 
else.”

“You now know about ten thousand times more about the people in the 
vats than your average cit. But you haven't even got the shadow of the 
picture yet, buddy. It took \emph{decades} of relationship-building for 
Ate to sell its first product to a vat-person.”

\emph{And we haven't sold anything else since,} Leon thought, but he 
didn't say it. No one would say it at Ate. The agency pitched itself as 
a powerhouse, a success in a field full of successes. It was \emph{the} 
go-to agency for servicing the “ultra-high-net-worth individual,” 
and yet...

One sale.

“And we haven't sold anything since.” Brautigan said it without a 
hint of shame. “And yet, this entire building, this entire agency, 
the salaries and the designers and the consultants: all of it paid for 
by clipping the toenails of that fortune. Which means that one 
\emph{more} sale --”

He gestured around. The offices were sumptuous, designed to impress the 
functionaries of the fortunes in the vats. A trick of light and scent 
and wind made you feel as though you were in an ancient forest glade as 
soon as you came through the door, though no forest was in evidence. 
The reception desktop was a sheet of pitted tombstone granite, the 
unreadable smooth epitaph peeking around the edges of the old fashioned 
typewriter that had been cunningly reworked to serve as a 
slightly-less-old-fashioned keyboard. The receptionist -- presently 
ignoring them with professional verisimilitude -- conveyed beauty, 
intelligence, and motherly concern, all by means of dress, bearing and 
makeup. Ate employed a small team of stylists that worked on all 
public-facing employees; Leon had endured a just-so rumpling of his 
sandy hair and some carefully applied fraying at the cuffs and elbows 
of his jacket that morning.

“So no, Leon, buddy, I am \emph{not} taking you down to meet my 
vat-person. But I \emph{will} get you started on a path that may take 
you there, some day, if you're very good and prove yourself out here. 
Once you've paid your dues.”

Leon had paid plenty of dues -- more than this blow-dried turd ever 
did. But he smiled and snuffled it up like a good little worm, hating 
himself. “Hit me.”

“Look, we've been pitching vat-products for six years now without a 
single hit. Plenty of people have come through that door and stepped 
into the job you've got now, and they've all thrown a million ideas in 
the air, and every one came smashing to earth. We've never 
systematically catalogued those ideas, never got them in any kind of 
grid that will let us see what kind of territory we've already 
explored, where the holes are...” He looked meaningfully at Leon.

“You want me to catalog every failed pitch in the agency's 
history.” Leon didn't hide his disappointment. That was the kind of 
job you gave to an intern, not a junior account exec.

Brautigan clicked his horsey teeth together, gave a laugh like a 
whinny, and left Ate's offices, admitting a breath of the boring air 
that circulated out there in the real world. The receptionist radiated 
matronly care in his direction. He leaned her way and her fingers 
thunked on the mechanical keys of her converted Underwood Noiseless, a 
machinegun rattle. He waited until she was done, then she turned that 
caring, loving smile back on him.

“It's all in your workspace, Leon -- good luck with it.”

\tb

It seemed to Leon that the problems faced by immortal quadrillionaires 
in vats wouldn't be that different from those facing mere mortals. Once 
practically anything could be made for practically nothing, everything 
was practically worthless. No one needed to discover anymore -- just 
\emph{combine}, just \emph{invent}. Then you could either hit a button 
and print it out on your desktop fab or down at the local depot for 
bigger jobs, or if you needed the kind of fabrication a printer 
couldn't handle, there were plenty of on-demand jobbers who'd have some 
worker in a distant country knock it out overnight and you'd have it in 
hermetic FedEx packaging on your desktop by the morning.

Looking through the Ate files, he could see that he wasn't the last one 
to follow this line of reasoning. Every account exec had come up with 
pitches that involved things that \emph{couldn't} be fabbed -- precious 
gewgaws that needed a trained master to produce -- or things that 
\emph{hadn't} been fabbed -- antiques, one-of-a-kinds, fetish objects 
from history. And all of it had met with crashing indifference from the 
vat-people, who could hire any master they wanted, who could buy entire 
warehouses full of antiques.

The normal megarich got offered experiences: a ticket to space, a 
chance to hunt the last member of an endangered species, the 
opportunity to kill a man and get away with it, a deep-ocean sub to the 
bottom of the Marianas trench. The people in the vat had done plenty of 
those things before they'd ended up in the vats. Now they were 
metastatic, these hyperrich, lumps of curdling meat in the pickling 
solution of a hundred vast machines that laboriously kept them alive 
amid their cancer-blooms and myriad failures. Somewhere in that tangle 
of hoses and wires was something that was technically a person, and 
also technically a corporation, and, in many cases, technically a 
sovereign state.

Each concentration of wealth was an efficient machine, meshed in a 
million ways with the mortal economy. You interacted with the vats when 
you bought hamburgers, Internet connections, movies, music, books, 
electronics, games, transportation -- the money left your hands and was 
sieved through their hoses and tubes, flushed back out into the world 
where other mortals would touch it.

But there was no easy way to touch the money at its most concentrated, 
purest form. It was like a theoretical superdense element from the 
first instant of the universe's creation, money so dense it stopped 
acting like money; money so dense it changed state when you chipped a 
piece of it off.

Leon's predecessors had been shrewd and clever. They had walked the 
length and breadth of the problem space of providing services and 
products to a person who was money who was a state who was a vat. Many 
of the nicer grace-notes in the office came from those failed pitches 
-- the business with the lights and the air, for example.

Leon had a good education, the kind that came with the mathematics of 
multidimensional space. He kept throwing axes at his chart of the 
failed inventions of Ate, Inc., mapping out the many ways in which they 
were similar and dissimilar. The pattern that emerged was easy to 
understand.

They'd tried \emph{everything}.

\tb

Brautigan's whinny was the most humiliating sound Leon had ever heard, 
in all his working life.

“No, of course you can't know what got sold to the vat-person! That 
was part of the deal -- it was why the payoff was so large. \emph{No 
one} knows what we sold to the vat-person. Not me, not the old woman. 
The man who sold it? He cashed out years ago, and hasn't been seen or 
heard from since. Silent partner, preferred shares, controlling 
interest -- but he's the invisible man. We talk to him through lawyers 
who talk to lawyers who, it is rumored, communicate by means of notes 
left under a tombstone in a tiny cemetery on Pitcairn Island, and row 
in and out in longboats to get his instructions.”

The hyperbole was grating on Leon. Third day on the job, and the 
sun-dappled, ozonated pseuodoforested environment felt as stale as an 
old gym bag (there was, in fact, an old gym bag under his desk, waiting 
for the day he finally pulled himself off the job in time to hit the 
complimentary gym). Brautigan was grating on him more than the 
hyperbole.

“I'm not an asshole, Brautigan, so stop treating me like one. You 
hired me to do a job, but all I'm getting from you is shitwork, 
sarcasm, and secrecy.” The alliteration came out without his 
intending it to, but he was good at that sort of thing. “So here's 
what I want to know: is there any single solitary reason for me to come 
to work tomorrow, or should I just sit at home, drawing a salary until 
you get bored of having me on the payroll and can my ass?”

It wasn't entirely spontaneous. Leon's industrial psychology background 
was pretty good -- he'd gotten straight A's and an offer of a post-doc, 
none of which had interested him nearly so much as the practical 
applications of the sweet science of persuasion. He understood that 
Brautigan had been pushing him around to see how far he'd push. No one 
pushed like an ad-guy -- if you could sweet-talk someone into craving 
something, it followed that you could goad him into hating something 
just as much. Two faces of a coin and all that.

Brautigan faked anger, but Leon had spent three days studying his 
tells, and Leon could see that the emotion was no more sincere than 
anything else about the man. Carefully, Leon flared his nostrils, 
brought his chest up, inched his chin higher. He \emph{sold} his 
outrage, sold it like it was potato chips, over-the-counter securities, 
or under-the-counter diet pills. Brautigan tried to sell his anger in 
return. Leon was a no-sale. Brautigan bought.

“There's a new one,” he said, in a conspiratorial whisper.

“A new what?” Leon whispered. They were still chest to chest, 
quivering with angry body-language, but Leon let another part of his 
mind deal with that.

“A new monster,” Brautigan said. “Gone to his vat at a mere 103. 
Youngest ever. Unplanned.” He looked up, down, left, right. “An 
accident. Impossible accident. Impossible, but he had it, which 
means?”

“It was no accident,” Leon said. “Police?” It was impossible 
not to fall into Brautigan's telegraphed speech-style. That was a 
persuasion thing, too, he knew. Once you talked like him, you'd 
sympathize with him. And vice-versa, of course. They were converging on 
a single identity. Bonding. It was intense, like make-up sex for 
co-workers.

“He's a sovereign three ways. An African republic, an island, one of 
those little Baltic countries. On the other side of the international 
vowel line. Mxlplx or something. They swung for him at the WTO, the UN 
-- whole bodies of international trade law for this one. So no regular 
cops; this is diplomatic corps stuff. And, of course, he's not dead, so 
that makes it more complicated.”

“How?”

“Dead people become corporations. They get managed by boards of 
directors who act predictably, if not rationally. Living people, 
they're \emph{flamboyant}. Seismic. Unpredictable. But. On the other 
hand.” He waggled his eyebrows.

“On the other hand, they buy things.”

“Once in a very long while, they do.”

\tb

Leon's life was all about discipline. He'd heard a weight-loss guru 
once explain that the key to maintaining a slim figure was to really 
“listen to your body” and only eat until it signaled that it was 
full. Leon had listened to his body. It wanted three entire pepperoni 
and mushroom pizzas every single day, plus a rather large cake. And 
malted milkshakes, the old fashioned kind you could make in your 
kitchen with an antique Hamilton Beech machine in avocado-colored 
plastic, served up in a tall red anodized aluminum cup. Leon's body was 
extremely verbose on what it wanted him to shovel into it.

So Leon ignored his body. He ignored his mind when it told him that 
what it wanted to do was fall asleep on the sofa with the video 
following his eyes around the room, one of those shows that followed 
your neural activity and tried to tune the drama to maximize your 
engrossment. Instead, he made his mind sit up in bed, absorbing many 
improving books from the mountain he'd printed out and stacked there.

Leon ignored his limbic system when it told him to stay in bed for an 
extra hour every morning when his alarm detonated. He ignored the 
fatigue messages he got while he worked through an hour of yoga and 
meditation before breakfast.

He wound himself up tight with will and it was will that made him stoop 
to pick up the laundry on the stairs while he was headed up and fold it 
neatly away when he got to the spacious walk-in dressing room attached 
to the master bedroom (the apartment had been a good way to absorb his 
Ate signing bonus -- safer than keeping the money in cash, with the 
currency fluctuations and all. Manhattan real estate was a century-long 
good buy and was more stable than bonds, derivatives or funds). It was 
discipline that made him pay every bill as it came in. It was all that 
which made him wash every dish when he was done with it and assiduously 
stop at the grocer's every night on the way home to buy anything that 
had run out the previous day.

His parents came to visit from Anguilla and they teased him about how 
\emph{organized} he was, so unlike the fat little boy who'd been 
awarded the “Hansel and Gretelprize” by his sixth grade teacher for 
leaving a trail behind him everywhere he went.

What they didn't know was that he was still that kid, and every act of 
conscientious, precise, buttoned-down finicky habit was, in fact, the 
product of relentless, iron determination not to be that kid again. He 
not only ignored that inner voice of his that called out for pizzas and 
told him to sleep in, take a cab instead of walking, lie down and let 
the video soar and dip with his moods, a drip-feed of null and nothing 
to while away the hours -- he actively denied it, shouted it into 
submission, locked it up and never let it free.

And that -- \emph{that} -- that was why he was going to figure out how 
to sell something new to the man in the vat: because anyone who could 
amass that sort of fortune and go down to life eternal in an 
ever-expanding kingdom of machines would be the sort of person who had 
spent a life denying himself, and Leon knew \emph{just} what that felt 
like.

\tb

The Lower East Side had ebbed and flowed over the years: poor, rich, 
middle-class, super-rich, poor. One year the buildings were funky and 
reminiscent of the romantic squalor that had preceded this era of 
lightspeed buckchasing. The next year, the buildings were merely 
squalorous, the landlords busted and the receivers in bankruptcy 
slapping up paper-thin walls to convert giant airy lofts into rooming 
houses. The corner stores sold blunt-skins to trustafarian hipsters 
with a bag of something gengineered to disrupt some extremely specific 
brain structures; then they sold food-stamp milk to desperate mothers 
who wouldn't meet their eyes. The shopkeepers had the knack of sensing 
changes in the wind and adjusting their stock accordingly.

Walking around his neighborhood, Leon sniffed change in the wind. The 
shopkeepers seemed to have more discount, high-calorie wino-drink; less 
designer low-carb energy food with FDA-mandated booklets explaining 
their nutritional claims. A sprinkling of FOR RENT signs. A 
construction site that hadn't had anyone working on it for a week now, 
the padlocked foreman's shed growing a mossy coat of graffiti.

Leon didn't mind. He'd lived rough -- not just student-rough, either. 
His parents had gone to Anguilla from Romania, chasing the tax-haven 
set, dreaming of making a killing working as bookkeepers, security 
guards. They'd mistimed the trip, arrived in the middle of an 
econopocalyptic collapse and ended up living in a vertical slum that 
had once been a luxury hotel. The sole Romanians among the smuggled 
Mexicans who were de-facto slaves, they'd traded their ability to write 
desperate letters to the Mexican consulate for Spanish lessons for 
Leon. The Mexicans dwindled away -- the advantage of de-facto slaves 
over de-jure slaves was that you could just send the de-facto slaves 
away when the economy tanked, taking their feed and care off your books 
-- until it was just them there, and without the safety of the crowd, 
they'd been spotted by local authorities and had to go underground. 
Going back to Bucharest was out of the question -- the airfare was as 
far out of reach as one of the private jets the tax-evaders and 
high-rolling gamblers flew in and out of Wallblake Airport.

From rough to rougher. Leon's family spent three years underground, 
living as roadside hawkers, letting the sun bake them to an ethnically 
indeterminate brown. A decade later, when his father had successfully 
built up his little bookkeeping business and his mother was running a 
smart dress-shop for the cruise-ship day-trippers, those days seemed 
like a dream. But once he left for stateside university and found 
himself amid the soft, rich children of the fortunes his father had 
tabulated, it all came back to him, and he wondered if any of these 
children in carefully disheveled rags would ever be able to pick 
through the garbage for their meals.

The rough edge on the LES put him at his ease, made him feel like he 
was still ahead of the game, in possession of something his neighbors 
could never have -- the ability to move fluidly between the worlds of 
the rich and the poor. Somewhere in those worlds, he was sure, was the 
secret to chipping a crumb off one of the great fortunes of the world.

\tb

“Visitor for you,” Carmela said. Carmela, that was the 
receptionist's name. She was Puerto Rican, but so many generations in 
that he spoke better Spanish than she did. “I put her in the Living 
Room.” That was one of the three board rooms in at Ate, the name a 
bad pun, every stick of furniture in it an elaborate topiary sculpture 
of living wood and shrubbery. It was surprisingly comfortable, and the 
very subtle breeze had an even more subtle breath of honeysuckle that 
was so real he suspected it was piped in from a nursery on another 
level. That's how he would have done it: the best fake was no fake at 
all.

“Who?” He liked Carmela. She was all business, but her business was 
compassion, a shoulder to cry on and an absolutely discreet gossip 
repository for the whole firm.

“Envoy,” she said. “Name's Buhle. I ran his face and name against 
our dossiers and came up with practically nothing. He's from 
Montenegro, originally, I have that much.”

“Envoy from whom?”

She didn't answer, just looked very meaningfully at him.

The new vat-person had sent him an envoy. His heart began to thump and 
his cuffs suddenly felt tight at his wrists. “Thanks, Carmela.” He 
shot his cuffs.

“You look fine,” she said. “I've got the kitchen on standby, and 
the intercom's listening for my voice. Just let me know what I can do 
for you.”

He gave her a weak smile. This was why she was the center of the whole 
business, the soul of Ate. \emph{Thank you,} he mouthed, and she ticked 
a smart salute off her temple with one finger.

\tb

The envoy was out of place in Ate, but she didn't hold it against them. 
This he knew within seconds of setting foot into the Living Room. She 
got up, wiped her hands on her sensible jeans, brushed some iron-grey 
hair off her face, and smiled at him, an expression that seemed to say, 
“Well, this is a funny thing, the two of us, meeting here, like 
this.” He'd put her age at around 40, and she was hippy and a little 
wrinkled and didn't seem to care at all.

“You must be Leon,” she said, and took his hand. Short fingernails, 
warm, dry, palm, firm handshake. “I \emph{love} this room!” She 
waved her arm around in an all-encompassing circle. “Fantastic.”

He found himself half in love with her and he hadn't said a word. 
“It's nice to meet you, Ms --”

“Ria,” she said. “Call me Ria.” She sat down on one of the 
topiary chairs, kicking off her comfortable hush puppies and pulling 
her legs up to sit cross-legged.

“I've never gone barefoot in this room,” he said, looking at her 
calloused feet -- feet that did a lot of barefooting.

“Do it,” she said, making scooting gestures. “I insist. Do it!”

He kicked off the handmade shoes -- designed by an architect who'd 
given up on literary criticism to pursue cobblery -- and used his toes 
to peel off his socks. Under his feet, the ground was -- warm? cool? -- 
it was \emph{perfect}. He couldn't pin down the texture, but it made 
every nerve ending on the sensitive soles of his feet tingle pleasantly.

“I'm thinking something that goes straight into the nerves,” she 
said. “It has to be. Extraordinary.”

“You know your way around this place better than I do,” he said.

She shrugged. “This room was clearly designed to impress. It would be 
stupid to be so cool-obsessed that I failed to let it impress me. I'm 
impressed. Also,” she dropped her voice, “also, I'm wondering if 
anyone's ever snuck in here and screwed on that stuff.” She looked 
seriously at him and he tried to keep a straight face, but the chuckle 
wouldn't stay put in his chest, and it broke loose, and a laugh 
followed it, and she whooped and they both laughed, hard, until their 
stomachs hurt.

He moved toward another topiary easy-chair, then stopped, bent down, 
and sat on the mossy floor, letting it brush against his feet, his 
ankles, the palms of his hands and his wrists. “If no one ever has, 
it's a damned shame,” he said, with mock gravity. She smiled, and she 
had dimples and wrinkles and crowsfeet, so her whole face smiled. “Do 
you want something to eat? Drink? We can get pretty much anything here 
--”

“Let's get to it,” she said. “I don't want to be rude, but the 
good part isn't the food. I get all the food I need. I'm here for 
something else. The good part, Leon.”

He drew in a deep breath. “The good part,” he said. “OK, let's 
get to it. I want to meet your --” What? Employer? Patron? Owner? He 
waved his hand.

“You can call him Buhle,” she said. “That's the name of the 
parent company, anyway. Of course you do. We have an entire corporate 
intelligence arm that knew you'd want to meet with Buhle before you 
did.” Leon had always assumed that his workspaces and communications 
were monitored by his employer, but now it occurred to him that any 
system designed from the ground up to subject its users to scrutiny 
without their knowledge would be a bonanza for anyone \emph{else} who 
wanted to sniff them, since they could use the system's own 
capabilities to hide their snooping from the victims.

“That's impressive,” he said. “Do you monitor everyone who might 
want to pitch something to Buhle, or --” He let the thought hang out 
there.

“Oh, a little of this and a little of that. We've got a competitive 
intelligence subdepartment that monitors everyone who might want to 
sell us something or sell something that might compete with us. It 
comes out to a pretty wide net. Add to that the people who might 
personally be a threat or opportunity for Buhle and you've got, well, 
let's say an appreciable slice of human activity under close 
observation.”

“How close can it be? Sounds like you've got some big haystacks.”

“We're good at finding the needles,” she said. “But we're always 
looking for new ways to find them. That's something you could sell us, 
you know.”

He shrugged. “If we had a better way of finding relevance in 
mountains of data, we'd be using it ourselves to figure out what to 
sell you.”

“Good point. Let's turn this around. Why should Buhle meet with 
you?”

He was ready for this one. “We have a track-record of designing 
products that suit people in his...” Talking about the vat-born lent 
itself to elliptical statements. Maybe that's why Brautigan had 
developed that annoying telegraph-talk.

“You've designed one such product,” she said.

“That's one more than almost anyone else can claim.” There were two 
other firms like Ate. He thought of them in his head as Sefen and Nein, 
as though invoking their real names might cause them to appear. “I'm 
new here, but I'm not alone. We're tied in with some of the finest 
designers, engineers, research scientists...” Again with the 
ellipsis. “You wanted to get to the good part. This isn't the good 
part, Ria. You've got smart people. We've got smart people. What we 
have, what you don't have, is smart people who are impedance-mismatched 
to your organization. Every organization has quirks that make it 
unsuited to working with some good people and good ideas. You've got 
your no-go areas, just like anyone else. We're good at mining that 
space, the no-go space, the mote in your eye, for things that you 
need.”

She nodded and slapped her hands together like someone about to start a 
carpentry project. “That's a great spiel,” she said.

He felt a little blush creep into his cheeks. “I think about this a 
lot, rehearse it in my head.”

“That's good,” she said. “Shows you're in the right line of 
business. Are you a Daffy Duck man?”

He cocked his head. “More of a Bugs man,” he said, finally, 
wondering where this was going.

“Go download a cartoon called `The Stupor Salesman,' and get back to 
me, OK?”

She stood up, wriggling her toes on the mossy surface and then stepping 
back into her shoes. He scrambled to his feet, wiping his palms on his 
legs. She must have seen the expression on his face because she made 
all those dimples and wrinkles and crowsfeet appear again and took his 
hand warmly. “You did very well,” she said. “We'll talk again 
soon.” She let go of his hand and knelt down to rub her hands over 
the floor. “In the meantime, you've got a pretty sweet gig, don't 
you?”

\tb

The Stupor Salesman turned out to feature Daffy Duck as a traveling 
salesman bent on selling something to a bank robber who is holed up in 
a suburban bungalow. Daffy produces a stream of ever-more-improbable 
wares, and is violently rebuffed with each attempt. Finally, one of his 
attempts manages to blow up the robber's hideout, just as Daffy is once 
again jiggling the doorknob. As the robber and Daffy fly through the 
air, Daffy brandishes the doorknob at him and shouts, “Hey, bub, I 
know just what you need! You need a house to go with this doorknob!”

The first time he watched it, Leon snorted at the punchline, but on 
subsequent viewings, he found himself less and less amused. Yes, he was 
indeed trying to come up with a need that this Buhle didn't know he had 
-- he was assuming Buhle was a he, but no one was sure -- and then fill 
it. From Buhle's perspective, life would be just fine if Leon gave up 
and never bothered him again.

And yet Ria had been so \emph{nice} -- so understanding and gentle, he 
thought there must be something else to this. And she had made a point 
of telling him that he had a “sweet gig” and he had to admit that 
it was true. He was contracted for five years with Ate, and would get a 
hefty bonus if they canned him before then. If he managed to score a 
sale to Buhle or one of the others, he'd be indescribably wealthy. In 
the meantime, Ate took care of his every need.

But it was so \emph{empty} there -- that's what got him. There were a 
hundred people on Ate's production team, bright sorts like him, and 
most of them only used the office to park a few knick-knacks and 
impress out-of-town relatives. Ate hired the best, charged them with 
the impossible and turned them loose. They got lost.

Carmela knew them all, of course. She was Ate's den-mother.

“We should all get together,” he said. “Maybe a weekly staff 
meeting?”

“Oh, they tried that,” she said, sipping from the triple-filtered 
water that was always at her elbow. “No one had much to say. The 
collaboration spaces update themselves with all the interesting leads 
from everyone's research, and the suggestion engine is pretty good at 
making sure you get an overview of anything relevant to your work going 
on.” She shrugged. “This place is a show-room, more than anything 
else. I always figured you had to give creative people room to be 
creative.”

He mulled this over. “How long do you figure they'll keep this place 
open if it doesn't sell anything to one of the vat people?”

“I try not to think about that too much,” she said lightly. “I 
figure either we don't find something, run out of time and shut -- and 
there's nothing I can do about it; or we find something in time and 
stay open -- and there's nothing I can do about it.”

“That's depressing.”

“I think of it as liberating. It's like that lady said, Leon, you've 
got a sweet gig. You can make anything you can imagine, and if you hit 
one out of the park, you'll attain orbit and never reenter the 
atmosphere.”

“Do the other account execs come around for pep talks?”

“Everyone needs a little help now and then,” she said.

\tb

Ria met him for lunch at a supper-club in the living room of an 11th 
floor apartment in a slightly run-down ex-doorman building in midtown. 
The cooks were a middle-aged couple, he was Thai, she was Hungarian, 
the food was eclectic, light, and spicy, blending paprika and chilis in 
a nose-watering cocktail.

There were only two other diners in the tiny room for the early 
seating. They were another couple, two young gay men, tourists from the 
Netherlands, wearing crease-proof sportsjackets and barely-there 
barefoot hiking shoes. They spoke excellent English, and chatted 
politely about the sights they'd seen so far in New York, before 
falling into Dutch and leaving Ria and Leon to concentrate on each 
other and the food, which emerged from the kitchen in a series of 
ever-more-wonderful courses.

Over fluffy, caramelized fried bananas and Thai iced coffee, Ria 
effusively praised the food to their hosts, then waited politely while 
Leon did the same. The hosts were genuinely delighted to have fed them 
so successfully, and were only too happy to talk about their recipes, 
their grown children, the other diners they'd entertained over the 
years.

Outside, standing on 34th street between Lex and Third, a cool summer 
evening breeze and purple summer twilight skies, Leon patted his 
stomach and closed his eyes and groaned.

“Ate too much, didn't you?” she said.

“It was like eating my mother's cooking -- she just kept putting more 
on the plate. I couldn't help it.”

“Did you enjoy it?”

He opened his eyes. “You're kidding, right? That was probably the 
most incredible meal I've eaten in my entire life. It was like a 
parallel dimension of good food.”

She nodded vigorously and took his arm in a friendly, intimate gesture, 
led him toward Lexington. “You notice how time sort of stops when 
you're there? How the part of your brain that's going `what next? what 
next?' goes quiet?”

“That's it! That's \emph{exactly} it!” The buzz of the jetpacks on 
Lex grew louder as they neared the corner, like a thousand crickets in 
the sky.

“Hate those things,” she said, glaring up at the joyriders zipping 
past, scarves and capes streaming out behind them. “A thousand 
crashes upon your souls.” She spat, theatrically.

“You make them, though, don't you?”

She laughed. “You've been reading up on Buhle then?”

“Everything I can find.” He'd bought small blocks of shares in all 
the public companies in which Buhle was a substantial owner, charging 
them to Ate's brokerage account, and then devoured their annual 
reports. There was lots more he could feel in the shadows: blind trusts 
holding more shares in still more companies. It was the standard 
corporate structure, a Flying Spaghetti Monster of interlocking 
directorships, offshore holdings, debt parking lots, and exotic 
matrioshke companies that seemed on the verge of devouring themselves.

“Oy,” she said. “Poor boy. Those aren't meant to be parsed. 
They're like the bramble patch around the sleeping princess, there to 
ensnare foolhardy knights who wish to court the virgin in the tower. 
Yes, Buhle's the largest jetpack manufacturer in the world, through a 
layer or two of misdirection.” She inspected the uptownbound horde, 
sculling the air with their fins and gloves, making course corrections 
and wibbles and wobbles that were sheer, joyful exhibitionism.

“He did it for me,” she said. “Have you noticed that they've 
gotten better in the past couple years? Quieter? That was us. We put a 
lot of thought into the campaign; the chop-shops have been selling 
`loud pipes save lives' since the motorcycle days, and every tiny-dick 
flyboy wanted to have a pack that was as loud as a bulldozer. It took a 
lot of market smarts to turn it around; we had a low-end model we were 
selling way below cost that was close to those loud-pipe machines in 
decibel count; it was ugly and junky and fell apart. Naturally, we sold 
it through a different arm of the company that had totally different 
livery, identity and everything. Then we started to cut into our 
margins on the high-end rides, and at the same time, we engineered them 
for a quieter and quieter run. We actually did some preproduction on a 
jetpack that was so quiet it actually \emph{absorbed} noise, don't ask 
me to explain it, unless you've got a day or two to waste on the 
psycho-acoustics.

“Every swish bourgeois was competing to see whose jetpack could run 
quieter, while the low-end was busily switching loyalty to our loud 
junkmobiles. The competition went out of business in a year, and then 
we dummied-up a bunch of consumer-protection lawsuits that `forced' -- 
” she drew air-quotes -- “us to recall the loud ones, rebuild them 
with pipes so engineered and tuned you could use them for the woodwinds 
section. And here we are.” She gestured at the buzzing, whooshing 
fliers overhead.

Leon tried to figure out if she was kidding, but she looked and sounded 
serious. “You're telling me that Buhle dropped, what, a billion?”

“About eight billion, in the end.”

“Eight billion rupiah on a project to make the skies quieter?”

“All told,” she said. “We could have done it other ways, some of 
them cheaper. We could have bought some laws, or bought out the 
competition and changed their product line, but that's very, you know, 
\emph{blunt}. This was sweet. Everyone got what they wanted in the end: 
fast rides, quiet skies, safe, cheap vehicles. Win win win.”

An old school flyer with a jetpack as loud as the inside of an 
ice-blender roared past, leaving thousands scowling in his wake.

“That guy is plenty dedicated,” she said. “He'll be machining his 
own replacement parts for that thing. No one's making them anymore.”

He tried a joke: “You're not going to send the Buhle ninjas to off 
him before he hits Union Square?”

She didn't smile. “We don't use assassination,” she said. “That's 
what I'm trying to convey to you, Leon.”

He crumbled. He'd blown it somehow, shown himself to be the boor he'd 
always feared he was.

“I'm sorry,” he said. “I guess -- look, it's all kind of hard to 
take in. The sums are staggering.”

“They're meaningless,” she said. “That's the point. The sums are 
just a convenient way of directing power. Power is what matters.”

“I don't mean to offend you,” he said carefully, “but that's a 
scary sounding thing to say.”

“Now you're getting it,” she said, and took his arm again. 
“Drinks?”

\tb

The limes for the daquiris came from the trees around them on the 
rooftop conservatory. The trees were healthy working beasts, and the 
barman expertly inspected several limes before deftly twisting off a 
basket's worth and retreating to his workbench to juice them over his 
blender.

“You have to be a member to drink here,” Ria said, as they sat on 
the roof, watching the jetpacks scud past.

“I'm not surprised,” he said. “It must be expensive.”

“You can't buy your way in,” she said. “You have to work it off. 
It's a co-op. I planted this whole row of trees.” She waved her arm, 
sloshing a little daquiri on the odd turf their loungers rested on. 
“I planted the mint garden over there.” It was a beautiful little 
patch, decorated with rocks and favored with a small stream that wended 
its way through them.

“Forgive me for saying this,” he said, “but you must earn a lot 
of money. A \emph{lot}, I'm thinking.” She nodded, unembarrassed, 
even waggled her eyebrows a bit. “So you could, I don't know, you 
could probably build one of these on any of the buildings that Buhle 
owns in Manhattan. Just like this. Even keep a little staff on board. 
Give out memberships as perks for your senior management team.”

“That's right,” she said. “I could.”

He drank his daquiri. “I'm supposed to figure out why you don't, 
right?”

She nodded. “Indeed.” She drank. Her face suffused with pleasure. 
He took a moment to pay attention to the signals his tongue was 
transmitting to him. The drink was \emph{incredible}. Even the glass 
was beautiful, thick, handblown, irregular. “Listen, Leon, I'll let 
you in on a secret. \emph{I want you to succeed}. There's not much that 
surprises Buhle and even less that pleasantly surprises him. If you 
were to manage it...” She took another sip and looked intensely at 
him. He squirmed. Had he thought her matronly and sweet? She looked 
like she could lead a guerilla force. Like she could wrestle a mugger 
to the ground and kick the shit out of him.

“So a success for me would be a success for you?”

“You think I'm after money,” she said. “You're still not getting 
it. Think about the jetpacks, Leon. Think about what that power 
means.”

\tb

He meant to go home, but he didn't make it. His feet took him crosstown 
to the Ate offices, and he let himself in with his biometrics and his 
passphrase and watched the marvelous dappled lights go through their 
warm-up cycle and then bathe him with their wonderful, calming light. 
Then the breeze, and now it was a nighttime forest, mossier and heavier 
than in the day. Either someone had really gone balls-out on the 
product design, or there really was an indoor forest somewhere in the 
building growing under diurnal lights, there solely to supply soothing 
woodsy air to the agency's office. He decided that the forest was the 
more likely explanation.

He stood at Carmela's desk for a long time, then, gingerly, settled 
himself in her chair. It was plain and firm and well made, with just a 
little spring. Her funny little sculptural keyboard had keycaps that 
had worn smooth under her fingertips over the years, and there were 
shiny spots on the desk where her wrists had worn away the granite. He 
cradled his face in his palms, breathing in the nighttime forest air, 
and tried to make sense of the night.

The Living Room was nighttime dark, but it still felt glorious on his 
bare feet, and then, a moment later, on his bare chest and legs. He lay 
on his stomach in his underwear and tried to name the sensation on his 
nerve endings and decided that “anticipation” was the best word for 
it, the feeling you get just \emph{beside} the skin that's being 
scratched on your back, the skin that's next in line for a good 
scratching. It was glorious.

How many people in the world would ever know what this felt like? Ate 
had licensed it out to a few select boutique hotels -- he'd checked 
into it after talking with Ria the first time -- but that was it. All 
told, there were less than 3,000 people in the world who'd ever felt 
this remarkable feeling. Out of eight billion. He tried to do the 
division in his head but kept losing the zeroes. It was a thousandth of 
a percent? A ten thousandth of a percent? No one on Anguilla would ever 
feel it: not the workers in the vertical slums, but also not the mere 
millionaires in the grand houses with their timeshare jets.

Something about that...

He wished he could talk to Ria some more. She scared him, but she also 
made him feel good. Like she was the guide he'd been searching for all 
his life. At this point, he would have settled for Brautigan. Anyone 
who could help him make sense of what felt like the biggest, scariest 
opportunity of his entire career.

He must have dozed, because the next thing he knew, the lights were 
flickering on and he was mostly naked, on the floor, staring up into 
Brautigan's face. He had a look of forced jollity, and he snapped his 
fingers a few times in front of Leon's face.

“Morning, sunshine!”

Leon looked for the ghostly clock that shimmered in the corner of each 
wall, a slightly darker patch of reactive paint that was just outside 
of conscious comprehension unless you really stared at it. 4:12 AM. He 
stifled a groan.

“What are you doing here?” he said, peering at Brautigan.

The man clacked his horsey teeth, assayed a chuckle. “Early bird. 
Worm.”

Leon sat up, found his shirt, started buttoning it up. “Seriously, 
Brautigan.”

“Seriously?” He sat down on the floor next to Leon, his big feet 
straight out ahead of him. His shoes had been designed by the same 
architect that did Leon's. Leon recognized the style. “Seriously.” 
He scratched his chin. Suddenly, he slumped. “I'm shitting bricks, 
Leon. I am seriously shitting bricks.”

“How did it go with your monster?”

Brautigan stared at the architect's shoes. There was an odd flare they 
did, just behind the toe, just on the way to the laces, that was really 
graceful. Leon thought it might be a standard distribution bell-curve. 
“My monster is...” He blew out air. “Uncooperative.”

“Less cooperative than previously?”

Brautigan unlaced his shoes and peeled off his socks, scrunched his 
toes in the moss. His feet gave off a hot, trapped smell.

“What was he like on the other times you'd seen him?”

Brautigan tilted his head. “What do you mean?”

“He was uncooperative this time, what about the other times?”

Brautigan looked back down at his toes.

“You'd \emph{never} seen him before this?”

“It was a risk,” he said. “I thought I could convince him, face 
to face.”

“But?”

“I bombed. It was -- it was the -- it was \emph{everything}. The 
compound. The people. All of it. It was like a \emph{city}, a 
\emph{theme park}. They lived there, hundreds of them, and managed 
every tiny piece of his empire. Like Royal Urchins.”

Leon puzzled over this. “Eunuchs?”

“Royal Eunuchs. They had this whole culture, and as I got closer and 
closer to him, I realized, shit, they could just \emph{buy} Ate. They 
could destroy us. They could have us made illegal, put us all in jail. 
Or get me elected president. Anything.”

“You were overawed.”

“That's the right word. It wasn't a castle or anything, either. It 
was just a place, a well-built collection of buildings. In Westchester, 
you know? It had been a little town center once. They'd preserved 
everything good, built more on top of it. It all just...worked. You're 
still new here. Haven't noticed.”

“What? That Ate is a disaster? I figured that out a long time ago. 
There's several dozen highly paid creative geniuses on the payroll here 
who haven't seen their desks in months. We could be a creative 
powerhouse. We're more like someone's vanity project.”

“Brutal.”

He wondered if he'd overstepped himself. Who cared?

“Brutal doesn't mean untrue. It's like, it's like the money that came 
into this place, it became autonomous, turned into a strategy for 
multiplying itself. A bad strategy. The money wants to sell something 
to a monster, but the money doesn't know what monsters want, so it's 
just, what, beating its brains out on the wall. One day, the money runs 
out and...”

“The money won't run out,” Brautigan said. “Wrong. We'd have to 
spend at ten-ex what we're burning now to even approach the 
principal.”

“OK,” Leon said. “So it's immortal. That's better?”

Brautigan winced. “Look, it's not so crazy. There's an entire 
unserved market out there. No one's serving it. They're like, you know, 
like Communist countries. Planned economies. They need something, they 
just acquire the capacity. No market.”

“Hey, bub, I know just what you need! You need a house to go with 
this doorknob!” To his own surprise, Leon discovered that he did a 
passable Daffy Duck. Brautigan blinked at him. Leon realized that the 
man was a little drunk. “Just something I heard the other day,” he 
said. “I told the lady from my monster that we could provide the 
stuff that their corporate culture precluded. I was thinking of you 
know, how the samurai banned firearms. We can think and do the 
unthinkable and undo-able.”

“Good line.” He flopped onto his back. An inch of pale belly peeked 
between the top of his three-quarter-length culottes and the lower hem 
of his smart wraparound shirt. “The monster in the vat. Some skin, 
some meat. Tubes. Pinches of skin clamped between clear hard plastic 
squares, bathed in some kind of diagnostic light. No eyes, no top of 
the head where the eyes should be. Just a smooth mask. Eyes everywhere 
else. Ceiling. Floor. Walls. I looked away, couldn't make contact with 
them, found I was looking at something wet. Liver. I think.”

“Yeesh. That's immortality, huh?”

“I'm there, `A pleasure to meet you, an honor,' talking to the liver. 
The eyes never blinked. The monster gave a speech. `You're a 
low-capital, high-risk, high-payoff longshot Mr Brautigan. I can keep 
dribbling sums to you so that you can go back to your wonder factory 
and try to come up with ways to surprise me. So there's no need to 
worry on that score.' And that was it. Couldn't think of anything to 
say. Didn't have time. Gone in a flash. Out the door. Limo. Nice babu 
to tell me how good it had been for the monster, how much he'd been 
looking forward to it.” He struggled up onto his elbows. “How about 
you?”

Leon didn't want to talk about Ria with Brautigan. He shrugged. 
Brautigan got a mean, stung look on his face. “Don't be like that. 
Bro. Dude. Pal.”

Leon shrugged again. Thing was, he \emph{liked} Ria. Talking about her 
with Brautigan would be treating her like a... a \emph{sales-target}. 
If he were talking with Carmela, he'd say, “I feel like she wants me 
to succeed. Like it would be a huge deal for everyone if I managed it. 
But I also feel like maybe she doesn't think I can.” But to 
Brautigan, he merely shrugged, ignored the lizardy slit-eyed glare, 
stood, pulled on his pants, and went to his desk.

\tb

If you sat at your desk long enough at Ate, you'd eventually meet 
\emph{everyone} who worked there. Carmela knew all, told all, and 
assured him that everyone touched base at least once a month. Some came 
in a couple times a week. They had plants on their desks and liked to 
personally see to their watering.

Leon took every single one of them to lunch. It wasn't easy -- in one 
case, he had to ask Carmela to send an Ate chauffeur to pick up the 
man's kids from school (it was a half-day) and bring them to the 
sitter's, just to clear the schedule. But the lunches themselves went 
very well. It turned out that the people at Ate were, to a one, 
incredibly interesting. Oh, they were all monsters, narcissistic, 
tantrum-prone geniuses, but once you got past that, you found yourself 
talking to people who were, at bottom, damned smart, with a whole lot 
going on. He met the woman who designed the moss in the Living Room. 
She was younger than him, and had been catapulted from a mediocre 
academic adventure at the Cooper Union into more wealth and freedom 
than she knew what to do with. She had a whole rolodex of people who 
wanted to sublicense the stuff, and she spent her days toying with 
them, seeing if they had any cool ideas she could incorporate into her 
next pitch to one of the lucky few who had the ear of a monster.

Like Leon. That's why they all met with him. He'd unwittingly stepped 
into one of the agency's top spots, thanks to Ria, one of the 
power-broker seats that everyone else yearned to fill. The fact that he 
had no idea how he'd got there or what to do with it didn't surprise 
anyone. To a one, his colleagues at Ate regarded everything to do with 
the vat monsters as an absolute, unknowable crapshoot, as predictable 
as a meteor strike.

No wonder they all stayed away from the office.

\tb

Ria met him in a different pair of jeans, these ones worn and patched 
at the knees. She had on a loose, flowing silk shirt that was frayed 
around the seams, and had tied her hair back with a kerchief that had 
faded to a non-color that was like the ancient New York sidewalk 
outside Ate's office. He felt the calluses on her hand when they shook.

“You look like you're ready to do some gardening,” he said.

“My shift at the club,” she said. “I'll be trimming the lime 
trees and tending the mint patch and the cucumber frames all 
afternoon.” She smiled and stopped him with a gesture. She bent down 
and plucked a blade of greenery from the untidy trail-edge. They were 
in Central Park, in one of the places where it felt like a primeval 
forest instead of an artful garden razed and built in the middle of the 
city. She uncapped her water bottle and poured water over the herb -- 
it looked like a blade of grass -- rubbing it between her forefinger 
and thumb to scrub at it. Then she tore it in two and handed him one 
piece, held the other to her nose, then ate it, nibbling and making her 
nose wrinkle like a rabbit's. He followed suit. Lemon, delicious and 
tangy.

“Lemon grass,” she said. “Terrible weed, of course. But doesn't 
it taste amazing?”

He nodded. The flavor lingered in his mouth.

“Especially when you consider what this is made of -- smoggy rain, 
dog piss, choked up air, and sunshine, and DNA. What a weird flavor to 
emerge from such a strange soup, don't you think?”

The thought made the flavor a little less delicious. He said so.

“I love the idea,” she said. “Making great things from garbage.”

“About the jetpacks,” he said, for he'd been thinking.

“Yes?”

“Are you utopians of some kind? Making a better world?”

“By `you,' you mean `people who work for Buhle?'”

He shrugged.

“I'm a bit of a utopian, I'll admit. But that's not it. You know 
Henry Ford set up these work-camps in Brazil, `Fordlandia,' and 
enforced a strict code of conduct on the rubber plantation workers? He 
outlawed the caipirinha and replaced it with Tom Collinses, because 
they were more civilized.”

“And you're saying Buhle wouldn't do that?”

She waggled her head from side to side, thinking it over. “Probably 
not. Maybe, if I asked.” She covered her mouth as though she'd made 
an indiscreet admission.

“Are -- \emph{were} -- you and he...?”

She laughed. “Never. It's purely cerebral. Do you know where his 
money came from?”

He gave her a look.

“OK, of course you do. But if all you've read is the official 
history, you'll think he was just a finance guy who made some good 
bets. It's nothing like it. He played a game against the market, 
tinkered with the confidence of other traders by taking crazy 
positions, all bluff, except when they weren't. No one could outsmart 
him. He could convince you that you were about to miss out on the deal 
of the century, or that you'd already missed it, or that you were about 
to walk off onto easy street. Sometimes, he convinced you of something 
that was real. More often, it was pure bluff, which you'd only find out 
after you'd done some trade with him that left him with more money than 
you'd see in your whole life, and you facepalming and cursing yourself 
for a sucker. When he started doing it to national banks, put a run on 
the dollar, broke the Fed, well, that's when we all knew that he was 
someone who was \emph{special}, someone who could create signals that 
went right to your hindbrain without any critical interpretation.”

“Scary.”

“Oh yes. Very. In another era they'd have burned him for a witch or 
made him the man who cut out your heart with the obsidian knife. But 
here's the thing: he could never, ever kid \emph{me}. Not once.”

“And you're alive to tell the tale?”

“Oh, he likes it. His reality distortion field, it screws with his 
internal landscape. Makes it hard for him to figure out what he needs, 
what he wants, and what will make him miserable. I'm indispensable.”

He had a sudden, terrible thought. He didn't say anything, but she must 
have seen it on his face.

“What is it? Tell me.”

“How do I know that you're on the level about any of this? Maybe 
you're just jerking me around. Maybe it's all made up -- the jetpacks, 
everything.” He swallowed. “I'm sorry. I don't know where that came 
from, but it popped into my head --”

“It's a fair question. Here's one that'll blow your mind, though: how 
do you know that I'm not on the level, \emph{and} jerking you around?”

They changed the subject soon after, with uneasy laughter. They ended 
up on a park bench near the family of dancing bears, whom they watched 
avidly.

“They seem so \emph{happy},” he said. “That's what gets me about 
them. Like dancing was the secret passion of every bear, and these 
three are the first to figure out how to make a life of it.”

She didn't say anything, but watched the three giants lumber in a 
graceful, unmistakably joyous kind of shuffle. The music -- constantly 
mutated based on the intensity of the bears, a piece of software that 
sought tirelessly to please them -- was jangly and pop-like, with a 
staccato one-two/\-onetwothreefourfive/\-one-two rhythm that let the 
bears do something like a drunken stagger that was as fun to watch as a 
box of puppies.

He felt the silence. “So happy,” he said again. “That's the weird 
part. Not like seeing an elephant perform. You watch those old videos 
and they seem, you know, they seem --”

“Resigned,” she said.

“Yeah. Not unhappy, but about as thrilled to be balancing on a ball 
as a horse might be to be hitched to a plough. But look at those 
bears!”

“Notice that no one else watches them for long?” she said.

He had noticed that. The benches were all empty around them.

“I think it's because they're so happy,” she said. “It lays the 
trick bare.” She showed teeth at the pun, then put them away. “What 
I mean is, you can see how it's possible to design a bear that 
experiences brain reward from rhythm, keep it well-fed, supply it with 
as many rockin' tunes as it can eat, and you get that happy family of 
dancing bears who'll peacefully co-exist alongside humans who're going 
to work, carrying their groceries, pushing their toddlers around in 
strollers, necking on benches --”

The bears were resting now, lolling on their backs, happy tongues 
sloppy in the corners of their mouths.

“We made them,” she said. “It was against my advice, too. There's 
not much subtlety in it. As a piece of social commentary, it's a 
cartoon sledgehammer with an oversized head. But the artist had Buhle's 
ear, he'd been CEO of one of the portfolio companies and had been 
interested in genomic art as a sideline for his whole career. Buhle saw 
that funding this thing would probably spin off lots of interesting 
sublicenses, which it did. But just look at it.”

He looked. “They're \emph{so happy},” he said.

She looked too. “Bears shouldn't be that happy,” she said.

\tb

Carmela greeted him sunnily as ever, but there was something odd.

“What is it?” he asked in Spanish. He made a habit of talking 
Spanish to her, because both of them were getting rusty, and also it 
was like a little shared secret between them.

She shook her head.

“Is everything all right?” Meaning, \emph{Are we being shut down?} 
It could happen, might happen at any time, with no notice. That was 
something he -- all of them -- understood. The money that powered them 
was autonomous and unknowable, an alien force that was more emergent 
property than will.

She shook her head again. “It's not my place to say,” she said. 
Which made him even more sure that they were all going down, for when 
had Carmela ever said anything about her \emph{place}?

“Now you've got me worried,” he said.

She cocked her head back toward the back office. He noticed that there 
were three coats hung on the beautiful, anachronistic coat-stand by the 
ancient temple door that divided reception from the rest of Ate.

He let himself in and walked down the glassed-in double-rows of 
offices, the cubicles in the middle, all with their characteristic 
spotless hush, like a restaurant dining room set up for the meals that 
people would come to later.

He looked in the Living Room, but there was no one there, so he began 
to check out the other conference rooms, which ran the gamut from 
super-conservative to utter madness. He found them in the Ceile, with 
its barn-board floors, its homey stone hearth, and the gimmicked sofas 
that looked like unsprung old thrift-store numbers, but which sported 
adaptive genetic-algorithm-directed haptics that adjusted constantly to 
support you no matter how you flopped on them, so that you could play 
at being a little kid sprawled carelessly on the cushions no matter how 
old and cranky your bones were.

On the Ceile's sofa were Brautigan, Ria, and a woman he hadn't met 
before. She was somewhere between Brautigan and Ria's age, but with 
that made-up, pulled-tight appearance of someone who knew the world 
wouldn't take her as seriously if she let one crumb of weakness escape 
from any pore or wrinkle. He thought he knew who this must be, and she 
confirmed it when she spoke.

“Leon,” she said. “I'm glad you're here.” He knew that voice. 
It was the voice on the phone that had recruited him and brought him to 
New York and told him where to come for his first day on the job. It 
was the voice of Jennifer Torino, and she was technically his boss. 
“Carmela said that you often worked from here so I was hoping today 
would be one of the days you came by so we could chat.”

“Jennifer,” he said. She nodded. “Ria.” She had a poker-face 
on, as unreadable as a slab of granite. She was wearing her customary 
denim and flowing cotton, but she'd kept her shoes on and her feet on 
the ground. “Brautigan,” and Brautigan grinned like it was 
Christmas morning.

Jennifer looked flatly at a place just to one side of his gaze, a trick 
he knew, and said, “In recognition of his excellent work, Mr 
Brautigan's been promoted, effective today. He is now Manager for Major 
Accounts.” Brautigan beamed.

“Congratulations,” Leon said, thinking, \emph{What excellent work? 
No one at Ate has accomplished the agency's primary objective in the 
entire history of the firm!} “Well done.”

Jennifer kept her eyes coolly fixed on that empty, safe spot. “As you 
know, we have struggled to close a deal with any of our major 
accounts.” He restrained himself from rolling his eyes. “And so Mr 
Brautigan has undertaken a thorough study of the way we handle these 
accounts.” She nodded at Brautigan.

“It's a mess,” he said. “Totally scattergun. No lines of 
authority. No checks and balances. No system.”

“I can't argue with that,” Leon said. He saw where this was going.

“Yes,” Jennifer said. “You haven't been here very long, but I 
understand you've been looking deeply into the organizational structure 
of Ate yourself, haven't you?” He nodded. “And that's why Mr 
Brautigan has asked that you be tasked to him as his head of strategic 
research.” She smiled a thin smile. “Congratulations yourself.”

He said, “Thanks,” flatly, and looked at Brautigan. “What's 
strategic research, then?”

“Oh,” Brautigan said. “Just a lot of what you've been doing: 
figuring out what everyone's up to, putting them together, proposing 
organizational structures that will make us more efficient at design 
and deployment. Stuff you're good at.”

Leon swallowed and looked at Ria. There was nothing on her face. “I 
can't help but notice,” he said, forcing his voice to its absolutely 
calmest, “that you haven't mentioned anything to do with the, uh, 
\emph{clients}.”

Brautigan nodded and strained to pull his lips over his horsey teeth to 
hide his grin. It didn't work. “Yeah,” he said. “That's about 
right. We need someone of your talents doing what he does best, and 
what you do best is --”

He held up a hand. Brautigan fell silent. The three of them looked at 
him. He realized, in a flash, that he had them all in his power, just 
at that second. He could shout BOO! and they'd all fall off their 
chairs. They were waiting to see if he'd blow his top or take it and 
ask for more. He did something else.

“Nice working with ya,” he said. And he turned his back on the 
sweetest, softest job anyone could ask for. He said \emph{adios} and 
\emph{buena suerte} to Carmela on the way out, and he forced himself 
not to linger around the outside doors down at street level to see if 
anyone would come chasing after him.

\tb

The realtor looked at him like he was crazy. “You'll never get two 
million for that place in today's market,” she said. She was young, 
no-nonsense, black, and she had grown up on the Lower East Side, a fact 
she mentioned prominently in her advertising materials: \emph{A local 
realtor for a local neighborhood.}

“I paid two million for it less than a year ago,” he said. The 80 
percent mortgage had worried him a little but Ate had underwritten it, 
bringing the interest rate down to less than two percent.

She gestured at the large corner picture window that overlooked Broome 
Street and Grand Street. “Count the FOR SALE signs,” she said. “I 
want to be on your side. That's a nice place. I'd like to see it go to 
someone like you, someone decent. Not some \emph{developer} --” she 
spat the word like a curse “or some corporate apartment broker who'll 
rent it by the week to VIPs. This neighborhood needs real people who 
really live here, understand.”

“So you're saying I won't get what I paid for it?”

She smiled fondly at him. “No, sweetheart, you're not going to get 
what you paid for it. All those things they told you when you put two 
mil into that place, like `They're not making any more Manhattan' and 
`Location location location'? It's lies.” Her face got serious, 
sympathetic. “It's supposed to panic you and make you lose your head 
and spend more than you think something is worth. That goes on for a 
while and then everyone ends up with too much mortgage for not enough 
home, or for too much home for that matter, and then blooie, the bottom 
blows out of the market and everything falls down like a souffle.”

“You don't sugar-coat it, huh?” He'd come straight to her office 
from Ate's door, taking the subway rather than cabbing it or even 
renting a jetpack. He was on austerity measures, effective immediately. 
His brain seemed to have a pre-made list of cost-savers it had prepared 
behind his back, as though it knew this day would come.

She shrugged. “I can, if you want me to. We can hem and haw about the 
money and so on and I can hold your hand through the seven stages of 
grieving. I do that a lot when the market goes soft. But you looked 
like the kind of guy who wants it straight. Should I start over? Or, 
you know, if you want, we can list you at two mil or even two point 
two, and I'll use that to prove that some \emph{other} loft is a steal 
at 1.9. If you want.”

“No,” he said, and he felt some of the angry numbness ebb away. He 
liked this woman. She had read him perfectly. “So tell me what you 
think I can get for it?”

She put her fist under her chin and her eyes went far away. “I sold 
that apartment, um, eight years ago? Family who had it before you. Had 
a look when they sold it to you -- they used a different broker, kind 
of place where they don't mind selling to a corporate placement 
specialist. I don't do that, which you know. But I saw it when it sold. 
Have you changed it much since?”

He squirmed. “I didn't, but I think the broker did. It came 
furnished, nice stuff.”

She rolled her eyes eloquently. “It's never nice stuff. Even when it 
comes from the best showroom in town, it's not nice stuff. Nice is 
antithetical to corporate. Inoffensive is the best you can hope for.” 
She looked up, to the right, back down. “I'm figuring out the 
discount for how the place will show now that they've taken all the 
seams and crumbs out. I'm thinking, um, 1.8. That's a number I think I 
can deliver.”

“But I've only \emph{got} 200K in the place,” he said.

Her expressive brown eyes flicked at the picture window, the FOR SALE 
signs. “And? Sounds like you'll break even or maybe lose a little on 
the deal. Is that right?”

He nodded. Losing a little wasn't something he'd figured on. But by the 
time he'd paid all the fees and taxes -- “I'll probably be down a 
point or two.”

“Have you got it?”

He hated talking about money. That was one thing about Ria is that she 
never actually talked about money -- what money \emph{did}, sure, but 
never money. “Technically,” he said.

“OK, technical money is as good as any other kind. So look at it this 
way: you bought a place, a really totally amazing place on the Lower 
East Side, a place bigger than five average New York apartments. You 
lived in it for, what?”

“Eight months.”

“Most of a year. And it cost you one percent of the street price on 
the place. Rent would have been about eleven times that. You're up 
--” she calculated in her head -- “it's about 83 percent.”

He couldn't keep the look of misery off his face.

“What?” she said. “Why are you pulling faces at me? You said you 
didn't want it sugar-coated, right?”

“It's just that --” He dropped his voice, striving to keep any kind 
of whine out of it. “Well, I'd hoped to make something in the 
bargain.”

“For what?” she said, softly.

“You know, appreciation. Property goes up.”

“Did you do anything to the place that made it better?”

He shook his head.

“So you did no productive labor but you wanted to get paid anyway, 
right? Have you thought about what would happen to society if we 
rewarded people for owning things instead of doing things?”

“Are you sure you're a real estate broker?”

“Board certified. Do very well, too.”

He swallowed. “I don't expect to make money for doing nothing, but 
you know, I just quit my job. I was just hoping to get a little cash in 
hand to help me smooth things out until I find a new one.”

The realtor gave a small nod. “Tough times ahead. Winds are about to 
shift again. You need to adjust your expectations, Leon. The best you 
can hope for right now is to get out of that place before you have to 
make another mortgage payment.”

His pulse throbbed in his jaw and his thigh in counterpoint. “But I 
\emph{need} money to --”

“Leon,” she said, with some steel in her voice. “You're 
\emph{bargaining}. As in denial, anger, bargaining, depression and 
acceptance. That's healthy and all, but it's not going to get your 
place sold. Here's two options: one, you can go find another realtor, 
maybe one who'll sugar coat things or string you along to price up 
something else he's trying to sell. Two, you can let me get on with 
making some phone calls and I'll see who I can bring in. I keep a list 
of people I'd like to see in this `hood, people who've asked me to look 
out for the right kind of place. That place you're in is one of a kind. 
I might be able to take it off your hands in very quick time, if you 
let me do my thing.” She shuffled some papers. “Oh, there's a 
third, which is that you could go back to your apartment and pretend 
that nothing is wrong until that next mortgage payment comes out of 
your bank-account. That would be \emph{denial} and if you're 
bargaining, you should be two steps past that.

“What's it going to be?”

“I need to think about it.”

“Good plan,” she said. “Remember, depression comes after 
bargaining. Go buy a quart of ice-cream and download some weepy movies. 
Stay off booze, it only brings you down. Sleep on it, come back in the 
morning if you'd like.”

He thanked her numbly and stepped out into the Lower East Side. The 
bodega turned out to have an amazing selection of ice-cream, so he 
bought the one with the most elaborated name, full of chunks, swirls 
and stir-ins, and brought it up to his apartment, which was so big that 
it made his knees tremble when he unlocked his door. The realtor had 
been right. Depression was next.

\tb

Buhle sent him an invitation a month later. It came laser-etched into a 
piece of ancient leather, delivered by a messenger whose jetpack was so 
quiet that he didn't even notice that she had gone until he looked up 
from the scroll to thank her. His new apartment was a perch he rented 
by the week at five times what an annual lease would have cost him, but 
still a fraction of what he had been paying on the LES. It was jammed 
with boxes of things he hadn't been able to bring himself to get rid 
of, and now he cursed every knick-knack as he dug through them looking 
for a good suit.

He gave up. The invitation said, “At your earliest convenience,” 
and a quadrillionaire in a vat wasn't going to be impressed by his 
year-old designer job-interview suit.

It had been a month, and no one had come calling. None of his queries 
to product design, marketing, R{\&}D or advertising shops had been 
answered. He tried walking in the park every day, to see the bears, on 
the grounds that it was free and it would stimulate his creative flow. 
Then he noticed that every time he left his door, fistfuls of money 
seemed to evaporate from his pockets on little “necessities” that 
added up to real money. The frugality center of his brain began to 
flood him with anxiety every time he considered leaving the place and 
so it had been days since he'd gone out.

Now he was going. There were some clean clothes in one of the boxes, 
just sloppy jeans and tees, but they'd been expensive sloppy once upon 
a time, and they were better than the shorts and shirts he'd been 
rotating in and out of the tiny washing machine every couple days, when 
the thought occurred to him. The \$200 haircut he'd had on his last day 
of work had gone shaggy and lost all its clever style, so he just 
combed it as best as he could after a quick shower and put on his 
architect's shoes, shining them on the backs of his pants legs on his 
way out the door in a gesture that reminded him of his father going to 
work in Anguilla, a pathetic gesture of respectability from someone who 
had none. The realization made him \emph{oof} out a breath like he'd 
been gut-punched.

His frugality gland fired like crazy as he hailed a taxi and directed 
it to the helipad at Grand Central Terminus. It flooded him with so 
much cheapamine that he had to actually pinch his arms a couple times 
to distract himself from the full-body panic at the thought of spending 
so much. But Buhle was all the way in Rhode Island, and Leon didn't 
fancy keeping him waiting. He knew that to talk to money you had to act 
like money -- impedance-match the money. Money wouldn't wait while he 
took the train or caught the subway.

He booked the chopper-cab from the cab, using the terminal in the 
back-seat. At Ate, he'd had Carmela to do this kind of organizing for 
him. He'd had Carmela to do a hundred other things, too. In that 
ancient, lost time, he'd had money and help beyond his wildest dreams, 
and most days he couldn't imagine what had tempted him into giving it 
up.

The chopper clawed the air and lifted him up over Manhattan, the 
canyons of steel stretched out below him like a model. The racket of 
the chopper obliterated any possibility of speech, so he could ignore 
the pilot and she could ignore him with a cordiality that let him 
pretend, for a moment, that he was a powerful executive who 
nonchalantly choppered around over the country. They hugged the 
coastline and the stately rows of windmills and bobbing float-homes, 
surfers carving the waves, bulldozed strips topped with levees that 
shot up from the ground like the burial mound of some giant serpent.

Leon's earmuffs made all the sound -- the sea, the chopper -- into a 
uniform hiss, and in that hiss, his thoughts and fears seemed to recede 
for a moment, as though they couldn't make themselves heard over the 
white noise. For the first time since he'd walked out of Ate, the 
nagging, doubtful voices fell still and Leon was alone in his head. It 
was as though he'd had a great pin stuck through his chest that had 
been finally removed. There was a feeling of lightness, and tears 
pricking at his eyes, and a feeling of wonderful \emph{obliteration}, 
as he stopped, just for a moment, stopped trying to figure out where he 
fit in the world.

The chopper touched down on a helipad at Newport State Airport, to one 
side of the huge X slashed into the heavy woods -- new forest, 
fast-growing carbon sinkers garlanded with extravagances of moss and 
vine. From the moment the doors opened, the heavy earthy smell filled 
his nose and he thought of the Living Room, which led him to think of 
Ria. He thanked the pilot and zapped her a tip and looked up and there 
was Ria, as though his thoughts had summoned her.

She had a little half-smile on her face, uncertain and somehow 
childlike, a little girl waiting to find out if he'd be her friend 
still. He smiled at her, grateful for the clatter of the chopper so 
that they couldn't speak. She shook his hand, hers warm and dry, and 
then, on impulse, he gave her a hug. She was soft and firm too, a 
middle-aged woman who kept fit but didn't obsess about the pounds. It 
was the first time he'd touched another human since he left Ate. And, 
as with the chopper's din, this revelation didn't open him to fresh 
miseries -- rather, it put the miseries away, so that he felt 
\emph{better}.

“Are you ready?” she said, once the chopper had lifted off.

“One thing,” he said. “Is there a town here? I thought I saw one 
while we were landing.”

“A little one,” she said. “Used to be bigger, but we like them 
small.”

“Does it have a hardware store?”

She gave him a significant look. “What for? An axe? A nailgun? Going 
to do some improvements?”

“Thought I'd bring along a door-knob,” he said.

She dissolved into giggles. “Oh, he'll \emph{like} that. Yes, we can 
find a hardware store.”

\tb

Buhle's security people subjected the doorknob to millimeter radar and 
a gas chromatograph before letting it past. He was shown into an 
anteroom by Ria, who talked to him through the whole procedure, just 
light chatter about the weather and his real-estate problems, but she 
gently steered him around the room, changing their angle several times, 
and then he said, “Am I being scanned?”

“Millimeter radar in here too,” she said. “Whole body imaging. 
Don't worry, I get it every time I come in. Par for the course.”

He shrugged. “This is the least offensive security scan I've ever 
been through,” he said.

“It's the room,” she said. “The dimensions, the color. Mostly the 
semiotics of a security scan are either \emph{you are a germ on a 
slide} or \emph{you are not worth trifling with, but if we must, we 
must.} We went for something a little... sweeter.” And it was, a 
sweet little room, like the private study of a single mom who's stolen 
a corner in which to work on her secret novel.

Beyond the room -- a wonderful place.

“It's like a college campus,” he said.

“Oh, I think we use a better class of materials than most 
colleges,” Ria said, airily, but he could tell he'd pleased her. 
“But yes, there's about 15,000 of us here. A little city. Nice cafes, 
gyms, cinemas. A couple artists in residence, a nice little Waldorf 
school...” The pathways were tidy and wended their way through 
buildings ranging from cottages to large, institutional buildings, but 
all with the feel of endowed research institutes rather than finance 
towers. The people were young and old, casually dressed, walking in 
pairs and groups, mostly, deep in conversation.

“15,000?”

“That's the head office. Most of them doing medical stuff here. We've 
got lots of other holdings, all around the world, in places that are 
different from this. But we're bringing them all in line with HQ, fast 
as we can. It's a good way to work. Churn is incredibly low. We 
actually have to put people back out into the world for a year every 
decade, just so they can see what it's like.”

“Is that what you're doing?”

She socked him in the arm. “You think I could be happy here? No, I've 
always lived off campus. I commute. I'm not a team person. It's OK, 
this is the kind of place where even lone guns can find their way to 
glory.”

They were walking on the grass now, and he saw that the trees, 
strangely oversized red maples without any of the whippy slenderness he 
associated with the species, had a walkway suspended from their 
branches, a real Swiss Family Robinson job with rope-railings and 
little platforms with baskets on pulleys for ascending and descending. 
The people who scurried by overhead greeted each other volubly and 
laughed at the awkwardness of squeezing past each other in opposite 
directions.

“Does that ever get old?” he said, lifting his eyebrows to the 
walkways.

“Not for a certain kind of person,” she said. “For a certain kind 
of person, the delightfulness of those walkways never wears off.” The 
way she said “certain kind of person” made him remember her saying, 
“bears shouldn't be that happy.”

He pointed to a bench, a long twig-chair, really, made from birch 
branches and rope and wire all twined together. “Can we sit for a 
moment? I mean, will Buhle mind?”

She flicked her fingers. “Buhle's schedule is his own. If we're five 
minutes late, someone will put five minutes' worth of interesting and 
useful injecta into his in-box. Don't you worry.” She sat on the 
bench, which looked too fragile and fey to take a grown person's 
weight, but then she patted the seat next to him, and when he sat, he 
felt almost no give. The bench had been very well built, by someone who 
knew what she or he was doing.

“OK, so what's going on, Ria? First you went along with Brautigan 
scooping my job and exiling me to Siberia --” he held up a hand to 
stop her from speaking and discovered that the hand was shaking and so 
was his chest, shaking with a bottled-up anger he hadn't dared admit. 
“You could have stopped it at a word. You envoys from the vat-gods, 
you are the absolute monarchs at Ate. You could have told them to have 
Brautigan skinned, tanned and made into a pair of boots, and he'd have 
measured your foot-size himself. But you let them do it.

“And now, here I am, a minister without portfolio, about to do 
something that would make Brautigan explode with delight, about to meet 
one of the Great Old Ones, in his very vat, in person. A man who might 
live to be a thousand, if all goes according to plan, a man who is a 
\emph{country}, sovereign and inviolate. And I just want to ask you, 
\emph{why}? Why all the secrecy and obliqueness and funny gaps? 
\emph{Why}?”

Ria waited while a pack of grad students scampered by overhead, deep in 
discussion of teleomeres, the racket of their talk and their bare feet 
slapping on the walkway loud enough to serve as a pretence for silence. 
Leon's pulse thudded and his armpits slicked themselves as he realized 
that he might have just popped the bubble of unreality between them, 
the consensual illusion that all was normal, whatever normal was.

“Oh, Leon,” she said. “I'm sorry. Habit here -- there's some 
things that can't be readily said in utopia. Eventually, you just get 
in the habit of speaking out of the back of your head. It's, you know, 
\emph{rude} to ruin people's gardens by pointing out the snakes. So, 
yes, OK, I'll say something right out. I like you, Leon. The average 
employee at a place like Ate is a bottomless well of desires, trying to 
figure out what others might desire. We've been hearing from them for 
decades now, the resourceful ones, the important ones, the ones who 
could get past the filters and the filters behind the filters. We know 
what they're like.

“Your work was different. As soon as you were hired by Ate, we 
generated a dossier on you. Saw your grad work.”

Leon swallowed. His resume emphasized his grades, not his final 
projects. He didn't speak of them at all.

“So we thought, well, here's something different, it's possible he 
may have a house to go with our doorknob. But we knew what would happen 
if you were left to your own devices at a place like Ate: they'd bend 
you and shape you and make you over or ruin you. We do it ourselves, 
all too often. Bring in a promising young thing, subject him to the 
dreaded Buhle Culture, a culture he's totally unsuited to, and he 
either runs screaming or... \emph{fits in}. It's worse when the latter 
happens. So we made sure that you had a good fairy perched on your 
right shoulder to counterbalance the devil on your left shoulder.” 
She stopped, made a face, mock slapped herself upside the head. 
“Talking in euphemism again. Bad habit. You see what I mean.”

“And you let me get pushed aside...”

She looked solemn. “We figured you wouldn't last long as a 
button-polisher. Figured you'd want out.”

“And that you'd be able to hire me.”

“Oh, we could have hired you any time. We could have bought Ate. Ate 
would have given you to us -- remember all that business about making 
Brautigan into a pair of boots? It applies all around.”

“So you wanted me to... what? Walk in the wilderness first?”

“Now you're talking in euphemisms. It's catching! Let's walk.”

\tb

They gave him a bunny-suit to wear into the heart of Buhle. First he 
passed through a pair of double-doors, faintly positively pressurized, 
sterile air that ruffled his hair on the way in. The building was 
low-slung, nondescript brown brick, no windows. It could have been a 
water sterilization plant or a dry-goods warehouse. The inside was good 
tile, warm colors with lots of reds and browns down low, making the 
walls look like they were the inside of a kiln. The building's interior 
was hushed, and a pair of alert-looking plainclothes security men 
watched them very closely as they changed into the bunny suits, loose 
micropore coveralls with plastic visors. Each one had a small, 
self-contained air-circ system powered by a wrist canister, and when a 
security man helpfully twisted the valve open, Leon noted that there 
were clever jets that managed to defog the visor without drying out his 
eyeballs.

“That be enough for you, Ria?” the taller of the two security men 
said. He was dressed like a college kid who'd been invited to his 
girlfriend's place for dinner: smart slacks a little frayed at the 
cuffs, a short-sleeved, pressed cotton shirt that showed the bulge of 
his substantial chest and biceps and neck.

She looked at her canister, holding it up to the visor. “30 minutes 
is fine,” she said. “I doubt he'll have any more time than that for 
us!” Turning to Leon, she said, “I think that the whole air-supply 
thing is way overblown. But it does keep meetings from going long.”

“Where does the exhaust go?” Leon said, twisting in his suit. “I 
mean, surely the point is to keep my cooties away from,” he 
swallowed, “\emph{Buhle}.” It was the first time he'd really used 
the word to describe a person, rather than a \emph{concept}, and he was 
filled with the knowledge that the person it described was somewhere 
very close.

“Here,” she said, and pointed to a small bubble growing out of the 
back of her neck. “You swell up, one little bladder at a time, until 
you look like the Michelin man. Some joke.” She made a face. “You 
can get a permanent suit if you come here often. Much less awkward. But 
Buhle likes it awkward.”

She led him down a corridor with still more people, these ones in 
bunny-suits or more permanent-looking suits that were form-fitting and 
iridescent and flattering. “Really?” he said, keeping pace with 
her. “Elegant is a word that comes to mind, not awkward.”

“Well, sure, elegant on the other side of that airlock door. But 
we're inside Buhle's body now.” She saw the look on his face and 
smiled. “No, no, it's not a riddle. Everything on this side of the 
airlock is Buhle. It's his lungs and circulatory and limbic system. The 
vat may be where the meat sits, but all this is what makes the vat 
work. You're like a gigantic foreign organism that's burrowing into his 
tissues. It's intimate.” They passed through another set of doors and 
now they were almost alone in a hall the size of his university's 
basketball court, the only others a long way off. She lowered her voice 
so that he had to lean in to hear her. “When you're outside, speaking 
to Buhle through his many tendrils, like me, or even on the phone, he 
has all the power in the world. He's a giant. But here, inside his 
body, he's very, very weak. The suits, they're there to level out the 
playing field. It's all head-games and symbolism. And this is just Mark 
I, the system we jury-rigged after Buhle's... \emph{accident}. They're 
building the Mark II about five miles from here, and half a mile 
underground. When it's ready, they'll blast a tunnel and take him all 
the way down into it without ever compromising the skin of Buhle's 
extended body.”

“You never told me what the accident was, how he ended up here. I 
assumed it was a stroke or --”

Ria shook her head, the micropore fabric rustling softly. “Nothing 
like that,” she said.

They were on the other side of the great room now, headed for the 
doors. “What is this giant room for?”

“Left over from the original floor-plan, when this place was just 
biotech R{\&}D. Used for all-hands meetings then, sometimes a little 
symposium. Too big now. Security protocol dictates no more than ten 
people in any one space.”

“Was it assassination?” He said it without thinking, quick as 
ripping off a band-aid.

Again, the rustle of fabric. “No.”

She put her hand on the door's crashbar, made ready to pass into the 
next chamber.

“I'm starting to freak out a little here, Ria,” he said. “He 
doesn't hunt humans or something?”

“No,” and he didn't need to see her face, he could see the smile.

“Or need an organ? I don't think I have a rare blood-type, and I 
should tell you that mine have been indifferently cared for --”

“Leon,” she said, “if Buhle needed an organ, we'd make one right 
here. Print it out in about forty hours, pristine and virgin.”

“So you're saying I'm not going to be harvested or hunted, then?”

“It's a very low probability outcome,” she said, and pushed the 
crashbar.

It was darker in this room, a mellow, candlelit sort of light, and 
there was a rhythmic vibration coming up through the floor, a whoosh 
whoosh.

Ria said, “It's his breath. The filtration systems are down there.” 
She pointed a toe at the outline of a service hatch set into the floor. 
“Circulatory system overhead,” she said, and he craned his neck up 
at the grate covering the ceiling, the troughs filled with neatly 
bundled tubes.

One more set of doors, another cool, dark room, this one nearly silent, 
and one more door at the end, an airlock door, and another plainclothes 
security person in front of it; a side-room with a glass door bustling 
with people staring intently at screens. The security person -- a 
woman, Leon saw -- had a frank and square pistol with a bulbous butt 
velcroed to the side of her suit.

“He's through there, isn't he?” Leon said, pointing at the airlock 
door.

“No,” Ria said. “No. He's here. We are inside him. Remember that, 
Leon. He isn't the stuff in the vat there. In some sense you've been in 
Buhle's body since you got off the chopper. His sensor array network 
stretches out as far as the heliport, like the tips of the hairs on 
your neck, they feel the breezes that blow in his vicinity. Now you've 
tunneled inside him, and you're right here, in his heart or his 
liver.”

“Or his brain.”

A voice, then, from everywhere, warm and good-humored. “The brain is 
overrated.” Leon looked at Ria and she rolled her eyes eloquently 
behind her faceplate.

“Tuned sound,” she said. “A party trick. Buhle --”

“Wait,” Buhle said. “Wait. The brain, this is important, the 
brain is \emph{so} overrated. The ancient Egyptians thought it was used 
to cool the blood, you know that?” He chortled, a sound that felt to 
Leon as though it began just above his groin and rose up through his 
torso, a very pleasant and very invasive sensation. “The heart, they 
thought, the heart was the place where the \emph{me} lived. I used to 
wonder about that. Wouldn't they think that the thing between the 
organs of hearing, the thing behind the organs of seeing, that must be 
the me? But that's just the brain doing one of its little stupid games, 
backfilling the explanation. We think the brain is the obvious seat of 
the me because the brain already knows that it is the seat, and can't 
conceive of anything else. When the brain thought it lived in your 
chest, it was perfectly happy to rationalize that too -- \emph{Of 
course it's in the chest, you feel your sorrow and your joy there, your 
satiety and your hunger...} The brain, pffft, the brain!”

“Buhle,” she said. “We're coming in now.”

The nurse/guard by the door had apparently only heard their part of the 
conversation, but also hadn't let it bother her. She stood to one side, 
and offered Leon a tiny, incremental nod as he passed. He returned it, 
and then hurried to catch up with Ria, who was waiting inside the 
airlock. The outer door closed and for a moment, they were pressed up 
against one another and he felt a wild, horny thought streak through 
him, all the excitement discharging itself from yet another place that 
the me might reside.

Then the outer door hissed open and he met Buhle -- he tried to 
remember what Ria had said, that Buhle wasn't this, Buhle was 
everywhere, but he couldn't help himself from feeling that this was 
\emph{him}.

Buhle's vat was surprisingly small, no bigger than the sarcophagus that 
an ancient Egyptian might have gone to in his burial chamber. He tried 
not to stare inside it, but he couldn't stop himself. The withered, 
wrinkled man floating in the vat was intertwined with a thousand fiber 
optics that disappeared into pinprick holes in his naked skin. There 
were tubes: in the big highways in the groin, in the gut through a 
small valve set into a pucker of scar, in the nose and ear. The 
hairless head was pushed in on one side, like a pumpkin that hasn't 
been turned as it grew in the patch, and there was no skin on the flat 
piece, only white bone and a fine metallic mesh and more ragged, 
curdled scar tissue.

The eyes were hidden behind a slim set of goggles that irised open when 
they neared him, and beneath the goggles they were preternaturally 
bright, bright as marbles, set deep in bruised-looking sockets. The 
mouth beneath the nostril-tubes parted in a smile, revealing teeth as 
neat and white as a toothpaste advertisement, and Buhle spoke.

“Welcome to the liver. Or the heart.”

Leon choked on whatever words he'd prepared. The voice was the same one 
he'd heard in the outer room, warm and friendly, the voice of a man 
whom you could trust, who would take care of you. He fumbled around his 
suit, patting it. “I brought you a doorknob,” he said, “but I 
can't reach it just now.”

Buhle laughed, not the chuckle he'd heard before, but an actual, barked 
\emph{Ha!} that made the tubes heave and the fiber optics writhe. 
“Fantastic,” he said. “Ria, he's fantastic.”

The compliment made the tips of Leon's ears grow warm.

“He's a good one,” she said. “And he's come a long way at your 
request.”

“You hear how she reminds me of my responsibilities? Sit down, both 
of you.” Ria rolled over two chairs, and Leon settled into one, 
feeling it noiselessly adjust to take his weight. A small mirror 
unfolded itself and then two more, angled beneath it, and he found 
himself looking into Buhle's eyes, looking at his face, reflected in 
the mirrors.

“Leon,” Buhle said, “tell me about your final project, the one 
that got you the top grade in your class.”

Leon's fragile calm vanished, and he began to sweat. “I don't like to 
talk about it,” he said.

“Makes you vulnerable, I know. But vulnerable isn't so bad. Take me. 
I thought I was invincible. I thought that I could make and unmake the 
world to my liking. I thought I understood how the human mind worked -- 
and how it broke.

“And then one day in Madrid, as I was sitting in my suite's breakfast 
room, talking with an old friend while I ate my porridge oats, my old 
friend picked up the heavy silver coffee jug, leaped on my chest, 
smashed me to the floor, and methodically attempted to beat the brains 
out of my head with it. It weighed about three pounds, not counting the 
coffee, which was scalding, and she only got in three licks before they 
pulled her off of me, took her away. Those three licks though --” He 
looked intently at them. “I'm an old man,” he said. “Old bones, 
old tissues. The first blow cracked my skull. The second one broke it. 
The third one forced fragments into my brain. By the time the medics 
arrived, I'd been technically dead for about 174 seconds, give or take 
a second or two.”

Leon wasn't sure the old thing in the vat had finished speaking, but 
that seemed to be the whole story. “Why?” he said, picking the word 
that was uppermost in his mind.

“Why did I tell you this?”

“No,” Leon said. “Why did your old friend try to kill you?”

Buhle grinned. “Oh, I expect I deserved it,” he said.

“Are you going to tell me why?” Leon said.

Buhle's cozy grin disappeared. “I don't think I will.”

Leon found he was breathing so hard that he was fogging up his 
faceplate, despite the air-jets that worked to clear it. “Buhle,” 
he said, “the point of that story was to tell me how vulnerable you 
are so that I'd tell you my story, but that story doesn't make you 
vulnerable. You were beaten to death and yet you survived, grew 
stronger, changed into this --” He waved his hands around. “This 
body, this monstrous, town-sized giant. You're about as vulnerable as 
fucking Zeus.”

Ria laughed softly but unmistakably. “Told you so,” she said to 
Buhle. “He's a good one.”

The exposed lower part of Buhle's face clenched like a fist and the 
pitch of the machine noises around them shifted a half-tone. Then he 
smiled a smile that was visibly forced, obviously artificial even in 
that ruin of a face.

“I had an idea,” he said. “That many of the world's problems 
could be solved with a positive outlook. We spend so much time worrying 
about the rare and lurid outcomes in life. Kids being snatched. 
Terrorists blowing up cities. Stolen secrets ruining your business. 
Irate customers winning huge judgments in improbable lawsuits. All this 
\emph{chickenshit}, bed-wetting, hand-wringing \emph{fear}.” His 
voice rose and fell like a minister's and it was all Leon could do not 
to sway in time with him. “And at the same time, we neglect the 
likely: traffic accidents, jetpack crashes, bathtub drownings. It's 
like the mind can't stop thinking about the grotesque, and can't stop 
forgetting about the likely.”

“Get on with it,” Ria said. “The speech is lovely, but it doesn't 
answer the question.”

He glared at her through the mirror, the marble-eyes in their mesh of 
burst blood vessels and red spider-tracks, like the eyes of a demon. 
“The human mind is just \emph{kinked wrong}. And it's correctable.” 
The excitement in his voice was palpable. “Imagine a product that let 
you \emph{feel} what you \emph{know} -- imagine if anyone who heard 
`Lotto: you've got to be in it to win it' immediately understood that 
this is \emph{so much bullshit} That statistically, your chances of 
winning the lotto are not measurably improved by buying a lottery 
ticket. Imagine if explaining the war on terror to people made them 
double over with laughter! Imagine if the capital markets ran on 
realistic assessments of risk instead of envy, panic and greed.”

“You'd be a lot poorer,” Ria said.

He rolled his eyes eloquently.

“It's an interesting vision,” Leon said. “I'd take the cure, 
whatever it was.”

The eyes snapped to him, drilled through him, fierce. “That's the 
problem, \emph{right there}. The only people who'll take this are the 
people who don't need it. Politicians and traders and oddsmakers know 
how probability works, but they also know that the people who make them 
fat and happy \emph{don't} understand it a bit, and so they can't 
afford to be rational. So there's only one answer to the problem.”

Leon blurted, “The bears.”

Ria let out an audible sigh.

“The fucking bears,” Buhle agreed, and the way he said it was so 
full of world-weary exhaustion that it made Leon want to hug him. 
“Yes. As a social reform tool, we couldn't afford to leave this to 
the people who were willing to take it. So we --”

“Weaponized it,” Ria said.

“Whose story is this?”

Leon felt that the limbs of his suit were growing stiffer, his exhaust 
turning it into a balloon. And he had to pee. And he didn't want to 
move.

“You dosed people with it?”

“Leon,” Buhle said, in a voice that implied, \emph{Come on, we're 
bigger than that}. “They'd consented to being medical research 
subjects. And it \emph{worked}. They stopped running around shouting 
\emph{The sky is falling, the sky is falling} and became -- \emph{zen}. 
Happy, in a calm, even-keeled way. Headless chickens turned into 
flinty-eyed air-traffic controllers.”

“And your best friend beat your brains in --”

“Because,” Buhle said, in a little Mickey Mouse falsetto, 
“\emph{it would be unethical to do a broad-scale release on the 
general public}”

Ria was sitting so still he had almost forgotten she was there.

Leon shifted his weight. “I don't think that you're telling me the 
whole story.”

“We were set to market it as an anti-anxiety medication.”

“And?”

Ria stood up abruptly. “I'll wait outside.” She left without 
another word.

Buhle rolled his eyes again. “How do you get people to take 
anti-anxiety medication? Lots and lots of people? I mean, if I assigned 
you that project, gave you a budget for it --”

Leon felt torn between a desire to chase after Ria and to continue to 
stay in the magnetic presence of Buhle. He shrugged. “Same as you 
would with any pharma. Cook the diagnosis protocol, expand the number 
of people it catches. Get the news media whipped up about the anxiety 
epidemic. That's easy. Fear sells. An epidemic of fear? Christ, that'd 
be too easy. Far too easy. Get the insurers on board, discounts on the 
meds, make it cheaper to prescribe a course of treatment than to take 
the call-center time to explain to the guy why he's \emph{not} getting 
the meds.”

“You're my kind of guy, Leon,” Buhle said. “So yeah.”

“Yeah?”

Another one of those we're-both-men-of-the-world smiles. “Yeah.”

Oh.

“How many?”

“That's the thing. We were trying it in a little market first. Basque 
country. The local authority was very receptive. Lots of chances to 
fine-tune the message. They're the most media-savvy people on the 
planet these days -- they are to media as the Japanese were to 
electronics in the last century. If we could get them in the door --”

“How many?”

“About a million. More than half the population.”

“You created a bioweapon that infected its victims with numeracy, and 
infected a million Basque with it?”

“Crashed the lottery. That's how I knew we'd done it. Lottery tickets 
fell by more than 80 percent. Wiped out.”

“And then your friend beat your head in?”

“Well.”

The suit was getting more uncomfortable by the second. Leon wondered if 
he'd get stuck if he waited too long, his overinflated suit incapable 
of moving. “I'm going to have to go, soon.”

“Evolutionarily, bad risk-assessment is advantageous.”

Leon nodded slowly. “OK, I'll buy that. Makes you entrepreneurial 
--”

“Drives you to colonize new lands, to ask out the beautiful monkey in 
the next tree, to have a baby you can't imagine how you'll afford.”

“And your numerate Vulcans stopped?”

“Pretty much,” he said. “But that's just normal shakedown. Like 
when people move to cities, their birthrate drops. And nevertheless, 
the human race is becoming more and more citified and still, it isn't 
vanishing. Social stuff takes time.”

“And then your friend beat your head in?”

“Stop saying that.”

Leon stood. “Maybe I should go and find Ria.”

Buhle made a disgusted noise. “Fine. And ask her why she didn't 
finish the job? Ask her if she decided to do it right then, or if she'd 
planned it? Ask her why she used the coffee jug instead of the 
bread-knife? Because, you know, I wonder this myself.”

Leon backpedaled, clumsy in the overinflated suit. He struggled to get 
into the airlock, and as it hissed through its cycle, he tried not to 
think of Ria straddling the old man's chest, the coffee urn rising and 
falling.

She was waiting for him on the other side, also overinflated in her 
suit. “Let's go,” she said, and took his hand, the rubberized palms 
of their gloves sticking together. She half-dragged him through the 
many rooms of Buhle's body, tripping through the final door, then 
spinning him around and ripping, hard, on the release cord that split 
the suit down the back so that it fell into two lifeless pieces that 
slithered to the ground. He gasped out a breath he hadn't realized he'd 
been holding in as the cool air made contact with the thin layer of 
perspiration that filmed his body.

Ria had already ripped open her own suit and her face was flushed and 
sweaty, her hair matted. Small sweat-rings sprouted beneath her 
armpits. An efficient orderly came forward and began gathering up their 
suits. Ria thanked her impersonally and headed for the doors.

“I didn't think he'd do that,” she said, once they were outside of 
the building -- outside the core of Buhle's body.

“You tried to kill him,” Leon said. He looked at her hands, which 
had blunt, neat fingernails and large knuckles. He tried to picture the 
tendons on their backs standing out like sail-ropes when the wind blew, 
as they did the rhythmic work of raising and lowering the heavy silver 
coffee pot.

She wiped her hands on her trousers and stuffed them in her pockets, 
awkward now, without any of her usual self-confidence. “I'm not 
ashamed of that. I'm proud of it. Not everyone would have had the guts. 
If I hadn't, you and everyone you know would be --” She brought her 
hands out of her pockets, bunched into fists. She shook her head. “I 
thought he'd tell you what we like about your grad project. Then we 
could have talked about where you'd fit in here --”

“You never said anything about that,” he said. “I could have 
saved you a lot of trouble. I don't talk about it.”

Ria shook her head. “This is Buhle. You won't stop us from doing 
anything we want to do. I'm not trying to intimidate you here. It's 
just a fact of life. If we want to replicate your experiment, we can, 
on any scale we want --”

“But I won't be a part of it,” he said. “That matters.”

“Not as much as you think it does. And if you think you can avoid 
being a part of something that Buhle wants you for, you're likely to be 
surprised. We can get you what you want.”

“No you can't,” he said. “If there's one thing I know, it's that 
you can't do that.”

\tb

Take one normal human being at lunch. Ask her about her breakfast. If 
lunch is great, she'll tell you how great breakfast is. If lunch is 
terrible, she'll tell you how awful breakfast was.

Now ask her about dinner. A bad lunch will make her assume that a bad 
dinner is forthcoming. A great lunch will make her optimistic about 
dinner.

Explain this dynamic to her and ask her again about breakfast. She'll 
struggle to remember the actual details of breakfast, the texture of 
the oatmeal, whether the juice was cold and delicious or slightly warm 
and slimy. She will remember and remember and remember for all she's 
worth, and then, if lunch is good, she'll tell you breakfast was good. 
And if lunch is bad, she'll tell you breakfast was bad.

Because you just can't help it. Even though you know you're doing it, 
you can't help it.

But what if you could?

\tb

“It was the parents,” he said, as they picked their way through the 
treetops, along the narrow walkway, squeezing to one side to let the 
eager, gabbling researchers past. “That was the heartbreaker. Parents 
only remember the good parts of parenthood. Parents whose kids are 
grown remember a succession of sweet hugs, school triumphs, sports 
victories, and they simply forget the vomit, the tantrums, the sleep 
deprivation... It's the thing that lets us continue the species, this 
excellent facility for forgetting. That's what should have tipped me 
off.”

Ria nodded solemnly. “But there was an upside, wasn't there?”

“Oh, sure. Better breakfasts, for one thing. And the weight-loss -- 
amazing. Just being able to remember how shitty you felt the last time 
you ate the chocolate bar or pigged out on fries. It was amazing.”

“The applications do sound impressive. Just that weight-loss one --”

“Weight-loss, addiction counseling, you name it. It was all killer 
apps, wall to wall.”

“But?”

He stopped abruptly. “You must know this,” he said. “If you know 
about Clarity -- that's what I called it, Clarity -- then you know 
about what happened. With Buhle's resources, you can find out anything, 
right?”

She made a wry smile. “Oh, I know what history records. What I don't 
know is what \emph{happened}. The official version, the one that put 
Ate onto you and got us interested --”

“Why'd you try to kill Buhle?”

“Because I'm the only one he can't bullshit, and I saw where he was 
going with his little experiment. The competitive advantage to a firm 
that knows about such a radical shift in human cognition -- it's 
massive. Think of all the products that would vanish if numeracy came 
in a virus. Think of all the shifts in governance, in policy. Just 
imagine an \emph{airport} run by and for people who understand risk!”

“Sound pretty good to me,” Leon said.

“Oh sure,” she said. “Sure. A world of eager consumers who know 
the cost of everything and the value of nothing. Why did evolution 
endow us with such pathological innumeracy? What's the survival 
advantage in being led around by the nose by whichever witch-doctor can 
come up with the best scare-story?”

“He said that entrepreneurial things -- parenthood, businesses...”

“Any kind of risk-taking. Sports. No one swings for the stands when 
he knows that the odds are so much better on a bunt.”

“And Buhle \emph{wanted} this?”

She peered at him. “A world of people who understand risk are nearly 
as easy to lead around by the nose as a world of people who are 
incapable of understanding risk. The big difference is that the 
competition is at a massive disadvantage in the latter case, not being 
as highly evolved as the home team.”

He looked at her, really looked at her for the first time. Saw that she 
was the face of a monster, the voice of a god. The hand of a massive, 
unknowable machine that was vying to change the world, remake it to 
suit its needs. A machine that was \emph{good at it}.

“Clarity,” he said. “Clarity.” She looked perfectly attentive. 
“Do you think you'd have tried to kill Buhle if you'd been taking 
Clarity?”

She blinked in surprise. “I don't think I ever considered the 
question.”

He waited. He found he was holding his breath.

“I think I would have succeeded if I'd been taking Clarity,” she 
said.

“And if Buhle had been taking Clarity?”

“I think he would have let me.” She blurted it so quickly it 
sounded like a belch.

“Is anyone in charge of Buhle?”

“What do you mean?”

“I mean -- that vat-thing. Is it volitional? Does it steer this, this 
\emph{enterprise}? Or does the enterprise tick on under its own power, 
making its own decisions?”

She swallowed. “Technically, it's a benevolent dictatorship. He's 
sovereign, you know that.” She swallowed again. “Will you tell me 
what happened with Clarity?”

“Does he actually make decisions, though?”

“I don't think so,” she whispered. “Not really. It's more like, 
like --”

“A force of nature?”

“An emergent phenomenon.”

“Can he hear us?”

She nodded.

“Buhle,” he said, thinking of the thing in the vat. “Clarity made 
the people who took it very angry. They couldn't look at advertisements 
without wanting to smash something. Going into a shop made them nearly 
catatonic. Voting made them want to storm a government office with 
flaming torches. Every test subject went to prison within eight 
weeks.”

Ria smiled. She took his hands in hers -- warm, dry -- and squeezed 
them.

His phone rang. He took one hand out and answered it.

“Hello?”

“How much do you want for it?” Buhle's voice was ebullient. Mad, 
even.

“It's not for sale.”

“I'll buy Ate, put you in charge.”

“Don't want it.”

“I'll kill your parents.” The ebullient tone didn't change at all.

“You'll kill everyone if Clarity is widely used.”

“You don't believe that. Clarity lets you choose the course that will 
make you happiest. Mass suicide won't make humanity happiest.”

“You don't know that.”

“Wanna bet?”

“Why don't you kill yourself?”

“Because dead, I'll never make things better.”

Ria was watching intently. She squeezed the hand she held.

“Will you take it?”

There was a long pause.

Leon pressed on. “No deal unless you take it,” he said.

“You have some?”

“I can make some. I'll need to talk to some lab-techs and download 
some of my research first.”

“Will you take it with me?”

He didn't hesitate. “Never.”

“I'll take it,” Buhle said, and hung up.

Ria took his hand again. Leaned forward. Gave him a dry, firm kiss on 
the mouth. Leaned back.

“Thank you,” she said.

“Don't thank me,” he said. “I'm not doing you any favors.”

She stood up, pulling him to his feet.

“Welcome to the team,” she said. “Welcome to Buhle.”

\section{Afterword}

I wrote this story for a Frederic Pohl tribute anthology, inspired by 
\emph{The Space Merchants,} the classic novel he co-wrote with CM 
Kornbluth. I've been on a real neuroeconomics bender lately, fascinated 
with the idea that happiness, perceived happiness, and predicted 
happiness can be completely disjoint. It raises an interesting question 
in an era like ours, where governments increasingly see their role as 
managing people, rather than representing them. If you can make people 
happier by changing the way they feel about their circumstances, isn't 
that easier and more likely to succeed than changing the circumstances 
themselves?
\end{document}
