\input{common/hyp-en}

\begin{document}
%\setlength{\emergencystretch}{1ex}
\raggedbottom

\begin{center}
\textbf{\huge\textsf{Human Readable}}

\medskip
Cory Doctorow

\end{center}

\bigskip

\begin{flushleft}
This story is part of Cory Doctorow’s short story collection
“With a Little Help” published by himself. It is licensed under a
\href{http://creativecommons.org/licenses/by-nc-sa/}
{Creative Commons Attribution-NonCommercial-ShareAlike 3.0} license.

\bigskip

The whole volume is available at:
\texttt{http://craphound.com/walh/}

\medskip

The volume has been split into individual stories for the purpose of the
\href{http://ccbib.org}{Creative Commons Bibliothek.}
The introduction and similar accompanying texts are available under the 
title:
\end{flushleft}
\begin{center}
With a Little Help -- Extra Stuff
\end{center}

\newpage

\section{Human Readable}

\subsection{1. Nice networks don't go down}

It was unthinkable that the invisible ants that governed all human 
endeavor should catastrophically fail, but fail they did, 
catastrophically, on the occasion of Trish's eighth date with Rainer. 
It took nineteen seconds for the cascade of errors to slow every car on 
the Interstate to a halt, to light up the dashboard with a grim xmas 
tree of errors, to still the stereo and freeze the tickers of 
information and context that they had come to think of as the crawling 
embodiment of the colony that routed all the traffic that made up their 
universe.

“We are going to be: So. Late,” Rainer said, and Trish swiveled in 
her seat to look at him. He was Fretting again, his forehead wrinkled 
and his hands clenched on the steering wheel. When they traded massages 
(third date) and she'd rubbed at his hands, she'd found them tensed 
into claws that crackled with knuckle-fluid when she bent each finger 
back and rubbed sandalwood-scented oil into it. He was mighty cute for 
a neurotic -- at least he knew it when he was being nuts. Not that he'd 
stop being nuts, but he'd cheerfully admit it.

“We are not going to be late,” she said. “We just need to 
manually route ourselves out of the dead spot and get back on the grid 
and we'll be on our way. We've got plenty of time.”

“Dead spot?”

“Yes,” she said. His forehead wrinkles were looking more klingon by 
the second. “Dead spot.” She forced a chuckle. “You didn't think 
that the whole world was down, did you?”

He relaxed his knuckles. “Course not,” he said. “Dead spot. 
Probably ends up at the turn-off.”

“Right,” she said. “We need a map. I'm navigatrix. You're pilot. 
Tell me where your maps are, then get onto the shoulder and drive 
straight.”

“Where my maps are? Jesus, what century do you live in? My maps are 
with the sextant and sundial, between my leeches and my obsidian 
sacrifice-knife.”

She laughed. “OK, pal, I'll find a michelin, you drive. Every car has 
a couple maps. They self-assemble from happy meal boxes.” She opened 
the glove-compartment and started rooting through it while he pulled 
onto the shoulder and gunned the tiny two-seater along it.

“This is: So. Illegal,” he said.

“Naw,” she said. “I'm pretty sure you're allowed on the shoulder 
when the routing goes down. It's in the written-test manual. Learned it 
while I was helping my little cousin Leelee study. Aha!” she said, 
holding something up.

“You have a cousin named Leelee? That's uniquely horrible.”

“Shut up,” she said. “Look at this.” It was an old-fashioned 
phone, of a certain handsome retro line that made it look like a 
dolphin fucking a silver dildo, the kind of thing marketed to old 
people who wanted a device with its affordances constrained to collapse 
the universe of all possible uses for things that fit into your hand 
into the much smaller universe of, say, a cellphone.

“Yeah, my mom left that behind a couple years ago. I looked 
everywhere for it but couldn't find it. She must've been snooping in 
the glove-box. Serves her right. So what?”

“These things can unmesh and talk straight to a tower at a long 
distance, can't they?”

“I dunno, can they?”

“Oh yes, they can. Which means that they work in dead spots. So we 
can call and get directions.”

“You think you're pretty smart, huh, dumpling?”

She put her finger to her temple and made an adorable frowny thinky 
face, and held it until he looked at her and laughed. They'd discovered 
their ability to make one another laugh when he'd farted while taking 
off his kilt (second date) and had reflexively swung the hem back to 
make it appear that his mighty gust was ruffling the pleats.

“What's your mom's number?” she said.

He recited it and she tapped it in.

“Hi there! This is Trish, Rainer's friend? We're on the way, but the, 
well the, but the -- I mean to say, the grid's down or something. The 
car doesn't have any nav system, the dolby's out, the Interstate's a 
parking-lot\ldots{} Oh, you too? God. Wonder if it's the whole country! So, 
we need directions from San Luis Obispo, to the cemetery, if 
possible.”

:\ldots{}:

“Why yes, it's venti nice to be meeting \emph{you},” she said. 
“I've heard a lot about you, too. Yes, I'm giving directions, he's 
driving. Oh, that's so sweet of you. Yes, he \emph{does} look like he's 
going to scrunch his forehead into his upper lip. I think it's cute, 
too. Right. Got it. Left, then right, then left, then a slight left, 
then up the hill. Got it. Whups! That's the duracell! Better go. Soon! 
Yes. Whoops.”

“So?” he said.

“So, your mom sounds nice.”

“You got the directions?”

“She gave me directions.”

“So you know where we're going?”

“I don't have a single, solitary clue. Your mother gives 
\emph{terrible} directions, darling. Pull off at the next exit and 
we'll buy a map.”

“We are going to be: So. Late.”

“But now \emph{they know} we're late. We have an excuse. You: stop 
Fretting.”

\tb

Once they were on the secondary roads, the creepiness of the highway 
full of stopped cars and crane-necked drivers gave way to a wind-washed 
soughing silence of waves and beach and palms. Trish rolled down the 
window and let the breeze kiss the sweat off her lip, watching the 
surfers wiping out in the curl as the car sped toward the boneyard.

“Are you \emph{sure} this is the kind of thing you're supposed to 
bring a date to?”

“Yes,” he said. “Don't Fret. That's my job.”

“And you don't think it's even a \emph{little} weird to take a girl 
to a cemetery on a date?”

“We're not \emph{burying} anyone,” he said. “It's just an 
unveiling.”

“I still don't get that,” she said. “I keep picturing your mom 
cutting a ribbon with a giant pair of gold scissors.”

“Right, let's take it from the top,” he said. “And you'd better 
not be getting me to talk to stop my Fretting, because appealing to my 
pedantic nature to distract me is a \emph{very} cheap trick.”

“I'm fluttering my eyelashes innocently,” she said.

He laughed and stole a hand through the vent in her apron-trousers and 
over her thigh. “Achtung!” she said. “Eyes on road, hands on 
wheel, mind in gutter, \emph{this instant!}” She put her hand over 
his and he put down the pedal. His hand felt nice there -- too nice, 
for only eight dates and 20-some phone calls and about 100 emails. She 
patted it again.

“This is kind of fun,” he said, as they zipped past some surfer 
dudes staring glumly at their long-boards' displays, their perfect tits 
buoyant and colored like anodized aluminum with electric-tinted 
sun-paste.

“Ahem,” Trish said, squeezing his hand tight enough to make his 
knuckles grind together. “You were about to explain 
tombstone-unveiling to me,” she said. “When you got distracted by 
the athletic twinkies on the roadside. But I am sweet-natured and good 
and forgiving and so I will pretend not to have seen it and thus save 
us both the embarrassment of tearing out your Islets of Langerhans, all 
right?” She fluttered her eyes innocently in a way that she happened 
to know made him melt.

“Explaining! Yes! OK, remember, I'm not particularly Jewish. I mean, 
not that my parents are, either: they're just Orthodox. They don't 
believe in God or anything, they just like Biblical Law as a way of 
negotiating life. I renounced that when I dropped out of Yeshiva when I 
was 12, so I am not an authority on this subject.”

“Let the record show that the witness declared his utter 
ignorance,” she said. “But I don't get this atheist-Orthodox thing 
either --”

“Just think of them as Mennonites or something. They find the old 
ways to be a useful set of rules for navigating the universe's curves. 
God is irrelevant to the belief.”

“So they don't believe in God, but they pray to him?”

“Yeah,” he said. The surfers were all coming in now, jiggling their 
boards and rebooting them and staring ruefully at the radical cutback 
off the lip, dude, gnarly, as they plodded up the beach. “The ritual 
is the important part. Thinking good thoughts. Having right mind.

“It's good advice, most of it. It doesn't matter where it comes from 
or how it got there. What matters is that if you follow the Law, you 
get to where you're going, in good time, with little pain. You don't 
know why or how, but you do.”

“It's like following the ants,” she said, watching the stop-and-go 
traffic in the other direction. “Don't know why they tell us to go 
where they do, but they do, and it works.”

“Well, I guess,” he said, using the tone of voice that told her 
that he was avoiding telling her how wrong she was. She smiled.

“\emph{Anyway}. The thing about Jews -- ethnic Jews, cultural Jews, 
forget the religion here -- is that we're pretty much on the 
melodramatic end of the grieving scale. We like to weep and tear at our 
hair and throw ourselves on top of the coffins, right? So there's like 
5,000 years of this, and during that time, a bunch of social scientists 
-- Rabbinical scholars -- have developed a highly evolved protocol for 
ensuring that you grieve your dead enough that you don't feel haunted 
by guilt for having failed to honor them, but not grieving so much that 
you become a drag on the tribe.

“When someone dies, you bury him right away, usually within 24 hours. 
This means that you spend an entire day running around like your ass 
was on fire, calling everyone, getting the word out, booking 
last-minute travel, ordering in from the caterers, picking out a box, 
fielding consoling phone-calls, getting the rabbi on the phone, booking 
the limo, so much crap that you can't spare even a second to fall to 
pieces. And then you bury him, and while you're at it, your family 
extrudes a volunteer to go over to your house and take all the cushions 
off of one of the sofas, hang sheets over all the mirrors, and set out 
enough food to feed the entire state, along with an urn of starbucks 
the size of an oil-drum.

“Before the service starts, the rabbi gives you a razor-blade and you 
slash a hole in your lapel, so that you've got the rent in your heart 
hanging out there in plain sight, and once you get back home, you spend 
\emph{seven days} grieving. You pray two times a day with a quorum of 
ten men, facing east and singing the Kaddish, this really, really 
depressing song-prayer-dirge that's specially engineered to worm its 
way into the melancholy receptors of the Semitic hindbrain and make you 
feel really, really, really miserable. Other people come over and cook 
for you, all three meals. You don't see yourself in the mirrors, you 
don't sit on cushions, you don't do anything \emph{except mourn} for a 
whole week.

“Then it's over. You take a walk, leaving by one door and coming back 
in by the other. You put the mourning behind you and start your new 
life without your dear departed. You've given over your whole life for 
a whole week, done nothing but mourn, and you're completely sick of it 
by then, so you're almost glad to be done.

“Then, six months or a year later, usually just before Jewish New 
Year's, which is in the fall, you have a tombstone erected at the 
gravesite. The stone-cutters tie a white cloth around it, and everyone 
gathers there, and there's a sermon, and that dirge again, and more 
prayer, and everyone has a good hard cry as the scabs you've 
accumulated are ripped away and all your pain comes back fresh and 
scalding, and you feel it all again in one hot second, and realize with 
a guilty start that you \emph{have} been neglecting the memory of the 
loved one, which is to say that you've gotten on with your life even 
though his is over, which is to say that you've done perfectly healthy, 
normal stuff, but you feel totally, completely overwhelmed with guilt 
and love, which are kind of flipsides of the same emotion --”

“You don't believe that, do you?” She held her breath.

“Well, kind of. Not that they \emph{should be}, but hell they 
\emph{are}, most of the time, then.”

“Good thing we're not in love, then, right?” she said, in reference 
to their sixth date, when they'd decided that they would hold off on 
any declarations of love for at least an entire year, since they were 
most often moved to utter the Three Words of Significance when they 
were besotted with e.g. post-orgasmic brain-juice or a couple of 
cocktails.

“Yes, counselor.”

She shook her head. He \emph{knew} she was an academic, not a 
practicing lawyer, but he loved to tease her about it, ever since she'd 
revealed (after third date, on the phone) that she'd spent about ten 
seconds in private practice after she'd worked for her congressman and 
before she'd joined the faculty at UCLA.

“You're out of order,” she said.

“This whole damned car is out of order!” he said. “So that's the 
ritual. You \emph{said} you wanted to meet the parents and sisters and 
aunts and grandmothers and cousins and uncles and nephews and in-laws 
the next time we all got together. This is it.”

“Right,” she said. “I asked for this.” And she had, of course. 
Hadn't asked for the graveside elements, but she'd been curious to meet 
this big sprawling enterprise of a family that he was always nattering 
on about. This seemed as good an occasion as any. “So,” she said. 
“Is this a traditional date among Your People?”

He chuckled. “Yes, this is Yom Shiksa, the ritual bringing of the 
gentile woman to the family so that she may become the subject of 
intense, relentless scrutiny and speculation.”

She started to laugh, then saw that the tractors were stilled in the 
fields they were passing, that a train was stopped in its tracks, that 
the surfers were unable to get their roll-cage dune-buggies to take to 
the road.

“You all right, babe?” he said, after a couple minutes of this.

“Just wondering about the dead spot,” she said. “I wish we knew 
what had happened.”

“Nothing too bad, I'm sure,” he said. “It's all self-healing. I'm 
sure we'll be back online soon enough.”

\tb

They rolled into the parking lot for his family's \emph{shul}'s section 
of the giant graveyard a few minutes after 1PM, just over an hour late, 
along with the majority of the other attendees, all of whom had had to 
navigate manually.

“Where are your sisters?” Rainer's mother said, even before he'd 
kissed her cheek.

Rainer screwed his face up in a scowl and dug in his pocket for a 
yarmulke. “Do I know? Stuck in traffic, Ma. The grid's down 
everywhere.”

Trish watched this bemusedly, in her cool loose cotton apron-trousers 
and blouse. She scuffed her toe conspicuously and Rainer turned to her, 
and it was as though he'd forgotten she was there. She felt a second's 
irritation, then a wave of sympathy as she saw the spasm of anxiety 
cross his face. He was nervous about her meeting his fam, and nervous 
about who would arrive when, and nervous about where his sisters were 
with their enormous families and meek husbands, trapped somewhere on 
southern California's squillion-mile freeway network.

“Ma,” he said. “This is my friend, Trish.”

“Pleased to meet you, Mrs. Feinstein,” Trish said. The old woman 
was remarkably well-preserved, her soft skin glowing with heat-flush, 
her thick hair caught in a tight bun and covered with a little scarf 
that reminded Trish of Rainer's yarmulke. She wondered if she should be 
wearing one, too. Mrs. Feinstein's eyes flicked quickly to her shoes, 
up her legs and boobs, to her face and hair, and then back to her face. 
She opened her arms and drew Trish into a hug that smelled of good, 
subtle perfume, though Trish knew so little about scent that she 
couldn't have said which. “Call me Reba, darling,” she said. 
“It's so good of you to come.”

And then she was off, hustling to corral a wayward knot of 
horse\-play-aged cousins, stopping to shake hands with the deceased 
great-uncle's poker buddies in their old-man pants, golf shirts and 
knit yarmulkes bobby-pinned to their thinning hair.

Trish took stock. Looked like every other graveyard she'd been in, 
which wasn't that many. At 35, she'd been to half a dozen family 
funerals, a couple of college buddies who OD'ed or cracked up their 
cars, and one favorite poli-sci teacher's service, so she was hardly an 
expert on boneyards, but something was amiss.

“What's with the pebbles on the headstones?” she whispered to 
Rainer, who was scanning the road for signs of his sisters.

“Huh? Oh. You drop those on the monument when you visit the grave, as 
a sign that someone's been there.”

“Oh,” she said, and began to cast about for a pebble she could put 
on his great-uncle's headstone once it was unveiled. There were none to 
be found. The ground had been picked completely clean. Looking at the 
thousands and thousands of ranged marble headstones, each topped with a 
cairn of stones -- and not just stones, either, toys and seashells and 
small sculptures, she saw now -- and she understood why.

“What are you doing?” Rainer asked. He might have been irritated, 
or just nervous. It was hard to tell when he was Fretting, and he was 
clearly going coo-coo for coco-puffs.

“Looking for pebbles,” she said.

He said :fuck: very quietly. “I meant to bring some. Damnit. I've got 
twenty relatives buried here and we're going to go past every single 
tombstone before we get to leave and I don't have a single rock.”

“Can you leave toys or other stuff, like on those stones?”

“Yeah,” he said. “I suppose. If I had other stuff.”

She opened her purse and pulled out the dolphin-dildo cellphone. “You 
still need this?” she said.

He smiled and his forehead uncreased. “You're a genius,” he said.

She set it down on the pavement and brought her heel down on it hard, 
breaking it into dozens of fragments. “All the pebbles we'll ever 
need,” she said, picking them up and handing them to Rainer.

He put his arm around her shoulders and squeezed. “I'm awfully fond 
of you, Counselor,” he said, kissing her earlobe. His breath tickled 
her ear and made her think of the crazy animal new-relationship-energy 
sex they'd had the night before -- she was still limping, and so was he 
-- and she shivered.

“You too, steakypaste,” she said. “Now, introduce me to all of 
your relatives.”

“Introduce you?” He groaned. “You don't think I remember all of 
their names, do you?”

\tb

Afterward, they formed a long convoy back to the nearest family 
member's house -- a great aunt? a second cousin? Rainer was vague -- 
navigating by keeping everyone in sight, snaking along the traffic jam 
that appeared to have engulfed the entire state, if not the whole coast.

“You made that law, yes? We've all heard about you.” This was the 
sixth time someone had said this to her since they'd arrived and Rainer 
had made her a plate of blintzes, smoked salmon, fresh bagels, boiled 
eggs, and baby greens salad with raspberry dressing, then had been 
spirited away into an endless round of cheek-pinching and intense 
questioning. She'd been left on her own, and after having a couple of 
grave conversations with small children about the merits of different 
toys, she'd been latched upon by one of the Relatives and passed from 
hand to hand.

“I was involved in it, but I didn't write the law,” she said.

“Look at you, so modest, you're blushing!” the Relative said. She 
reached out to steady a cut-glass vase as it wobbled in the wake of two 
small boys playing keep-away with a third's yarmulke, and Trish 
realized that this was probably the hostess.

“This place is just supercalifragilistic,” she said, with an 
economical gesture at the tasteful Danish furnishings, the paper books 
in a handsome oak bookcase, the pretty garden out one side window and 
the ocean out the back window.

“Thank you,” the great-aunt said. “My Benny loved it here.” She 
misted up. Trish finally added two and two, remembered the BENJAMIN 
chiseled into the marble headstone, and the blank spot on the other 
half of the tombstone, realized that this wasn't just the hostess, this 
was the \emph{widow}, and felt about for a thing to say.

“It was a beautiful ceremony,” she said. She had a couple napkins 
tucked in the waistband of her pants, and without thinking, she 
extracted one and folded an angle into it, reaching for the corner of 
the great-aunt's eye. “Look up,” she said, and blotted the tear 
before it could draw a line of mascara down the widow's cheek.

The old woman smiled a well-preserved smile that reminded Trish of 
Rainer's mom. “You're a sweet girl,” she said. “Me, I'm not so 
good with names, and so I've forgotten yours.”

“I'm Trish,” Trish said, bemusedly. Rainer's grammar got 
yiddishized when he wasn't paying attention, and she adored the 
contrast between its shtetl credibility and his witty, smooth public 
ban\-ter-persona. It had attuned her to little phrases like, “Me, I'm 
not so good.”

The widow shook her hand. “I'm Dorothy. It's a pleasure to make your 
acquaintance. Would you like to come out to the garden with me?”

\tb

Once they were seated, young male Relatives materialized and set up 
shade-umbrellas and brought out trays of iced juice.

“They're not after the inheritance, you know,” the old woman said 
with a snort. “Their parents are \emph{very} well-off. They don't 
need from money. They just adore me because I've spoiled them rotten 
since they were babies and I'd take them swimming and to Disneyland.”

“You have a beautiful family,” Trish said.

“Do you have a big family, too?” The old woman put on a pair of 
enormous sunglasses and sipped at her pink grapefruit juice.

“Not like this one,” she said. There were a couple hundred people 
in the house, and Rainer had spent the whole car-ride back from the 
cemetery Fretting about all the relations who \emph{hadn't} made it.

“Oh, this one! Well, this is a special case. This family accumulates 
other families. My Benny had a small family, and when he married me, 
they just joined us. All the high holidays, we ate here, or at my 
parents' place, God rest them. Your family is in DC?”

“All over.”

“But you're from DC, no?”

“Not really. I grew up in Chicago and Seattle.”

“But you made that law --”

“I really didn't, honestly! I was clerking for a Supreme Court judge 
when the case was heard, and I wrote his dissenting opinion, and when 
we lost, I quit and went to work for a PAC that was agitating for 
copyright reform to accommodate free expression, and then when Senator 
Sandollar got voted in and they started the Intellectual Property 
committee and made her chairman, I joined her staff as a policy wonk. 
So I worked on it, along with a couple thousand other people, not 
counting the millions who contributed to the campaign and the people 
who knocked on doors and so on.”

“How old are you, darling?”

“35,” she said.

“At 35, I was having babies. You -- listen to you. Listen to what 
you've \emph{accomplished}! I'm proud just to \emph{listen} to you. 
Rainer is lucky to have you. You two will get married?”

Trish squirmed and felt her face grow hot. Neither of them really 
believed in marriage. Whenever anyone brought the subject up around 
Rainer, he'd grimace and say, “Are you kidding? It'd make my mother 
\emph{far} too happy -- she'd keel over from joy.”

There was some kind of disturbance down the beach, one that had been 
growing steadily over the past several minutes, and now the Relatives 
were all turning their attention that way, to a couple of small boys in 
miniature suits who were ruining the shine on their shoes running in 
the sand like lunatics.

Something in the way they were running, the distant expressions she 
couldn't quite make out on their faces. It made her think back to 
high-school, to working as a beach lifeguard on Lake Michigan in the 
summers, and before she knew what she was doing, she'd kicked off her 
shoes and was running for the shore, her legs flashing immodestly 
through the vents in her apron-trousers.

She was still yards away from the hissing surf when she began to assess 
the situation. There was the small boy, bobbing in the ocean, where the 
undertow had spit him up after sucking him under. There was the 
swimmer, unconscious on the beach, face down. Couldn't tell if his 
chest was moving, but the small boy was in a suit, not swim-trunks like 
the swimmer, and that meant that he was part of Rainer's Family, which 
she had begun (on the eighth date, no less!) to think of as her own, 
and so she had him as her primary target before she reached the sea.

She didn't bother finicking with the buttons on her top, just grabbed 
her collar and yanked, leaving her in a bra that revealed less than 
some bikinis she owned, but did so through a cunning arrangement of 
lace, mesh, and structural engineering that was probably illegal in 
Texas. She undid the bows on each hip holding up her pants and stepped 
out of them, leaving behind a very small pair of white panties whose 
primary design consideration had been to avoid showing lines through 
thin trousers, with modest coverage of all her nethers coming in a 
distant second.

She plunged into the water without hesitation, moving swiftly but 
surely, taking care to keep her feet dug in against the undertow as she 
waded out toward the young boy. She was a strong swimmer, but the water 
was shockingly cold after the heat of the garden and the buzzing 
afternoon and it sucked at her calves and legs like a jacuzzi intake. 
Her breath roared in her ears as she rode the swells, and then she was 
soaked by a succession of breakers, and then she had the boy's little 
hand.

She hauled him to her, seeing that he was only five or six, and that 
his pouting lips were alarmingly blue and that his skin was as pale as 
cream. She scooped the water out of his mouth, hooked her arm around 
his neck and tilted his head back and began to slosh back toward shore. 
When she was waist-deep -- immodestly revealed in a bra that she was 
quite certain had gone completely transparent -- she pinched his nose 
and blew into his mouth, not quite getting her mouth out of the way 
before he vomited up a gush of salt-water, blintzes, diet coke, and 
bile. She spat and wished that she could duck her head and get a 
mouthful of ocean to rinse with, but she couldn't without dunking the 
boy, too, so she hauled him up out of the water and handed him to the 
Relative who was standing with his arms on the shore, his fine leather 
shoes soaked with cold seawater.

She looked for the swimmer, and saw that he was still face-down in the 
sand. “You, you and you,” she said, pointing at three young cousins 
whose wide eyes were flicking from her boobs to her crotch -- white 
underwear, Christ, why white underwear today? -- to the boy on the 
sand, who was mobbed now with Relatives whose hubbub had reached 
deafening proportions, “Go to the house, find an old-fashioned phone 
and call emergency services. Tell them where we are, and that we have 
two drowning victims, one a child, neither breathing. What are you 
going to do?”

The tallest of the three managed to make eye-contact long enough to 
say, “Find a cell-phone, call emergency, tell them where we are, two 
drownings, not breathing.”

“Right,” she said. “Come back when you're done and tell me that 
it's done.”

“You, you,” she said, picking out two tall uncles who looked like 
they'd worked out or played sports before they found whatever careers 
had paid for the nice suits they were wearing, “Carry him here and 
lay him down on his side.”

She looked for Rainer and found his ass sticking out of the scrum 
around the boy. She snagged him by the belt and dragged him back. 
“Rainer,” she shouted. His forehead was scrunched, but he was 
clear-eyed and grim and looked like he was listening to her, which she 
found very pleasing. “You need to get everyone back at least five 
steps from that kid, and make them quiet down,” she said.

“Right,” he said, and took off his jacket and handed it to her.

“Hold it yourself,” she snapped, “I've got things to do.”

“It's to wear,” he said.

She surprised herself with a grin. “Thanks,”

The Relatives were murmuring, or crying, or bickering, but Rainer 
\emph{hollered}. “LISTEN UP,” he said. “All of you get over there 
by that rock, NOW, or my girlfriend won't be able to save Jory's life. 
GO!”

And they went, amazingly, crushing back so quickly they looked like a 
receding tide. The tall uncles deposited the swimmer in the sand 
between them, and she checked his breathing and saw that it was good.

“Turn him on his side and tell me if he starts to choke,” she said, 
and turned to the little boy, struggling to remember her rescue 
breathing.

\tb

She got the boy breathing and ended up with more puke on her face, on 
Rainer's jacket, in her hair. His pulse was thready but there. She 
turned to the swimmer and saw that he was a muscular surfer dude in 
board shorts with a couple of bitchun tatts and a decent body-paint 
job. He was breathing, too, but his heart was erratic as hell. She 
pressed two fingers to his throat.

“What happened?” she said. “Who saw it happen?”

One of the aunts stepped forward and said, “My son says they were 
playing --”

She held up her hand. “Where's your son?” she said.

“He's back at the house,” the aunt said, startling back.

“Send someone for him, then tell me what happened.”

The aunt looked like she'd been slapped, but the other Relatives were 
staring at her and so she had to talk, and then the boy arrived and he 
told it again and it was pretty much the same story, but she was able 
to get more details, as she began to examine both the boy and the 
surfer's bodies for cuts, bruises, breaks and punctures. She gave the 
boy's clothes the same treatment she'd given her own, gently but 
forcefully tearing them off, using a seashell to start the tears at 
first, then a pocket-knife that someone put in her hand.

The story was that the kids had been playing when they'd seen the 
surfer floating in the breakers, and they'd dared each other to fish 
him out, and the undertow had sucked them out to sea. One had gotten 
away, the other had ended out beyond the waves, and meanwhile, the 
surfer had beached himself on his own.

“Right,” she said. “Blankets and pillows. Elevate their feet and 
wrap them up good.” She stood up and staggered a step or two before 
Rainer caught her, and the crowd made a noise that was at once 
approving and scandalized.

“Get me to the sea,” she said. “I need to soak my head.”

So he walked her into the water, he still in his suit-pants and dress 
shirt and tie, and held onto her while she dunked her head and swirled 
a mouthful of salt water in her mouth.

“Where are the fucking paramedics?” she said, as she sloshed back 
out with him.

“There,” he said, and pointed at the horizon, where a Coast Guard 
clipper was zooming for the shore. “The cell-phone was dead, so I 
fired up a couple flares. You didn't hear them?”

“No,” she said. He could have set off a cannon and she wouldn't 
have noticed it.

She got back to the shore just in time to see the surfer convulse. She 
was on him in a second, kneeling at his side, doing 
airway-breathing-circulation checks, finding no pulse, and slamming him 
onto his back and beginning CPR.

Some time later, she was lifted off him and two paramedics went to work 
on him. Someone put a robe over her shoulders and a cup of juice in her 
hands. She dropped the juice in the sand and sticky liquid and beach 
sand covered her legs, which she realized now that she hadn't depilated 
in a week, and that made her realize that she'd spent a pretty crucial 
amount of time prancing around naked in front of her date's family, and 
that that was probably not on the timetable until the fifteenth date at 
\emph{least}.

She looked up at Rainer, who was still in his shoes and as she was in 
bare feet loomed over her. “God,” she said, “Rainer --”

He kissed her. “I love you, Patricia,” he said.

“Ooh,” she said, with a weak smile. “You're breaking the rules!”

“Can you let it go this once?”

She made her scrunchy thinky face and then nodded. “Just don't make a 
habit of it, you lunk.”

\tb

It would have been perfect if only the surfer hadn't died.

They didn't get home until well after midnight. Parts of LA appeared to 
be on fire as they inched their way along the freeway. It was weird to 
see LA at this speed. They were used to clipping along at 60 or 70 -- 
over 80 if the traffic was light -- flying over the freeway so fast 
that the scenery was just a blur. Only the year before, the \emph{New 
Yorker} had run a 40-page paean to LA, a public apology declaring it 
the most livable city in America, now that it had licked its traffic 
problems. It balanced lots of personal space with thorough urbanization 
and urbanity. It was why they both lived there.

Now they seemed to have traveled back 50 years in time, to the bad old 
traffic-jam-and-smog days. Looters danced below, torching stores, and 
the traffic moved so slowly that some people were apparently abandoning 
their cars to \emph{walk} home -- which made the traffic even worse. 
The smoke from the fires turned the sunset into a watercolor of reds 
and mustards and golds, tones that had blown away with the smog when 
the last gas-sucking Detroitmobile was retired for a plastic 
Nickel-Metal Hydride jellybean, and all the lanes were repainted to cut 
them in half.

It was nightmarish. When they got off the ramp at Studio City, they 
found homeless guys directing traffic with gas-tubes they'd torn out of 
the bus-shelters. The tubes glowed in the presence of microwave 
radio-frequency radiation, and as each of the trillions of invisible 
ants in the system attempted to connect with its neighbors and get the 
traffic set to rights again, the RF noise made the tubes glow like 
sodium lamps.

They coasted into Trish's driveway and collapsed in her living room.

“You were \emph{wonderful}, darling,” Rainer said, peeling off the 
tracksuit that one of his cousins had scrounged from the gym-bag in her 
trunk and donated to Trish. Her skin was gritted with sand and streaked 
with stripes of sunburn.

“God,” Trish said, lolling back on the sofa, just letting him 
gently brush away the sand and rub lotion into her skin. “You spoil 
me,” she said.

“You're unspoilable,” he said. “Wonderful girl. You saved their 
lives,” he said.

“What a fucking day,” she said. “You think that my lifeguard 
training made up for my scandalous undergarments in your family's 
minds?”

He snorted and she felt his breath tickle the fine hairs on her tummy. 
“You're kidding. My mom told me that if I didn't marry you, she'd 
have me killed and then fix you up with someone else from the family -- 
told me it was my duty to see to it that you didn't get away, just in 
case someone else fell in the ocean.”

He looked around at the blank walls. “Creepy not to have any news at 
all,” he said.

“There's a TV in the garage,” she said. “Or maybe the attic. You 
could find it and plug it in and find out that no one else knows what's 
going on, if you feel like it.”

“Or I could escort you to the bathtub and we could scrub each other 
clean and then I could give you a massage,” he said.

“Yes, or you could do that.”

“Where did you say the television was?” he said.

“You are going to be: In. So. Much. Trouble.” She twined her 
fingers in his hair and pulled him up to kiss her.

%\tb

\subsection{2. Progress pilgrims}

It took three days for even the thinnest crawls to return to the walls. 
In the meantime, people dug out old one-to-many devices like radios and 
televisions and set them up on their lawns so they could keep track of 
the aftermath of The Downtime.

He slept over those three nights, because no one was going anywhere, 
anyway, and they had a running argument over how many dates this 
counted as, but truth be told, they had a wonderful time, making 
omelets for one-another, washing each other's backs in the shower, 
stealing moments of sex in the living room at two in the afternoon 
without worrying about being interrupted by a chime, ringer, bell or 
vibe.

When they weren't enjoying each other, they took coolers of fizzy 
drinks onto the lawn and watched the neighbor's TV set and saw the 
pundits describing The Downtime. The news-shows were having a drunken 
ball with this one: as the only game in town, they were free to bring a 
level of craft to their newsmongering that hadn't been seen since 
Trish's parents' day, when news-networks turned catastrophes into light 
operas, complete with soundtracks, brand-identities, logo-marks and 
intermissions where buffoons worked the audience for laughs.

“Oh, she's your favorite, isn't she?” Trish asked, goosing Rainer's 
bicep and taking a sip of his peach ginger-ade. The pundit had been in 
heavy rotation since the TV went back on the air. She was a Norwegian 
academic mathematician who wrote books of popular philosophy. She was a 
collection of trademark affectations: a jacket with built-up shoulders, 
a monocle, a string tie, nipple tassles, and tattooed cross-hatching on 
her face that made her look like a woodcut of a Victorian 
counting-house clerk. Rainer loathed her -- she'd been on the committee 
to which he'd defended his Philosophy of Networks thesis, and she'd 
busted his balls so hard that they still ached a decade later when he 
saw her on the tube.

The pundit explained the packet-switching, using trains versus 
automobiles as a metaphor: “In a circuit uniwerse, every 
communication gets its own dedicated line, like a train on a track. Ven 
I vant to talk to you, ve build a circuit -- a train track -- betveen 
our dewices. No one else can use those tracks, even if ve're not 
talking. But packet-svitching is like a freevay. Ve break the 
information up into packets and ve give every packet its own little 
car, and it finds its own vay to the other end. If vun car doesn't 
arrive, ve make a copy of its information and send it again. The cars 
have brakes and steering veels, and so they can all share the same road 
vithout too much trouble.”

Rainer grit his teeth and hissed at the set. “She's faking the 
accent,” he said. “She thinks that Americans believe that anyone 
with a European accent is smarter than we are. She can pronounce vee 
and doubleyou perfectly well when she wants to -- she speaks better 
English than I do! Besides, \emph{she stole that line from me},” he 
said, “from my \emph{thesis},” he said, his face scrunching up 
again.

“Shh, shh,” Trish said, laughing at him. He wasn't really 
angry-angry, she knew. Just a little stir crazy. He was a networking 
guy -- he should have been out there trying to make the network go 
again, but he was on sabbatical and no one at UCLA wanted to hear from 
him just then.

And then the pundit was off onto ants -- networks modeled on 
ant-colonies that use virtual pheromones to explore all possible routes 
in realtime and emerge a solution to the problem of getting everything, 
everywhere, in shortest time. Rainer kept barking at the TV, and Trish 
knew he was doing it to entertain her as much as for any reason, so she 
laughed more and egged him on.

The TV cut back to the news-dude, who was a very cuddly ewok who'd made 
his name hosting a wheel-of-fortune, jumping up and down and squeaking 
excitedly and adorably whenever a contestant won the grand prize, his 
fur-plugs quivering. He cupped his paws to his cheeks and grinned.

“But ants aren't perfect, are they?” the ewok said.

“He's feeding her!” Rainer said. “She's going to go off on her 
stupid walking-in-circles bit --”

“The thing about using wirtual ants to map out the vorld and make 
routing recommendations is that ve can't really tell the difference 
betveen a good solution and a bad vun, without trying it. Sometimes, 
ants end up valking in circles, reinforcing their scent, until they 
starve to death. Ve might find that our cars tell us that the best vay 
from San Francisco to San Jose is via a 1500 mile detour to Las Wegas. 
It may be true -- if all the traffic eweryvhere else is bad enough, 
that might be the fastest vay, but it may just be the ants going in 
circles.”

“God, talk about taking a metaphor too far,” he said. Trish thought 
that Rainer was perfectly happy to think about the ants as ants, except 
when someone raised a point like this, but she didn't see any reason to 
raise that point just then.

The ewok turned to the camera: “One scientist says we \emph{should} 
expect more Downtimes to come. When we come back from this break, we'll 
talk to a University of Waterloo researcher who claims that this is 
just the first of many more Downtimes to come.”

The screen cut over to a beautiful, operatic advertisement for some 
Brazilian brand of coca-cola, wittily written, brilliantly shot, with 
an original score by a woman who'd won three grammies at the 
Independent Music Awards in Kamchatka the year before. They watched it 
with mild attention, and Trish absently fished another bottle out of 
the cooler and chewed the lid off with her side-molars.

She looked at Rainer. He was gripping the arm-rests of his inflatable 
chair tightly, dimpling the hard plastic. She held the bottle to his 
lips and he took it, then she rubbed at his shoulders while he took a 
swallow.

“Let's go back inside and play,” she said. “They won't have 
anything new to tell us for days.”

\tb

The crawls were alive the next morning, exuberantly tracking across the 
walls and over the mirror and down the stairs. They picked out the 
important ones and trailed them to convenient spots with a fingertip 
and devoured them, reading interesting bits aloud to one another.

Soon the crawls had been tamed and only a few personal messages 
remained. Trish dragged hers over to the tabletop, next to her cereal 
bowl, and opened them up while she ate. Outside, she could hear the 
whisper of cars speeding down the road, and she supposed with a 
mingling of regret and relief that she should probably go into her 
office.

She opened her personal mail. It had been three days since she'd read 
it, but for all that, a surprisingly small amount had accumulated. Of 
course -- everyone else had been without connectivity, too. This was 
mostly stuff from the east coast and Europe, people who'd been awake 
for a couple hours.

She read, filed and forwarded, tapping out the occasional one-word 
answer to simple questions or bouncing back messages with a form letter.

Then she came to the note from the Coast Guard medic. He didn't mince 
words. It was in the first sentence: the surfer dude she'd rescued had 
had a second cardiac arrest on the boat. They'd tried what they could, 
but he hadn't recovered. He was a freak statistic of The Downtime, 
another person who'd lost his life when the ants spazzed out. They'd 
recovered his board and found its black-box. The accelerometer and GPS 
recorded the spill he'd taken after the loss of climate and 
wave-condition data from the other surfers strung out on the coast. 
He'd stayed up for about ten seconds before going under.

She stared numbly at the note, the spoon halfway to her mouth, and then 
she dropped the spoon into the bowl, not noticing that it splashed milk 
down her blouse.

She got up from the table and went into the kitchen. Rainer was there, 
in a change of clothes they'd bought from a mom-n-pop gap at the mall 
on the corner that had been taking IOUs from anyone who could show a 
driver's license with a local address. She grabbed his wrist, making 
him slosh starbucks down his front, she took the cup out of his hand 
and set it down on the counter, then put her arms around his chest and 
hugged him. He didn't protest or ask any questions, he just put his 
arms around her and hugged back.

Eventually, she cried. Then she told him what she was crying about. She 
let him tell her that she was a hero, that she'd saved Jory's life and 
almost saved the surfer's life, and she let him tell her that it wasn't 
her fault for sloshing into the ocean to rinse off the barf, and she 
let him tell her that he loved her, and she cried until she thought she 
was cried out, and then she started again.

He took her upstairs and he laid her down on the bed. He undressed her, 
and she let him. He put her in fluffy jammies, and she let him. He 
wiped away her makeup and her hot tears with a cool face-cloth, and she 
let him. He took her hand and ran his fingers over her fingernails, 
squeezing each one a little, the way she liked, and she let him.

“You're going to have a nice lie-down for a couple hours, and I'm 
going to be right beside you. I'll call the department secretary and 
tell him you're taking a personal day and will be in tomorrow. Then 
we're going to go see Jory and his family, so that you can see the boy 
whose life you saved, and then we are going to go for a walk in the 
hills, and then I'm going to put you to bed. When you get up in the 
morning, you can make an appointment to see a grief counsellor or not. 
Today, I'm in charge, all right?”

Her heart swelled with love and she felt a tear slip down her cheek. 
“Rainer,” she said, “you're a wonder.”

“You inspire me, darling,” he said, and kissed her eyelids shut.

\tb

Their thirty-fifth date was their last.

“You're going back to Washington,” he said, when he saw the boxes 
in her office.

“Yes,” she said.

He stood in the doorway of her office. Trish was painfully aware of the 
other faculty members in the corridor watching him. Their romance was 
no secret, of course. Everyone in the law department knew about him, 
all the network engineers knew about her, and they both took a 
substantial amount of ribbing about “mixed marriages” and 
“interfaith dating.”

Trish realized with a pang that it was likely that everyone in the law 
department knew that she'd decided to go back to the Hill but that he'd 
only suspected it until this instant.

“Well, good for you,” he said, putting on a brave face that was 
belied by the Fret wrinkles in his forehead.

“I'm sorry,” she said. “I should have told you once I decided, 
but I didn't want to do it over the phone --”

“I'm glad you didn't,” he said, holding up his hand. “Do you want 
to come out for dinner with me anyway?”

She gestured at the half-packed office. “The movers are coming in the 
morning.”

“Well then, do you suppose you could use some help? I could get some 
burger king or taco bell.”

She looked at him for a long moment, swallowing the knob in her throat. 
“That would be lovely. Mexican. I mean, `taco bell,'” she said. 
“Thank you.”

He let her pay for it -- “You're making the big bucks now,” he said 
-- and he was a surprisingly conscientious packer, padding her framed 
pictures carefully and wrapping her knick-knacks in individual sheets 
of spun fiber.

“Well then,” he said, once he'd finished writing out a description 
of his latest box's contents on its outside, “you always told me that 
Hill Rats were Hill Rats for life, I suppose.”

“Yeah,” she said. She knew she should explain, but they'd had the 
argument about it three times since the new PAC had contacted her and 
offered her the executive director position. The explanation wouldn't 
get any better now that she'd made up her mind.

The new PAC, The Association for a Human-Readable World, was the 
brainchild of some people she'd worked with while she was on the Hill. 
They'd asked her to hire a team, to scout an office, and then to camp 
out in the offices of various important committee chairmen until they 
passed a law limiting the scope of emergent networking meshes. The 
Europeans had enacted legislation requiring cops, hydroelectric 
agencies, banks, hospitals and aviation authorities to use 
“interrogatable” networks within ten days of The Downtime. With 
fifteen thousand dead in Western Europe alone, with Florence in flames 
and Amsterdam under two meters of water, it was an easy call. The US 
had scoffed at them and pointed to the economic efficiencies of a 
self-governing network, but the people who were funding Human-Readable 
World wanted to know where old concepts like “transparency” and 
“accountability” and “consent of the governed” fit in when the 
world's essential infrastructure was being managed by nonsentient 
ant-colony simulations.

“Be gentle with us, OK?” he said.

“Oh, I wish I had your confidence in my abilities,” she said, 
sucking on her big-gulp of coke.

He put down his food and looked hard at her. He stared longer than was 
polite, even for (ex-) lovers, and she began to squirm.

“What?” she said.

“You're not putting me on. Amazing. Patricia Lourdes McCavity, you 
have felled an empire and you are setting yourself up to fell another 
-- and it's one that I'm pretty heavily invested in, both 
professionally and financially.”

`Come on,“ she said. ”I'm good, but I'm not superwoman. I was part 
of a team.”

“I've read your briefs. Position papers. Opinions. Speeches. Hell, 
your press-releases. They were the most cogent, convincing explanations 
for intellectual property reform I'd ever read. You weren't the judge, 
but you were his clerk. You weren't the committee chairman, but you 
were her head staffer. Taco Bell underestimated you. Coke 
underestimated you. Starbucks underestimated you. Disney underestimated 
you. Vivendi and Sony underestimated you. Now you're running your own 
organization, and it's pointed at me, and I'm scared shitless, you want 
to know the truth. I'm not underestimating you.” He'd drawn his dark 
eyebrows together while he spoke, and lowered his head, so that he was 
looking up at her from under his brow, looking intense as the day 
they'd met, when he was delivering a brilliant lecture on ant-colony 
optimization to a large lay audience at the law-school, fielding the 
Q{\&}A with such convulsive humor and scalding lucidity that he'd 
melted her heart.

She felt herself blushing, then wondered if she was flushing. She still 
loved him and still craved the feeling of his skin on hers, wanted 
nothing more than another lost weekend with him, taking turns being the 
strong one and being the one who surrendered, soothing each other and 
spoiling each other. Thinking of that first meeting brought back all 
those feelings with keen intensity that made her breasts ache and her 
hands flutter on the box she was eating off of.

“Rainer,” she began, then stopped. She took a couple deep breaths. 
“I'm not gunning for you, you know. You and I want the same thing: a 
world that we can be proud to live in. Your family's company has 
contributed more to the public good than any of us can really 
appreciate --”

He blushed now, too. She never talked about his father's role in the 
earliest build-outs of ant-based emergent routing algorithms, about the 
family fortune that he'd amassed through the company that bore his name 
still, 30 years after he'd stepped down as Chairman of the Board. 
Rainer was a genius in his own right, she knew, and his own 
contributions to the field were as important as his father's, but he 
was haunted by the idea that his esteem in the field was due more to 
his surname than his research. He waved his hands at her and she waved 
hers back.

“Shush. I'm trying to explain something to you. Between your father 
and you, the world has increased its capacity and improved its quality 
of life by an order of magnitude. You've beaten back Malthus for at 
least another century. That makes you heroes.

“But your field has been co-opted by corrupt interests. When you 
study the distributions, you can see it clearly: the rich and the 
powerful get to their destinations more quickly; the poor are routed 
through franchise ghettoes and onto toll-roads; the more important you 
are, the fewer number of connections you have to make when you fly, the 
better the chance that you'll get a kidney when you need it. The 
evidence is there for anyone to see, if only you look. We need 
standards for this -- we need to be able to interrogate the system and 
find out why it does what it does. That's an achievable goal, and a 
modest one: we're just asking for the same checks and balances that we 
rely on in the real world.”

He looked away and set down his taco. “Trish, I have a lot of respect 
for you. Please remember that when I tell you this. You are talking 
nonsense. The network is, by definition, above corruption. You simply 
can't direct it to give your cronies a better deal than the rest of the 
world. The system is too complex to game. Its behavior can't be 
\emph{predicted} -- how could it possibly be \emph{guided}? Statistics 
can be manipulated to `prove' anything, but everyone who has any clue 
about this understands that this is just paranoid raving --”

She narrowed her eyes and sucked in a breath, and he clamped his lips 
shut, breathed heavily through his nose, and went on.

“Sorry. It's just wrong, is all. Science isn't like law. You deal 
with shades of grey all the time, make compromises, seek out balance. 
I'm talking about mathematical truths here, not human-created political 
constructs. There's no one to compromise \emph{with} -- a 
human-readable emergent network just doesn't exist. Can't exist. It 
doesn't make sense to say it. It's like asking for me to make Pi equal 
three. Pi \emph{means} something, and what it means \emph{isn't} three. 
Emergent networks \emph{mean} not-human-readable.”

She looked at him, and he looked at her, and they looked at each other. 
She felt a sad smile in the corners of her lips, and saw one tug at 
his, and then they both broke out in grins.

“We're going to be seeing a lot of each other,” she said.

“Oh yes, we are,” he said.

“Across a committee room.”

“A podium.”

“On talk-shows.”

“Opposite sides.”

“Right.”

“No fighting dirty, OK?” he said, raising his eyebrows and showing 
her his big brown eyes. She snorted.

“Give me a hug and go home,” she said. “I'll see you at the 
hearings when they introduce my bill.”

He hugged her, and she smelled him, thinking, \emph{this is the last 
time I'll smell this smell.}

“Rainer,” she said, holding him at arm's length.

“Yes?” he said.

“I'm going to call you, when I have questions about ant-colony 
optimization, all right?”

He looked at her.

“I need the best expertise I can get. It's in your interest to see to 
it that I'm well-informed.”

Slowly, he nodded. “Yes, you're right. I'd like that. I'll call you 
when I have questions about policy, all right?”

“You're on,” she said, and they hugged again, fiercely.

Once he was gone, she permitted herself the briefest of tears. She knew 
that she was right and that she was going to make a fool out of him, 
but she didn't want to think of that right then. She felt the place 
behind her ear where he'd kissed her before going home and looked 
around her office, five years of her life in thirty banker's boxes 
ready to be shipped across the country tomorrow, according to a route 
that would be governed from moment to moment by invisible, notional, 
\emph{ridiculous} insects.

She ate more taco bell. The logo was a pretty one, really, and now that 
it had been adopted by every mom-and-pop burrito joint in the world, 
they'd really levelled the playing field. She thought about the old 
Taco Bell mystery-meat and plastic cheese and took a bite of the ground 
beef and sharp Monterey Jack that had come from her favorite little 
place on the corner, and permitted herself to believe, for a second, 
anyway, that she'd made that possible.

She was going to kick ant ass on the Hill.

%\tb

\subsection{3. Conflict of Insect}

Trish gathered her staff in the board room and wrote the following in 
glowing letters on the wall with her fingertip, leaving the text in her 
expressive schoolmarm's handwriting rather than converting it to some 
sterile font: “First they ignore you. Then they laugh at you. Then 
they fight you. Then you win.”

Her staff, all five of them, chuckled softly. “Recognize it?” she 
asked, looking round at them.

“Pee-Wee Herman?” said the grassroots guy, who was so young it 
ached to look at him, but who could fire a cannonload of email into any 
congressional office on 12 hours' notice. He never stopped joking.

The lawyer cocked an eyebrow at him and stroked her moustache, a 
distinctive gesture that you could see in any number of courtv archives 
of famous civil-rights battles, typically just before she unloaded both 
barrels at the jury-box and set one or another of her many precedents. 
“It's Martin Luther King, right?”

“Close,” Trish said.

“Geronimo,” guessed the paralegal, who probably wasn't going to 
work out after all, being something of a giant flake who spent more 
time on the phone to her girlfriend than filing papers and looking up 
precedents.

“Nope,” Trish said, looking at the other two staffers -- the office 
manager and the media guy -- who shrugged and shook their heads. 
“It's Gandhi,” she said.

They all went, “Ohhhh,” except the grassroots guy, who crossed to 
the wall and used his fingertip to add, “And then they assassinate 
you.”

“I'm too tough to die,” the lawyer said. “And you're all too 
young. So I think we're safe.”

“OK,” Trish said. “This is an official pep talk. They're playing 
dirty now. Last night, my car tried to take me to Arlington via 
Detroit. My email is arriving on a 72 hour time-delay. My phone doesn't 
ring, or it rings all night long. I've had to switch it off.

“But what all of this means is that I've got more uninterrupted 
work-time than ever and I'm getting reacquainted with my bicycle.”

“Every number I call rings at my ex-girlfriend's place,” the 
grassroots guy said. “I think we're going to get back together!”

“That's the right attitude, boy-o,” the lawyer said. “When life 
gives you SARS, make sarsaparilla. I appear to be unable to access any 
of my personal files, and any case-law I query shows up one sentence at 
a time. I've discovered that the Georgetown University law-library 
makes a very nice latte and serves a terrific high tea, and I've set 
Giselle to work on refiling and cross-indexing twenty years' worth of 
yellow pads that had previously sat mouldering in a storage locker that 
I was paying far too much for.”

“Which has given Giselle a rare opportunity to explore the rich civil 
rights history that you embody,” Trish said, looking pointedly at the 
paralegal. “But I suspect that she could use a hand, possibly from a 
grad student or two who could get some credit for this. Let's ask 
around at Georgetown, OK?”

The lawyer nodded. The office manager pointed out that their 
bill-payments were going astray after they'd been dispatched to their 
suppliers but before they were debited from their -- dwindling -- 
account, which meant that they were getting a couple days' worth of 
free cash-flow. Only the media guy was glum, since he couldn't field, 
make or review calls or press-releases, which made him pretty useless 
indeed.

“Right,” she said, and scribbled something on one of the steno pads 
she'd bought for everyone when their email started going down three 
times a day. “This guy owes me from back in the copyright wars -- I 
fed him some good stories that he used to launch his career. He was the 
ABNBC Washington bureau chief until last year and now he's teaching 
J-School at Columbia. Take the afternoon train to Manhattan and bring 
him back with you tonight. Don't take no for an answer. Tell him to 
bring his three most promising proteges, and tell him that they'll have 
all the access they need to produce an entire series on the campaign. 
Sleeping on our sofas. Following us to the toilet. Everything on the 
record. Do-able?”

“It's do-able,” the media guy said. “I'm on it.”

Once they'd all cleared out, the lawyer knocked on her door. “You 
going to be all right?” she asked.

Trish waved her hands at the piles of briefing books, red-lined 
hardcopies, marked-up magazine articles and memos from her Board of 
Directors. “Of course!” she said. She shook her head. “Probably. 
We never thought we'd get this far, remember? All this psy-ops shit 
they're pulling, it's just more proof that we're on the right track. No 
one should be able to do this. It's the opposite of democracy. It's the 
opposite of civil discourse.”

The lawyer smoothed her moustache. “Right on,” she said. “You 
should be proud. This is a hell of a fight, and I'm glad to be part of 
it. You know we'd follow you into the sun, right?”

Trish fluttered her hands. “God, don't give me that kind of 
responsibility.”

“All right then, into the ocean. We're making this happen, is what's 
important.”

“Thanks, babe,” Trish said. She put on a brave smile until the 
lawyer had backed out of the office, then stared down at her calendar 
and looked at her morning schedule. Three congressional staffers, a 
committee co-chair, an ACLU researcher, and the head of the newly 
formed Emergent Network Suppliers' Industry Association -- a man she 
had last seen in her office at UCLA, backing away from a long and 
melancholy hug.

\tb

When he rang off the phone and joined her, finally, she straightened 
out her smart cardigan and said, “Rainer, you're certainly looking\ldots{} 
well.”

“\ldots{}funded,” he finished, with a small smile. The Emergent Network 
Suppliers' Industry Association's new offices were in a nice Federal 
Revival building off Dupont Circle, with lots of stained glass that 
nicely set off the sculptural and understated furniture. “It's not as 
grand as appearances suggest, Trish. We got it for a song from the 
receivers in the Church of Scientology's bankruptcy, furnishings 
included. It \emph{is} nice though. Don't you think?”

“It's lovely,” she said. Around her, staffers bustled past in good 
suits and good shoes and smart haircuts. “Hard to believe you only 
set up shop a week ago,” she said.

“It came furnished, remember,” he said.

“Oh yes, so you said,” she said, watching a kid who looked like 
he'd gone tops in his class at the Naval Academy put his ankles up on 
the plasticized return beside his desk and tilt his chair, throwing his 
head back with wild laughter at whatever it was some other Hill Rat (in 
her mind, it was a key Congressman's aide -- some old frat buddy of Mr 
Navy 2048) was saying at the phone's other end.

She looked back \erratum{and}{at} Rainer and saw that he was staring where she had.

“Well, it's a far cry from academic research,” he said.

“I know you'll be very good at it. You can explain things without 
making it seem like an explanation. The first lesson I ever learned on 
the Hill was, `If you're explaining --'”

“'-- you're losing,'” he said. “Yeah, I've heard that. Well, 
you're the old hand here, I'm just learning as I go. Trying not to make 
too many mistakes and to learn from the ones I do make.”

“Do you want some free advice, Rainer?”

He sat down in one of the chairs, which bulged and sloshed as it 
conformed itself to his back and butt. He patted the upholstered jelly 
beside him. “You may always assume that I would be immensely grateful 
for your advice, Trish,” he said.

She sat down and crossed her legs, letting her sensible shoe hang 
loose. “Right. DC is a \emph{busy} place. In academic circles, in 
tech circles, you might get together to feel out your opponent, or to 
make someone's acquaintance, or to see an old friend. You might get 
together to enjoy the company of another human being.

“We do that in DC, \emph{after} working hours. Strictly evenings and 
weekends. When you schedule a meeting during office hours, it has to 
have a purpose. Even if it appears to have no purpose, it has a 
purpose. There's a protocol to meetings, a secret language, that's 
known to every Hill Rat and written nowhere. What time you have the 
meeting, who's there, who's invited, who knows it, how long you 
schedule, whether you cater: they all say little things about the 
purpose of the meeting. Even if you have no reason to call the meeting, 
one will be read into it.

“If this was any other city in the world, it would make perfect sense 
for you to look me up once you got to DC. We're still friends, I still 
think about you from time to time, but here in DC, you calling me over 
for a meeting, this kind of meeting, at this time of day, it means 
you're looking to parley. You want to strike a deal before my bill goes 
to the committee. I don't know how well you know the Hill, so I don't 
want to impute any motives to you. But if you took a meeting like this 
with anyone else, that's what they'd assume.”

Rainer's forehead crinkled.

“No Fretting,” she said. Then she smiled a sad smile. “Oh, Fret 
if you want. You're a big boy.”

He twiddled his thumbs, caught himself at it, and folded his hands in 
his lap. “Huh,” he said. “Well. I \emph{did} want to talk to you 
because it's been a while and because we meant a lot to each other. I 
\emph{also} wanted to talk to you about the bill, because that's what 
I'm here to do, at a pretty decent salary. I \emph{also} wanted to see 
you because I had an idea that you'd be different here in your native 
habitat, and well, that's true.”

She refused to let that make her self-conscious. Of \emph{course} she 
was different, but it wasn't geographic. The last time they'd seen each 
other, they were lovers and friends. Now they were ex-lovers who were 
being paid to accomplish opposing, mutually exclusive objectives. She 
knew that there was a certain power in not saying anything, so she 
wrapped herself in silence and waited for him to say something. She 
didn't have to wait long.

“Your bill is going to committee?”

“Well, I certainly hope so,” she said. “That's what I'm here for, 
after all. The discussion draft has been circulating for a week, and 
we're confident we'll see it introduced and assigned to committee by 
the end of this week. That's what we're told, anyway. It's got strong 
bipartisan support. Selling Congress on the importance of 
human-generated governance is pretty easy. Wouldn't want to be in your 
shoes.”

He grinned. “You're trying to psych me out.”

“Maybe,” she said, grinning back. “But that's nothing compared to 
the psych job that we've been getting down at my office.” She told 
him about the phone weirdness, the oddball traffic-management. 
“Someone on your side has a funny sense of humor.”

His smile faded. “You're still trying to fake me out,” he said. 
“If you're seeing corruption in the net, it's because you're looking 
so hard, you can't help but find it. You're reading malice into 
accident. Dead spots aren't personal, you know. This is a law of nature 
-- the networks emerge solutions, they're the best they can come up 
with. If you don't like the results, talk to nature, not me.”

She shrugged. “Whatever, Rainer. I know what's happening. You'll 
believe what you want to believe.” She pursed her lips and made an 
effort at controlling her irritation. “It's really happening, and 
it's not helping your side. If you know who's responsible, you might 
let him know that the dirty tricks are what convinced Senator 
Beauchamp's staff to green-light the bill. Conspiracy is supposed to be 
beneath the surface. It doesn't look so good when it's exposed to fresh 
air and sunshine.”

“You've got to be kidding,” he said. “Doesn't matter, I suppose. 
All right, message received. If I happen to run into someone whom I 
think should hear it, I'll be sure to pass it on, OK?”

“That's all I ask,” she said.

“You want to talk about the bill now?”

“Have you seen the discussion draft?”

He squirmed. “No,” he admitted. “I didn't know it existed until 
just now.”

“I'd offer to send you a copy, but I expect it would take a week to 
arrive, if it ever did. Why don't you ask that guy,” she gestured at 
the Navy man, “to get you a copy? He looks like he knows his way 
around. And then drop by my office if you want to chat about it.”

She stood up and tugged at her cardigan again. “It's been very nice 
seeing you,” she said. She picked up her coat and her mitts. “Good 
luck settling in.”

He gave her a hug -- which felt weird, hugging was strictly 
west-of-the-Mississippi, and she broke it off firmly -- and showed her 
to the door. The first snows were coming in, and the steps were 
slightly icy, so she maneuvered them slowly, carefully. When she 
reached the road, he was no longer in the doorway. He was standing 
right behind her, breath coming out in foggy huffs.

“Trish,” he said, then stopped. His arms dropped to his sides and 
his shoulders slumped.

“Rainer,” she said, keeping her voice calm and neutral.

“God,” he said. “God. How'd this happen, Trish? Look, I've never 
been happy the way I was with you. I haven't been that happy since. 
God, Trish --”

“Rainer,” she said again, taking one of his hands, firmly, 
motherly. “Rainer. Stop it. You're here to do a job, and your job 
requires that you and I keep it on a professional level. It doesn't 
matter how it happened --” But it did, didn't it? She'd left him to 
come east and do something he thought of as wrong-headed and backwards 
and superstitious. But she'd left him, not the other way around. And 
he'd never recovered, though she'd built herself a new life here. It 
wasn't a contest (but she was winning anyway). “It doesn't matter. We 
respect each other. That's enough.”

He deflated and she said, “Oh, come here,” and gave him a long and 
soulful hug, right there on the street, knowing that she was giving the 
hug and he was taking it. Then she let him go, spun him round, and gave 
him a little push back toward his office.

By the time she reached the corner and looked back over her shoulder, 
he was nowhere to be seen.

\tb

That afternoon, her phone started ringing normally, with actual people 
on the other end. Her outbound calls were connected. Her email was 
delivered. Her car got her home in record time. She sighed as she eased 
it into her driveway and carried her briefcase inside and poured 
herself a very small glass of Irish whisky so rare that it had been 
known to make grown men weep. Normally, she saved it for celebrations, 
but if she was celebrating something, she was damned if she knew what 
it was.

Her phone rang as she was licking the last few drops of liquor from the 
little glass. It was the lawyer, with news.

“I just stopped by the office and found a messenger on the doorstep. 
He had hard-copy of a press-release from Senator Beauchamp's office. 
They're introducing the bill in the morning. Congrats, kid, you did 
it.”

Trish set the glass down and said :whoopee: very quietly and very 
emphatically.

“You're durned tootin',” the lawyer said. “And double for me.”

Of course, it wasn't over by a long shot. Getting a bill introduced was 
not the same as getting it through committee. Getting it through 
committee was not the same as getting it passed in the Senate, and 
getting it passed in the Senate was not the same as getting it passed 
in the House, and then who the hell knew what the hereditary 
Chimp-in-Chief in the Oval Office would do when it was passed through 
the bars of his cage with his morning banana.

But she had ridden back into town less than a year before, and she had 
gone from nothing to this. The ACLU was supporting the bill, and EFF, 
EPIC, all the old civ-lib mafia had opened their arms to her. She 
poured herself one more very small whisky, gave herself a fragrant bath 
and put herself to bed, grinning like a fool.

\tb

“There are four news-crews, six print reporters, and a couple of 
others here to see you,” the office-manager said. The office phones 
were out again, but that hadn't stopped a fair number of determined 
people from figuring out that they could actually move their physical 
being from one part of Washington to another and have a real, 
old-fashioned face-to-face. The lawyer and she had each taken a dozen 
press “calls” that morning, with their embedded reporters from 
Columbia J-School perched obtrusively in the corners of their offices, 
taking copious notes and filming constantly.

“Others?”

“A mixed bag. Some Hill people, some I'm not sure about.”

Trish stood and stretched out her back, listening to it pop. She 
usually worked in bursts, typing or talking for an hour, then taking a 
little walk to gather her thoughts and touch base with her co-workers. 
Today, she'd been glued to her seat from 7AM to after lunchtime, and 
her back and butt were shrieking at her.

She walked into the front area, trailed by her reporter. She recognized 
some of the journos and some of the Congressional staffers, and a local 
rep from a European Privacy think-tank in Brussels, and -- Rainer.

He was turned out in a very natty suit and a homburg, a fashion that 
had recently come back to DC, and she knew that he'd been put together 
by a personal shopper. Her own Board had suggested to her, 
matter-of-factly, that she should get one of her own once the bill 
cleared committee, since she'd be doing tons of press and as sharp a 
dresser as she fancied herself, she was no pro. Her prodigious talents, 
they assured her, lay elsewhere.

He took her hand with both of his and gave her a long, intense 
hand-shake that drew stares from the journos and the think-tank man.

“Nice to see you again, Ms. McCavity,” he said, somberly.

“A pleasure as always, Mr. Feinstein,” she said.

“I'm sorry to drop in on you unannounced,” he said, “but I hoped 
that I could have just a moment of your time.” Belatedly, he 
remembered to take off his silly hat and then he fumbled with the right 
way to hold it, settling for dropping it to his waist and upending it. 
She thought he looked like a panhandler in a Charlie Chaplin movie and 
she suppressed a smile. His curly hair had been gelled into a careful 
configuration that reminded her of the glossy ringlets of a black 
poodle.

“I suppose we can do that,” she said. She turned to her other 
visitors. “Who's got a 3PM deadline?” she said. Two of the 
print-reporters held up their hands. “You then you,” she said. 
“Who's got a 5PM filing deadline? 6PM? 10PM?” She triaged them all, 
promised to meet the think-tank man for dinner at an Ethiopian place in 
Adams-Morgan, and led Rainer into her office and closed the door.

He looked at her embedded reporter and cocked his head.

“Sorry, Rainer,” she said. “I have a shadow for the duration. 
Just pretend he isn't here. You don't mind, do you dear?” she said to 
the reporter, who was very young and very bright and missed nothing. He 
shook his head and made some notes.

“The bill's dead,” Rainer said, after he'd sat down.

“Oh really?” she said.

“Just heard from Senator Rittenhouse, personally. He takes the 
position that this should be in Commerce, not Judiciary, and is calling 
hearings to make that happen.”

Rittenhouse was another powerful committee chairman, and this wasn't 
good news. What's more, he was in the pocket of the network operators 
and had been for a decade, so much so that editorialists and talk-radio 
types called him “The Senator from The Internet.”

Still, it wasn't catastrophic. “That's interesting,” she said, 
“but it's a far cry from killing the bill. It's pretty standard, in 
fact. Just slows things down.” She smiled at him. He was just a kid 
sometimes, so out of his depth here. He reminded her of the Relatives 
she'd met that day, the little boys in their miniature suits running on 
the beach.

He shifted in his seat and fondled his hat-brim. “Well, I guess we'll 
see. My press-liaison has set up a post-mortem debate on one of the 
news-networks tonight, and I thought you might want to represent the 
other side?”

She smiled again. He was twice the rhetorician that she was, but he had 
no idea how to play the game. She'd have to be careful to bruise, not 
break him.

\tb

“We, as a society, make trade-offs all the time,” Rainer said. He 
was wearing a different suit this evening, something that Trish had to 
admit looked damned good on the studio monitors (better than her frumpy 
blouse and wool winter-weight trousers). “We trade a little bit of 
privacy for a little bit of security when we show identification before 
going into a federal building --”

The ewok held up his paw. “But how much should we be willing to 
trade, Ms. McCavity?”

She looked into the camera, keeping her eyes still, the way she'd been 
told to if she didn't want to appear tourettic. “Wickett, when 
Franklin said, `Those willing to give up a little liberty for a little 
security deserve neither security nor liberty,' he wasn't spouting 
empty rhetoric, he was laying the groundwork for this enduring 
democratic experiment that we all love. Look, we're not opposed to the 
use of autonomous networks for \emph{some} applications, even 
\emph{most} applications, with appropriate safeguards and checks and 
balances. No nation on earth has the reliance that we do on these 
networks. Are they an appropriate way of advising you on the best way 
to get to the mall on a busy Saturday? Absolutely, provided that 
everyone gets the best advice the system can give, regardless of 
economic status or influence. But should they be used to figure out 
whom the FBI should open an investigation into? Absolutely not. We use 
judges and grand juries and evidence to establish the sufficiency of a 
request to investigate a private citizen who is considered innocent 
until proven guilty. We learned that lesson the hard way, during the 
War on Terrorism and the Ashcroft witch-hunts. Should we trade grand 
juries and judges for ant-colonies? Do you want the warrant for your 
wiretap issued by an accountable human being or by a simulated 
ant-hill?”

The ewok turned to the camera. “Both sides make a compelling case. 
What do you think? When we come back, we'll take your calls and 
questions.” The lights dimmed and it adjusted its collar and cracked 
its hairy knuckles on the table before it. Ever since it had made the 
move to a pbs, it had been grooming its fur ever-more conservatively 
and trying out a series of waistcoats and short pants. It turned to her 
and stared at her with its saucer-sized black button eyes. “You know, 
I just wanted to say thanks -- I had self-identified as an ewok since I 
was five years old, but Lucasfilm just wouldn't license the surgery, so 
I went through every day feeling like a stranger in my body. It wasn't 
until your law got enacted that I was able to find a doctor who'd do it 
without permission.”

She shook its paw. “It wasn't my law, but I helped. I'm glad it 
helped you out.” She unconsciously wiped her palm on her thigh as the 
ewok turned to his make-up boy and let him comb out its cheeks. She 
stared at Rainer, who wasn't looking good. She'd had him on the ropes 
since their opening remarks, and the ewok kept interrupting him to let 
her rebut -- and now she knew why.

Rainer had his phone clamped to his head, and he was nodding vigorously 
and drumming his fingers. He was sweating, and it was making his hair 
come un-coiffed. Trish's own phone buzzed and she looked down at in 
surprise. It was her voice-mail, coming back to life again. It had 
started when she got to the studio -- when she got within a few yards 
of Rainer, she realized. Messages coming in. She'd transcribed a dozen 
in the green-room before they'd dragged her into makeup.

The studio lights blinked and Rainer popped the phone back into his 
pocket and the ewok turned to look back into the camera, examining the 
ticker scrolling past his prompter. He introduced them again, then 
turned to Trish.

“Ms. McCavity, Alberto in San Juan writes in wanting to know what 
changes we should institute in the networks.”

She said, “It's not my place to say what technical changes the 
networks need to have. That's where experts like Mr. Feinstein come in. 
We'd ask the administrative branch to solicit comments from people like 
him to figure out exactly what technical changes could be made to allow 
us to remain competitive without giving up our fundamental liberties in 
order to beat the occasional traffic jam.”

“Mr. Feinstein?”

He grinned and leaned forward. “It's interesting that Ms. McCavity 
should disavow any technical expertise, since that's what we've been 
saying all along. If she's getting stuck in traffic, it's because 
there's a \emph{lot} of traffic. The ant-nets route \emph{five thousand 
percent more traffic} than our nation's highways ever accommodated 
without them, and they've increased the 
miles-per-hour-per-capita-per-linear-mile by \emph{six thousand, four 
hundred percent}. You're stuck in traffic? Fine. I get stuck sometimes 
too. But for every hour you spend stuck today, you're saving 
\emph{hundreds} of hours relative to the time your parents spent in 
transit.

“The other side of this debate are asking for something impossible: 
they want us to modify the structure of the network, which is a 
technical construct, built out of bits and equations, to accommodate a 
philosophical objective. They assert that this is possible, but it's 
like listening to someone assert that our democracy would be better 
served if we had less gravity, or if two plus two equaled five. Whether 
or not that's true, it's not reasonable to ask for it.”

The ewok turned to her.

She said, “Well, we've heard a great deal about the impossibility of 
building democratic fundamentals into the network, but nothing about 
the possibilities. This hard, no-compromise line is belied by the fact 
that we know that the rich and powerful manipulate the network to their 
own advantage, something that statistics have proven out --”

“See, this is \emph{exactly} how these Human-Readable types do it, 
it's how their media training goes. They are here to ask for changes to 
\emph{technical} specifications, but they disavow any technical 
knowledge, and when they're called on this, they spout dubious 
`statistics' that `prove' that up is down, black is white, and that 
millionaires can get to the movies in half the time that paupers can. 
The Emergent Network Suppliers' Industry Association represents the 
foremost experts in this field, but you don't need to be an expert to 
know that these networks \emph{work}. The ants take us where we want to 
go, in the shortest time, with the highest reliability. Anyone who 
doubts that can dig out her map and compass and sextant and try to 
navigate the world without their assistance, the way they do in 
Europe.”

Her mouth was open. \emph{Media training}? Where did he get this 
business about \emph{media training}? “I'm not sure where Mr. 
Feinstein gets his information about my media training from, but 
personally, I'd rather talk about networks.” She paused. “Let's 
talk about Europe, where they \emph{have} found ways of creating 
transparency and accountability for these `unregulatable' algorithms, 
where the sky \emph{hasn't} fallen and the final trump hasn't sounded. 
What do they know that we don't?”

“What indeed?” the ewok said, breaking in and giving her the last 
word again. “More of your questions after this break.”

They got in their cars together after they'd scrubbed off their makeup 
and shaken paws with the ewok, riding down in the elevator shoulder to 
shoulder, slumped and sweaty and exhausted. They didn't speak, and the 
silence might have been mistaken for companionable by someone who 
didn't know any better.

They got off at the same floor in the parking garage and turned in the 
same direction, and Trish spied his car, parked next to hers, the last 
two on the floor. Quickening her step, she opened her door and turned 
the car on, backing up so that she was right behind Rainer.

He backed out slowly, looking at her quizzically in his rearview, but 
she refused to meet his eye, and when he pulled out, she rode his 
bumper.

“Sweet fancy Moses,” she breathed, as the traffic parted before 
them, allowing them to scythe through the streets, onto the beltway. 
She hung grimly onto his bumper, cutting off cars that tried to shift 
into her lane. Moving this fast after so much time stuck on the roads 
-- it felt like flying. She laughed and then got a devilish idea.

Spotting a gap in the passing lane, she zipped ahead of Rainer and 
swerved back into his lane so that she was in the lead. As though a 
door had slammed shut, the traffic congealed before them into a clot as 
thick as an aneurysm. She hissed out a note of satisfaction, then 
waited patiently while Rainer laboriously passed her again, and the 
traffic melted away once more.

It was tempting not to get off at her exit, but she had to get some 
sleep, and so she reluctantly changed lanes. There wasn't much traffic 
on the road, but every traffic light glowed vindictive red all the way 
to her house.

The Chairman of her Board messengered over a hand-written note of 
congratulations that was on her doorstep. Beneath it was a note from 
Rainer's great-aunt, with the best wishes of his mother in neat pen 
beneath it. She read its kind words as she boiled the kettle, and put 
it into her pile of correspondence to answer. Rainer's great-aunt 
wanted to know if she had met a nice boy in DC yet, but she didn't come 
right out and say it -- too subtle for that. The women in Rainer's 
family got all the subtlety, and they recognized their own kind. It was 
why she and the old lady kept writing to each other; that and so that 
the Relatives could reassure themselves that someone in full possession 
of lifeguardly skills and a level head was watching out for Rainer's 
interests.

This business of hand-written, hand-delivered notes and letters was 
actually kind of charming, she thought as she put her feet up on her 
coffee table and opened up her flask of very special Irish whisky again.

\tb

She and Rainer went head to head in half a dozen more skirmishes that 
month -- her phone popping back to life every time she got within 
shouting distance of him. The on-again/off-again hearings in both 
Judiciary and Commerce never quite materialized.

She was better at playing the game, but he was a fast learner, and he 
had much deeper pockets and working network infrastructure. Her Board 
approved her renting out an empty suite of offices below their office 
and converting them to bedrooms for her staff for days when their cars 
couldn't get them home. They secretly borrowed elderly network 
appliances from relatives or bought them in the dollar-a-pound bin at 
the Salvation Army, but always, within a few hours of being in the 
possession of someone in the employ of the Association for a 
Human-Readable World, the devices would seize up and lose their routes 
to the network. Their offices started to fill up with dead soldiers, 
abandoned network boxes that no one could get online.

The embedded journalists went home after the second week. Their own 
gear was seizing up, too, as though the curse of the Association for a 
Human-Readable World was rubbing off on them. They vowed to return when 
things got interesting again, but they were of no use to anyone without 
working cameras, mics, and notepads.

Christmas came and went, and New Year's, and then February arrived and 
the city turned to ice and slush and perpetual twilight. The paralegal 
quit -- she needed a job where the phones worked so that she could call 
her girlfriend. The media guy took a series of “personal days” and 
she wasn't sure if he'd show up again, but it didn't matter, because 
the press had stopped calling them.

Then came the second Downtime.

It struck during morning rush-hour on Valentine's Day, a Monday, and it 
juddered the whole country to a halt for eight long days. The hospitals 
overflowed and doctors used motorized scooters to go from one place to 
another, unable to spread their expertise around with telemedicine. 
Firemen perished in blazes. Cops arrived too late at crime-scenes. 
Grocery stores didn't get their resupplies, and schools dug out old 
chalk-boards and taught the few students who lived close enough to 
walk. Fed cops of all description went berserk, and could be seen 
walking briskly from one federal building to another, their faces grim.

And suddenly, miraculously, every journalist, policy-wonk, staffer, 
advisor, clerk and cop in DC wanted to have a chat with the Association 
for a Human-Readable World.

\tb

She hired three more people that week, and borrowed four more from 
fellow-traveler organizations. Paying their salaries for the next four 
weeks would bottom out the group's finances, but she knew that this was 
now or never, and the Board backed her, after some nail-biting debate.

Rainer showed up on the fourth day of the Downtime, and she found him 
standing, bewildered, in the hustle of her office as her staffers 
penned notes on steno pads to their contacts on the Hill and handed 
them to waiting bicycle couriers in space-program warmgear that swathed 
them from fingertips to eyeballs. She plucked him out of the bustle and 
brought him back to her office.

“I've got a hell of a nerve,” he said, sitting in her guest-chair.

“Really?” she said. “I hadn't noticed.”

“Well, I haven't been showing it off. But I'm about to. I need 
advice. My office is falling apart. You've been living with no 
communications and no travel for a year now, you know how to make it 
work. We're completely lost. I've come to throw myself on your 
mercy.” He looked up at her with his big brown eyes, and then they 
crumpled shut as he made his Fretting face.

“You're playing me, Rainer,” she said. “And it won't work. 
Whatever I feel for you, I've got a job to do, and if this Downtime 
tells us anything, it's that I'm doing the right thing, and you're 
doing the wrong thing.”

He hung his head. He wasn't even the slightest bit natty that day. She 
supposed that his personal assistant was stuck in Fall's Church or 
Baltimore or somewhere, unable to get into the city. Judging from the 
slush and road-salt on his shoes, he must have walked the two miles 
between their offices.

“What's more, I don't have any advice to give you, in particular. 
We're not faring well here because we're doing something differently -- 
we're faring well because we're doing what we've been at all along, 
because of a network outage that you claim is impossible, is a figment 
of our imagination. Those bike messengers: we've been their best 
customers for months now. Everyone else is begging for service from 
them, but they're always here when we need them. We've got beds and 
changes of clothes and toilet-kits in the offices downstairs. We've 
been living through a Downtime for a couple of quarters now -- we've 
hardly noticed the change. If you want to cope as well as we are, well, 
you can go back in time, rent out spare offices to house your staff, 
establish a good working relationship with a bike-messenger company, 
learn to navigate the Metro and the freeways by map, and all the other 
things we've done here.”

He looked defeated. He began to stand, to turn, to leave.

“Rainer,” she said.

He paused.

“Close the door and sit down,” she said.

He did, looking at her with so much hope that it made her eyes water.

“Here's my offer,” she said. “You and I will lock ourselves in 
this office with the last draft of my bill. My staff will run 
interference for me with the Judiciary committee, and we will draft a 
version of my bill that we can both live with. We will jointly take it 
to Senators Beauchamp and Rittenhouse, with our blessings, and ask them 
to expedite it through \emph{both} committees. Every Congresscritter on 
the Hill is sitting around with his thumb up his ass until the lights 
come back on. We can get this voted in by Tuesday.”

He stared down at his hands. “I can't do it,” he said. “My 
\emph{job} is \emph{not to compromise}. I just can't do it.”

“Come on, Rainer, think outside the box for a minute here.” Her 
heart was pounding. This could really be it. This could be the solution 
she'd been waiting for. “Even if the bill passes, there's going to be 
a long deliberation over the contours of the regulation, probably at 
the FCC. You'll be able to work on the bureau staffers and at the 
expert agencies, take ex-parte meetings and lobby on behalf of your 
employers. It's all we've ever asked for: an expert discussion where 
the public interest gets a hearing alongside of private enterprise and 
government.”

But he was shaking his head, standing up to go. “You're probably 
right, Trish,” he said. “I don't know. What I know is, I can't do 
what you're asking of me. They'd just fire me.”

“If the Downtime continues, they won't be \emph{able} to fire you -- 
they won't even know what you're up to until it's too late. And then 
they'll make the best that they can out of it. No one is better 
qualified to represent your side in the administrative agencies.”

He put his ridiculous hat on and wrapped his scarf around his neck, and 
they looked each other in the eyes for a long moment. She waited for 
the involuntary smile that looking into his eyes inevitably evoked, but 
it didn't come.

“I don't understand you, Trish. You won this incredible victory for 
cooperation, for collective ownership of our intellectual 
infrastructure. Ant-networks demand the same cooperation from the 
nodes, that my phone pass your car's messages to his desk. Let's just 
set aside the professional politics for a second. Just you and me. Tell 
me: how can you \emph{not} support this?” He looked at her out from 
under his brows, staring intensely. He swallowed and said, “It was 
the surfer, wasn't it?”

“What?” she said.

“The one who died. That's why you're doing this. You want to make up 
for him --”

She couldn't believe he'd said it. Taken such a cheap shot. “I'm 
surprised you didn't save that one for television, Rainer. Jesus. No, 
I'm doing this because it's \emph{right}. In case you haven't noticed, 
your self-healing, uncorruptible network is \emph{down}. People are 
suffering. The economy is tanking. The death toll is mounting. You 
won't even bend one \emph{inch}, one \emph{tenth of an inch}, because 
you're worried about losing your job.”

“Trish,” he said, “I'm sorry, I didn't mean --”

Her office door opened and there stood her embedded journalist. “I 
just got in from Manhattan,” he said. “Can I set up in that corner 
there again?”

“Be my guest,” she said, grateful for the distraction. Rainer 
looked at her, forehead scrunched, and then he left.

\tb

“It's a good thing you're not over him,” the lawyer said, pouring 
her another victory whisky. The bill had passed the House with only one 
opposing and two abstentions, and had squeaked through the Senate by 
five seats, at five minutes to midnight on the eighth day of the 
Downtime. They were halfway to the bar (where the office manager had 
been feeding twenties to the bartender to stay open) when the grid came 
back up, crawls springing to life on every surface and cars suddenly 
zipping forward in the characteristic high-speed ballet of efficiently 
routed traffic. They'd laughed themselves stupid all the way to the bar 
and after a brief but intense negotiation between the lawyer and the 
barman, he'd produced a bottle of Irish that was nearly half as good as 
the stuff Trish kept at home.

“I'm going to pretend you didn't say that,” Trish said, sipping 
tenderly at the booze.

“Come on, girl,” the lawyer said, twirling her moustache. “Be 
serious. You two had so much sexual energy in that room, it's a wonder 
you didn't make the bulbs explode. It's how you got inside each other's 
heads. You weren't selling the committee, you were selling \emph{him}, 
and that's what made you so effective. We're going to need that again 
at the FCC, too -- so no getting over him until after then.”

Trish drank her whisky. She didn't know what to say to that. He'd 
looked ten years older tonight, in the corridors, whispering to his 
committee members, to his staffers, his face drooping and wilted. She 
supposed she didn't look any better. It had been, what, three days? 
since she'd had more than an hour's sleep.

“I don't get it,” she said. “How could he be so dumb? I mean, 
it's obvious that the system is being gamed. Obvious that we're being 
targeted through it. Yet he sits there, insisting that white is black, 
that up is down, that the network is autonomous and immune to all 
corruption.”

“It's like a religion for them,” the lawyer said. “It doesn't 
need explaining. It's just right-living. It's the Law.”

Trish thought back to the ceremony in the graveyard, the dirge and the 
prayers to a god no one believed in. Had Rainer really renounced his 
faith when he dropped out of Yeshiva?

“Here's to a human-readable world,” Trish said, raising her glass. 
Around her, the staffers and borrowed staffers and hangers-on and even 
the barman raised their glasses and cheered. It was warm and the 
feeling swelled in her tummy and up her chest and through her face and 
she burst out in what felt like the biggest smile of her life.

\tb

She'd learned a long time ago never to send email while drunk, but it 
had been too much last night.

“What if, Rainer, what if -- what if the reason for the Downtimes is 
that someone is manipulating the network and that's breaking it. Did 
you ever wonder about that? Maybe the network \emph{is} as good as you 
say it is -- until someone screws it up by trying to get preferential 
treatment for his pals.

“Wouldn't that be a kick in the teeth? We get five squillion percent 
increases in across-the-board routing efficiency, but in the end, it's 
never enough for people who can't be happy unless they're happier than 
someone else.

“The thing that saves the human race, but if we adopt it, it will 
destroy us. Irony sucks.”

She'd signed it “Love,” but even drunk, she'd had the sense to take 
that out before sending it. Saying “Love” would have been no more 
appropriate than saying, “You know, I \emph{did} save your cousin's 
life.” She'd called in no favors, she'd run no blackmail, and she'd 
won anyway.

He rang her doorbell at 5AM. She was barely able to drag herself out of 
bed.

“I figured you'd be getting up to deal with the press soon,” he 
said, and she groaned. He was right. She'd earned some time off, but 
it'd be a month before she could take it. Too much press to do. She 
appreciated anew how much work it must have taken to be any of her old 
bosses from the copyright wars: the judge, the senator, the executive 
director of the PAC.

She was in her robe, and he was in jeans and a UCLA sweatshirt. He 
didn't have any gel in his hair, which was matted down by the knit cap 
he'd been wearing. He looked adorable.

“They fired me this morning,” he said.

“Oh, hon --” she said.

“I would have quit,” he said. “I'm outmatched.”

She felt herself blush. Or was she flushing? She was suddenly aware of 
his smell, the boy smell, the smell that she could smell in his chest, 
in his scalp, in his tummy, lower\ldots{} She straightened up and led him 
into the living room and started the coffee-maker going.

“When do you fly back, then?” she said.

He looked at her, smiling. “I don't know,” he said. “I haven't 
booked a ticket.”

She felt an answering smile at the corners of her mouth and turned into 
the fridge to fetch out some gourmet MREs. “Bacon and eggs or 
pancakes?” she said, then laughed. “I guess bacon is out,” she 
said.

“Oh, I'm willing to bet that that bacon hasn't been anywhere near a 
pig,” he said, “but I'll have the pancakes, if you don't mind.”

She set everything to perking and went into the bedroom to pull on 
something smart and camera-friendly, but everything was in the hamper, 
so she settled for jeans and a decent shirt from last-year's wardrobe.

When she opened the door, he was standing right there, taller than her. 
“I think you're right,” he said. “About the network. It's the 
best explanation I've heard so far.”

She wrapped herself in silence again, waited for him to say more.

“You see, the \emph{true}, neutral network is immune to corrupting 
influences and favoritism. So the existence of corruption and 
favoritism means that what we've got \emph{isn't} a true network. Which 
means you're right! We need to have a hearing to get to the bottom of 
this, so that we can build the true network.” He smiled bravely. “I 
thought maybe you could use an expert in your corner who'd say that in 
a hearing?”

“Thanks,” she said, and slipped under his arm and back into the 
kitchen. Suddenly, she wanted very much to be back at her office, back 
with her staff, talking to reporters and overseeing a million details. 
“I'll think about it.”

“I'm giving up my apartment at the end of the month -- next Monday. I 
won't be able to afford it without the Association's salary,” he said.

Her place was big. A bedroom, a home office, a living room and a dining 
room. It was a serious deal for DC, even outside the beltway. It could 
easily accommodate a second person, even if they weren't sleeping 
together.

Her office -- her staff -- the press -- the bill -- her Board.

“Well,” she said, “I've got to get going. I'll shower at the 
office. Got to get there in time to catch the Euro press-calls. Let me 
put your breakfast in a bag, OK?”

He looked whipsawed. “Uh, OK. Can I give you a ride?”

“No, I'll need my car this afternoon. Thanks, though.” She kept her 
voice light, didn't meet his eyes. Kept thinking: her office -- her 
staff -- the bill.

“Well,” he said. He turned for the door. Stopped. She tensed. He 
turned back to her. “Trish,” he said.

“It's OK,” she said. “It's OK. We just have religious 
differences, is all.”

She slipped past him and into her car, and left him standing in her 
driveway. As she asked the car to plot a route for her back to the 
Hill, she dug through her purse for a pocket-knife. At the next red 
light, she took her lapel and slashed at it, opening a rent in her 
shirt that reflected a little of what her heart was feeling. It made 
her feel a little better to do it.

-- For Alice

\section{Afterword:}

I wrote this story for Alice, who was then my girlfriend and is now my 
wife. Alice is smart as anything and I wanted to show off to her.

So I stole two ideas: Eric Bonabeau's ground-breaking work on 
“ant-colony optimisation” (the basis of his consulting firm 
Icosystem, built on work he did at the Santa Fe Institute) and Natalie 
Jeremijenko's provocative notion of “legible computing” -- that is, 
computing whose results can be interpreted by laypeople. Eric's handing 
the keys to the kingdom to emergent systems that are spookily good but 
can't be interrogated. Natalie wants a Protestant Reformation for 
computing in which we can all understand what's being discovered by our 
computer systems. These two ideas are both fantastic, but they can't 
peacefully coexist, or can they?

\end{document}
