\input{common/hyp-en}

\begin{document}
%\setlength{\emergencystretch}{1ex}
\raggedbottom

\begin{center}
\textbf{\huge\textsf{Liberation Spectrum}}

\medskip
Cory Doctorow

\end{center}

\bigskip

\begin{flushleft}
This story is part of Cory Doctorow’s short story collection
“With a Little Help” published by himself. It is licensed under a
\href{http://creativecommons.org/licenses/by-nc-sa/}
{Creative Commons Attribution-NonCommercial-ShareAlike 3.0} license.

\bigskip

The whole volume is available at:
\texttt{http://craphound.com/walh/}

\medskip

The volume has been split into individual stories for the purpose of the
\href{http://ccbib.org}{Creative Commons Bibliothek.}
The introduction and similar accompanying texts are available under the 
title:
\end{flushleft}
\begin{center}
With a Little Help -- Extra Stuff
\end{center}

\newpage

\section{Liberation Spectrum}

The tiny multinational lumbered across the Niagara Falls border in its 
tour-bus, Lee-Daniel at the wheel, sipping iced mocha from the flexible 
straw that he'd threaded through a series of eyelets on his jacket. 
He'd been driving all the way since Akwesahsne, reciting mnemonic 
sleep-dep chants and steadily consuming the lethal blend of bittersweet 
chocolate and espresso, but after 20 straight hours he was in deadly 
danger of falling straight to sleep and head-onning the bus into a 
Jersey barrier or a bullet train or a minivan.

Once they were on US soil, he pulled the bus over at a temporary 
roadhouse and set the handbrake. He eased himself out of the driver's 
perch, chafing his narrow ass and thighs to get the blood flowing there 
again, and gave forth a drawn out “\emph{gaaaah}” as the pins and 
needles stabbed into his sweat-marinated muscles. He heard the rest of 
the company rousing itself behind him. First, the investors in the 
front row, then the rest of the board of directors in the row behind 
them, then four rows of middle-managers and finally the great mass of 
front-line workers, techs, customer service reps, trouble-shooters, 
antennamen, switchwomen, chicken-pluckers and left-handed 
bottle-stretchers.

He flipped the windows to transparent and let the sun shine in, 
provoking groans from the company. MacDiarmid, the angel investor who'd 
been in since the multinational had been able to fit in a sedan, threw 
a strong arm around Lee-Daniel's shoulders. “You OK?” he said. The 
tone had phony solicitousness; MacDiarmid and Lee-Daniel had been 
through half a dozen disasters, from hostile takeover attempts to 
roadblocks to high-speed engine failure, and Lee-Daniel knew a fake 
when he heard it.

“I'm fixing to lay down and die,” Lee-Daniel said, stretching 
theatrically, his pipe-cleaner arms straining. “You're street-legal 
in New York, right? How about you drive the bus for the next couple 
shifts?”

“Seriously?” MacDiarmid said. His black hair was showing grey now, 
but his eyebrows were still fierce and black, his eyes still sharp in 
their nest of whiskey-cured crows-feet.

It was rare for Lee-Daniel to cede the wheel to anyone else -- it was 
his damned company and he'd drive the damned bus. Lee-Daniel saw the 
shareholder confidence eroding before his eyes.

“Just for a while, OK? Not permanent, just for a day or two, just 
long enough for me to get over the sleep-deficit and re-grow some 
stomach lining.” It was hard being CEO of a mobile multinational. The 
shareholder oversight was murder.

MacDiarmid looked closely at him, then smiled and gave him a burly 
man-hug that smelled of sandalwood soap and good liquor. “Yeah, of 
course, of course. I'll put it to the Board Meeting tonight at dinner. 
Can't have the CEO burning out at the wheel, that's what I'll say, 
don't worry about it, LD.”

“Thanks, Mac,” Lee-Daniel said. “How about we get some eats?” 
He put his hand on the geometry-reader beside the wheel, 
re-authenticated to the bus, then hit the hatches. Doors hissed open at 
the back, at the front, at the middle, fresh dusty air rushing in all 
at once in an ear-popping whoosh. The bus knelt ponderously and the 
company piled out.

MacDiarmid hustled away to join the rest of the investors, his 
exquisite hand-made leather shoes slapping the paving, the cuffs of his 
wool tailor-made slacks shushing over their gleaming upper, and as 
Lee-Daniel locked the bus down and armed it up, he watched the angel 
investor whisper in his co-shareholders' ears. Lee-Daniel couldn't hear 
the words, but six years at the wheel of Cognitive Radio, Inc. had 
schooled him well in the body-language of investors and he knew his 
days with CogRad were numbered.

\tb

The roadhouse was the kind of TAZ that got less entertaining at the 
square of the amount of time spent within its animated walls. The first 
minute was painful, an overbright eternity of authenticating to the 
roadhouse-area-network and establishing credit with the system. Once 
they had their tokens -- poker-chips adorned with grinning, dancing 
anthropomorphic dollar, Euro and Yen symbols -- there came the second 
minute, twice as horrible as the first, as they struggled in the guts 
of the giant vending machine, trying to fathom the actual products 
represented by the branded messages that tailored themselves to your 
personal demographic, your stated and implicit preferences, the 
messages that danced across your field of vision as you perused the 
racks in the roadhouse's aisles.

The third minute was twice as horrible as the first two minutes, as you 
finalized your selections by waving your poker-chip at different 
displays, then tried to take receipt of your goods from the floor-level 
fulfillment chutes while fending off the imprecations of the upsell 
displays set into the floor-tiles. “Lee-Daniel! People who bought 
tuna-melts also bought thousand-hour power-cells. People who bought 
OralCare mouth-kits also bought MyGuts brand edible oscopycams. People 
who bought banana-melatonin rice-shakes also bought tailor-made 
sailcloth shirts by Figaro's of London and Rangoon.”

The horribleness of the roadhouse went asymptotic to infinity at minute 
four, as you sat down and tried to eat your rubbery tuna-melt hunkered 
down at a table crowded with middle-managers in need of reassurance 
while swatting away the buzzing aerostats that probabilistically routed 
towards those diners with the highest credit ratings, delivering 
pitches whose tone and content had been honed by genetic algorithms 
that sharpened them to maximal intrusiveness and intriguingness. It 
took a genetic algorithm to make a high colonic sound like an afternoon 
at a spa.

“I'm getting too old for this shit,” said Joey Riel, a 17-year-old 
metis whose fluency in English, French and Ojibwa had made him the 
youngest middle manager in CogRad history, eight months before. He'd 
started griping about his road-weariness within days of his promotion 
up from antennaman. It made him fit in with the other, older middle 
managers, who were coffee-soured lifers whose time on the road had 
drummed out any footloose spirit they might have once possessed.

Further down the arcade, the investors were waving their tokens over a 
trading table, playing the instant futures market. An aerostat overhead 
mirrored the gameplay, and as Lee-Daniel watched, MacDiarmid doubled 
his money on a short-odds bet on two cherries and a lemon, then 
Earnshaw lost big when his long-odds investment on uranium and coal 
came back with two windmills and a photovoltaic array.

“Amen to that, bro,” said Elaine, who ran two squads of surveyors. 
She was all lean muscle and blackfly repellent and mail-order 
outdoorwear, handily capable of living off the land for weeks while 
trekking the bush, homing in on optimal repeater locations. At the 
Akwesahsne Sovereign, she'd broken the hearts of a half-dozen 
starry-eyed Mohawk Warriors who'd puppydogged after her as she shlepped 
the length and breadth of their territory, warchalking neon arrows to 
indicate RF shadows cast by especially leafy trees and outcroppings of 
granite Canadian Shield. That was before the Surete du Quebec arrived 
on the scene and it all went pear-shaped.

“Me, too,” said Mortimer, the security man who really \emph{was} 
too old for this shit, Lee-Daniel reflected, scratching at his fussy 
little caterpillar moustache. He'd been protecting the old dodderer 
from the Board of Directors, who saw him as an insurance nightmare. 
Mortimer's hands shook, he was nightblind, and he was 98 years old, and 
there wasn't enough rejuve in the world to give him the mental 
flexibility required by the modern age. Lee-Daniel had stripped him of 
his sidearms, even the nonlethals, at the same time as he'd promoted 
Joey Riel. Now Mortimer carried a loudhailer through which he could 
bark orders in his old cop voice, the voice that made your asshole 
clench up and your shoulders itch for a soon-come bullet.

The investors howled again, and the aerostat told them all that 
MacDiarmid had cleaned up bigtime, paying out 100-to-1 on an investment 
in Shell Oil collectibles -- two derricks and a shell. The Series 
A/Series B investors crowded around him, giving him awe-struck 
back-slaps. The other two might be the fronts for gigafunds, but that 
was all they were: fronts. They were the Voice of the Money while the 
company was on the road, junior associates who needed to make a good 
score on their wanderjahr if they wanted to make partner. Mac was solo 
money, a shrewd individual investor who'd acquired his 15-share in 
CogRad with no more investment than a year's worth of gas and roadhouse 
meals while Lee-Daniel was getting the show on the road.

“The rich get richer and the poor get children,” Joey Riel said, 
shaking his head at the investors and the board carrying MacDiarmid off 
to a private dining room for their dinner and nightly board-meeting.

“Those Mohawks got you all full of bolshy horseshit, didn't they?” 
Mortimer said. The Mohawk Warrior Society talked a good 
anarcho-syndiclist line. Most of the Sovereigns that CogRad unwired 
were interested in setting up a local telco as part of some economic 
development scheme -- as far as they were concerned, tax-free packets 
were the new tax-free cigarettes.

But the Mohawk Warriors in Quebec were in it for the samizdata. They 
had big plans for their cognitive radio network. They'd peered with two 
upstate New York networks and an Algerian satellite backbone, and they 
were reselling enciphered proxy-time on their network to anyone who 
wanted it, providing an anonymizing relay for any and all data, 
regardless of origin, destination or payload.

Lee-Daniel knew he should have gotten them to pay upfront. Nothing got 
the blackshirts interested in private wireless networking like routing 
suspicious real-time chatter between Burmese guerrilla cells and 
suspected movie-swappers in DC. But that wasn't how CogRadio had been 
built. The native bands that were desperate enough to assert that their 
ancestral treaties didn't encompass the RF spectrum couldn't afford to 
lay out cash for CogRadio's hardware, training and remote 
administration. CogRadio was as much a bank as a technology startup.

But the Canadian government took a hard line on anything that looked 
like separatism. Four CogRadio employees who'd been unlucky enough to 
get stuck on the wrong side of the barricades during one of the 
Warriors' traditional shoot-outs with the Surete wouldn't be coming 
back to work for 10-to-15, eight with good behavior.

With the investors off out of sight, the managers and the front-liners 
shucked their veneer of civility and began to get wild, ordering drinks 
and health-insurance-invalidating carbo-treats. Elaine sucked down 
three tequila cartons and glared bleary hostility at him.

“The fuck do you know about \emph{anything}, hey?” she said. 
“You're supposed to be in charge of things, but you don't take a shit 
without clearing it with those bastards.” She jerked her head over 
her shoulder at the closed door of the private dining room. “And when 
you \emph{do} make a decision, you fuck it up.” The smell of old 
sweat and booze made his eyes water.

Mortimer hitched himself erect, creaking up from his seat. “That's 
enough of that,” he said in his cop-voice, laying a still-strong hand 
on Elaine's shoulder. “If you don't like your job, you can give 
notice, but you'll keep it polite as long as you're working here.”

Elaine tried to shake his hand off, but he kept his grasp firm. 
Lee-Daniel had been through one or two of these in the first year, and 
he knew that Mortimer knew what he was doing. Things could get awfully 
heated up at times like this.

“You're hurting me,” Elaine said. “Let go.”

“Apologize to the man,” Mortimer said, the voice of authority. 
“You're out of line.”

Joey Riel leapt on Mortimer's back, his arms locked around Mortimer's 
neck. “Don't you touch her, you pig,” he hissed. Mortimer took hold 
of Joey's thumb and twisted it into a come-along and Joey let go, 
dancing around and clutching his hand.

“You broke my fucking thumb!” he said, and then Elaine was on her 
feet, shouting incoherently, right up in Mortimer's face, darting her 
head at him like a striking cobra. The front liners broke off their 
gaming and boozing and necking and rushed over, hooting for blood.

Lee-Daniel felt the old adrenalin, the “leadership” brain-reward 
that he got when it all came down to a crisis. He jumped up on their 
table, scattering their dinners' active packaging, which curled and 
waved as it flapped to the floor, cycling through its upsell ads.

“Enough!” he roared. It wasn't a cop-voice, but it was a voice 
nevertheless -- the voice of the man who signs the paycheck, the 
disappointed father who was going to turn the bus around and take the 
company home \emph{this instant} if he didn't get respect. Lee-Daniel 
didn't have to use that voice often, but its rarity was part of its 
effectiveness.

It didn't work. Elaine still shouted, Joey Riel was digging through the 
drifts of trash for a weapon, and the front-liners were still cheering 
their bosses on. “\emph{Enough}!” he said again, just to check, but 
it didn't work any better the second time around.

He got down off the table and circled Mortimer, who had the mic for his 
loudhailer clipped to his belt. Lee-Daniel snatched it up and hit the 
Talk button, dialing the volume up to max with his thumb.

“\emph{Enough}!” he said, and the loudhailer amplified his voice to 
staggering volume. At max, it was meant to be used to signal passing 
aircraft. Inside the vending-machine's claustrophobic bowels, it was 
like a bullet ricocheting through their skulls. Some of the more 
delicate antennamen dropped to their knees, their hands clutched to 
their heads, and Mortimer staggered back into Lee-Daniel, nearly 
knocking him off his feet.

Lee-Daniel cut the volume in half and hit talk again. The company shied 
back when the speaker array on Mortimer's bandolier \emph{pop}ped to 
life. “All right, enough. Company meeting. Get chairs, sit on the 
floor, whatever. Right here, right now.” He handed the mic back to 
Mortimer, who wiped it down with care and clipped it back to his belt.

He gave Mortimer his poker-chip. “Get a bag of ice for Joey,” he 
said. “And thanks, man.”

Mortimer gave him the cop-stare and trudged off to one of the vending 
banks and started prodding methodically at its display.

“All right,” Lee-Daniel said, again, looking into the expectant, 
upturned faces of his company. “All right.

“It's been a rough week for all of us. But we've had rough weeks 
before. Remember Wisconsin? That was in our first year, and the FCC 
looked like it was going to impound every bit of gear we owned. We 
spent a month on the Reservation, borrowing and borrowing to pay off 
the lawyers. No one got paid. I ate enough venison and corn-bread to 
last me a lifetime.”

Wisconsin was legendary. That was when they'd acquired the Series A 
investor, who'd converted its debt-instruments to capital when the Lac 
du Flambeau Ojibwa signed up their thousandth customer and started 
sending royalties to the company's e-gold account. Only Mortimer had 
actually been with the company long enough to remember the incident 
first-hand, which was good, since three quarters of the company had 
quit after the third week and he'd had to fire two more when they got 
caught hitting on some of the women on the Rez.

“But we saw it through,” he said, looking significantly at 
Mortimer, who kept mum. “And we'll see this through. You might think 
that it was a mistake to go to Canada, and I understand why it might 
seem that way to you. But let me put your mind at ease.

“It wasn't a mistake.

“It was a risk that we took to expand this business. If you want your 
options to be worth something, someday, this company's going to have to 
\emph{grow}. We've been growing at 20 percent per quarter for the past 
three years, and that's right on track. Maintaining that growth is 
going to necessitate excursions out of the USA. We'll be going back to 
Canada -- better prepared, wiser, more cautions -- but we'll be going 
back. The Caribbean, too. South America and Mexico. I shouldn't have to 
tell \emph{you} that radio has no borders. Wherever there's 
unencumbered spectrum, we'll be there. There's \emph{never} going to be 
a `routine' job, whatever that means. Every job will be different. If 
you're looking for a `routine' job, you're in the wrong business.

“We're headed for the Seneca sovereign in Cattaraugus next. There'll 
be a week of R{\&}R there: fishing, hunting, gaming. They have a decent 
theater there that's doing a Beckett revival, and I've got half-price 
tickets if you want `em.

“Half-price tickets for those who stay, that is. Because I want to 
make one thing clear: If you don't like the way I run this company, you 
shouldn't put up with it. Give me your notice, I'll cut you a check and 
you can get lost. That's your remedy. That's your \emph{only} remedy. 
I'll be sitting right here, any of you want to give your notice 
tonight.”

He sat down at a table and helped himself to someone's carton of 
crantini, gave it a shake to cool it down, then took a nonchalant sip.

The silence was broken by the door to the investors' dining room 
hissing open. The Series A investor stepped out into the chaos of the 
main concourse and crooked a finger at Lee-Daniel.

“We'd like to speak with you, if we may,” he said, and swung the 
door wide.

\tb

Akwesahsne was supposed to be a cakewalk. The Canadian Radio and 
Television Commission -- Canada's RF Feds -- were softies, more worried 
with ensuring that 30 percent of the entertainment product on the 
airwaves was “Canadian Content” than with monitoring 
ultra-low-power, ultra-wide-band cognitive radio experiments in rural 
Quebec. The Mohawk Warrior Society, whose reservation was a Siamese 
twin with another Rez in upstate New York, were accustomed to the 
American way of doing biz, had even underwritten MBAs for a bunch of 
the bros, which explained the animated growth-charts back-linked to 
hundreds of diverse spreadsheets maintained by research committees 
across the continental Mohawk Nation infrastructure.

But they did indeed talk a line of bolshy horseshit in the Mohawk 
Warriors Society.

The first hint came from the guard in the pillbox at the Akwesahsne 
main-gate. The CogRadio magic bus pulled up, abuzz with new-gig energy, 
the anticipation of thirty skilled professionals who'd been crammed 
into a bus for four solid days, ready to tear each others' throats out. 
The gate-woman was all of seventeen, not that you could tell at first, 
so crufted-up was she with obsolete martian armor/arms and sensory 
array.

But once she came onto the bus for her customs inspection and removed 
her immersive headgear, it was obvious that she was no older than the 
switch girls who drifted in and out of the CogRad bus, using it as a 
means of making a little e-gold between footloose adventures in the 
Great American Heartland.

A seventeen-year-old with a defensive array of fast-acting 
anti-serotonin misters was a lot less threatening than a 
thirty-year-old would have been, and orders of magnitude less 
terrifying than a similarly armed innovation-sick fifty-year-old would 
have been. Joey Riel came forward, stinking of something between 
sweat-socks and Doritos, and greeted her in familiar, colloquial 
French, something flirty by the sound of it, and she gave him a wry, 
patronizing smile.

“Why do you speak French, Brother? Why not greet me in Kani\-ien'keha, 
or Cree, or even Ojibwa? When we speak whiteman words, they make us 
think whiteman thoughts.” She turned to the bus and gave them a long 
stare. “Hello, whitemen,” she continued, “hello whitewomen. 
Welcome to the Mohawk Warrior Society autonomous zone. No weapons. No 
sex with First People. No drinks or drugs. No whiteman tobacco.

“Cook your own meals, wash your own plates, step lightly on the land. 
You can observe our nightly meetings if you are respectful, but it's 
more important that you come to the seminars afterwards. There are 
lectures, role-playing exercises, personal storytelling, theater of the 
oppressed, newsblogging, warblogging, linkblogging, puppet-making, 
outreach, filterbusting. Whiteman guests are welcome here, provided 
that they're willing to help the cause.”

Lee-Daniel had heard variations on this speech before, but they usually 
came from hotheads who argued against renewing CogRad's maintenance 
contract, not the official greeter before they'd even started the gig. 
He knew well enough to take it in stride and move on, but Joey Riel was 
blushing furiously at having been shot down for insufficient indianity 
by this highly macha hottie, and so he waved some verbal dick, asking 
something in Ojibwa, all testicular.

She fixed him with a withering stare. “You're not the first apple 
I've met,” she said. Apple -- red on the outside, white on the 
inside. “And you're not the most pathetic. But you're an apple and 
you've forgotten who you are, and that means that you don't mean 
anything to me except a sad story and a warning to other First 
People.”

Joey Riel's hands balled up into fists and the investors shifted 
nervously. Lee-Daniel got to his feet and interposed himself between 
them.

“Ya-tay-hay, madam,” he said. “Thank you for your welcome. Can 
you tell me where I should park the bus? We've got a lot of work to do 
today, while there's still light to work by.”

\tb

“You need to understand, it's not \emph{personal},” MacDiarmid 
said, for the third time.

Lee-Daniel set down his ridiculous second-hand crantini carton and 
climbed slowly to his feet. “\emph{You} need to understand, Mac, that 
I don't \emph{care} if it's personal. Whether you're forcing me out of 
this company, this company that \emph{I} built with my own two hands, 
this company that is hitting every goddamned milestone, this company 
that is returning good dividends on your preferred stock, whether 
you're forcing me out because you're not \emph{my friend anymore} --” 
he said this in a pinched, Mickey Mouse voice “-- or whether you're 
forcing me out because you think that it's `for the best' doesn't 
matter to me at all. I don't care if you're doing it because you're 
protecting your investment or because your astrologer told you to, I 
still won't stand for it.”

The Series A and Series B investors, who'd started off looking 
uncomfortable, visibly squirmed during this. They weren't accustomed to 
interpersonal conflict in the course of conducting their affairs. But 
Mac took it all in stride. Angels have to be prepared to slug it out to 
protect their investment.

“You don't get to stand for it, LD,” MacDiarmid said, sipping at a 
frosty can of slushy ginseng-infused Long Island iced tea. “You don't 
get a say in it. When the investors are united, you don't have the 
equity to overrule us. The severance package is generous, the 
noncompete is lightweight, and you get to go with your dignity 
intact.” He didn't need to add that fighting the board would mean a 
significant change to that picture.

“What'd they promise you, Mac?” Lee-Daniel asked. He'd shrewdly 
chosen his investors for their mutual animosity, believing that bitter 
enemies like the Series A gigafund and the Series B terafund would 
never come together, and that Mac, who'd been screwed on deals by 
principles from both funds, would never toss his lot in with them. 
“What do they have that's worth your throwing away this entire 
investment?”

“No one's throwing away anything. There comes a point in any 
business's life-cycle where the founders get out of their depth and we 
need to transition in a professional CEO. You've done a good job with 
CogRad, LD, and we recognize that, but if we're going to ensure steady 
growth, we need seasoned leadership.”

“Seasoned?” He barked a laugh. “Mac, I \emph{invented} this 
business! We're five years ahead of our closest competitors -- who only 
got that far by copying stuff \emph{I} invented. Who the hell could 
possibly be more `seasoned' than me?”

“You've never run a Fortune Five company,” the Series A man said. 
“You've never had more than fifty people working under you. Executive 
search firms --”

MacDiarmid waved a hand crusted with three class-rings at the 
gesticulating Series A punk, who barely looked old enough to smoke. 
He'd only been out of B-school for a year and he'd only been on the bus 
for a month, but here he was, telling Lee-Daniel that they'd blown 
corporate funds, \emph{money he'd earned}, on a slick-ass headhunter 
who'd spent it getting old frat-brothers laid at fancy hotels on 
Hawai'i while negotiating how much of Lee-Daniel's company they would 
end up with once they stole his job from him.

The punk shut up.

“Mac,” Lee-Daniel said, sitting down again, pulling up a chair. 
“Come here Mac, take a seat, talk to me. I want to hear this from 
you, from the beginning.”

Mac stood, exchanging significant looks with the Series A and Series B 
investors.

“Come on, Mac, screw that. You and me, end-to-end.” That was CogRad 
jargon from back in the old days. The Internet was end-to-end, which 
meant that any two points could communicate without an intermediary 
interfering in the bytestream. In CogRad, you didn't talk 
person-to-person or man-to-man, you talked end-to-end, just like the 
connectivity they brought to the Rez. “I own fifteen percent of this 
company, same as you -- you owe me a decent explanation.”

MacDiarmid stood fast.

“Get in the fucking chair, Mac,” Lee-Daniel said, hating the whine 
in his voice. “If you want me to go along with this, get in the 
fucking chair.

“Mac, I'm sorry. Sorry if I flew off the handle. I'm a grownup, 
you're a grownup and we both care about CogRad. Get in the chair and 
tell me about this. Please.”

MacDiarmid sat.

“Listen up, LD. You've done excellent work here. Be proud. You 
started something good, something that will grow and grow and that you 
can retire on. But you can't keep this up forever. If I thought for a 
second that you'd take orders from someone else, I'd offer to keep you 
on as COO or VP of Research and Development. There's no way, though -- 
no one would ever be able to tell you what to do in this company.

“You're great at the dirty work. You can get a crew onto a Rez, get 
the terminals sited and installed and burned in. You can boss a bunch 
of egomaniacs and social retards on long road-trips. For six years, 
we've needed someone at the helm who could do all that stuff.

“But it's time to settle into the next phase. We're going abroad, and 
it needs a delicate touch. If \emph{Canada} ends up in a firefight, 
what'll it be like in \emph{Guatemala}?”

\tb

Lee-Daniel and his people had had to work around a lot of surveying 
constraints. There had been a burial ground at the Moapa River Indian 
Reservation that was freaking \emph{perfect} for a repeater-array, with 
a commanding view of the entire goddamned Rez. But no matter how 
tempted the Paiute elders were by the thought of getting out of the 
cutthroat slots biz and instead taking a piece of every casino's action 
by offering secure connectivity for phones and data, they couldn't see 
their way clear to permitting CogRad's surveyor crew to head up there 
and start hammering in stakes for the repeaters.

The Akwesahsne Warriors took the cake, though. A fat, middle-aged man 
in camou fatigues decorated with pow-wow badges who called himself 
“Meatloaf” gave them a two-hour briefing. He had a topo map of the 
Rez and the surrounding areas stuck up on the wall of the school 
auditorium, and they sat around it in the fading light of the sun that 
streamed through the steel-reinforced windows.

“The areas that have post-its are strategic. No one except a Warrior 
goes within 20 meters of these.”

“Sixty feet,” Lee-Daniel translated for the surveyors and the 
antennamen, who were products of the American educational system and 
hence impedance-mismatched with the entire metric-speaking world.

“Sixty feet,” Meatloaf said. “You'll know you've gotten too close 
if you find yourself at the bottom of a ten-foot pit with two broken 
legs. Don't go near the strategic areas, OK?”

Elaine stood up and began to pace the map's length. She unsnapped a 
laserpointer from her gearpig bandolier and began to hit each strategic 
area in turn.

“All the high-ground, right?”

Meatloaf nodded.

“The perimeter, too, right?”

He nodded again.

Elaine gave Lee-Daniel a look, then ran the dot of her pointer over 
each of the strategic areas again. Some of the surveyors groaned and 
whispered to the antennamen and the switchgirls.

Lee-Daniel cleared his throat. “Meatloaf,” he said, “all respect, 
but well, this won't work. Our radios operate on line-of-sight. If we 
can see it, we can shoot it at half a gigabit a second -- slower if 
there are a lot of leaves and stuff in the way. If we can't see it, we 
can't shoot it. Zero bits per second. We need high-ground, we need 
perimeter, otherwise we're just wasting your time.”

Meatloaf shook his head. “Radio radiates. I can't see the cell-tower, 
but I can still reach it with my phone.”

“That's dumb radio,” Lee-Daniel said. “If we want to have a 
conversation and we're out of sight of one another, we can communicate, 
but only if we shout. That's fine for us, but it's not so good for the 
people between us, right, Mortimer?”

Mortimer, who'd been through one or two (hundred) of these demos 
before, took his cue from outside the doorway, hitting it with the 
loudhailer dialed up about half way. “Right,” he said.

“That's how dumb radio works. You had a bunch of bands that you could 
communicate in -- cellular, TV, AM, FM, cops, air-traffic, whatever -- 
and rules and licenses for each, governing how loud everyone gets to 
shout.” Taking their cues, the CogRads started to gabble all at once, 
in stripes through the ranked chairs, saying “AM AM AM” or “TV TV 
TV” or “cellular cellular cellular.”

“Smart radio -- \emph{cognitive radio} -- is much more clever. 
Instead of shouting loud enough to be heard across the entire distance, 
cognitive radios cooperate with one another. When I need to talk to 
Mortimer, I first check around to see what channels are least occupied 
and most close to me, then I send my message to the best candidate.” 
He turned to Elaine, who'd come to stand by his shoulder. “Tell 
Mortimer that it's time to come back,” he said.

Elaine turned to a switchgirl who'd positioned herself a few feet away 
and said, “Tell Mortimer it's time to come back,” she said.

The switchgirl turned, but the next person in the chain, a customer 
service rep, had his phone headset in and was having a hushed support 
call -- it was faked, just part of the script, but he gave a good 
impression of helping someone tech a network problem at a distance, 
tracking a nonexistent support-script across his HUD and prodding at 
the air with a dataglove.

“Aha,” Lee-Daniel said, “here's where it gets tricky. What if one 
of the radios between us is too busy to relay a message? We've got two 
options. We can wait -- which we'll do, if we have to, but it adds 
latency to the message -- or we can find an alternate path.”

The switchgirl -- a network engineer he'd hired himself from a 
backwater DeVry at a job-fair in Tulsa, who ran a little to fat but was 
still broad-shouldered from her time on the rowing machine she shlepped 
compulsively from gig to gig, facts that Lee-Daniel could recall with 
ease even if he couldn't remember her name -- turned back and passed 
the word onto a surveyor who was standing a little ways out of the way, 
who relayed it to Joey Riel, who was by the doorway, who stuck his head 
into the corridor.

Mortimer sauntered back into the auditorium. He put the mic to his lips 
and boomed “You want something, boss?”

Lee-Daniel clamped his hands to his ears along with the rest of the 
crew. “No need to shout,” he said to Mortimer. “Is there?” he 
said to Meatloaf.

\tb

“So, what's the critical path, Mac?” Lee-Daniel asked. “Who's 
going to run this circus between tonight and your \emph{executive 
search} coming through with an empty suit to sit in the driver's 
seat?”

“We thought you'd stay on, LD, help with a smooth transition.”

“Why would I do that?” Lee-Daniel said.

The Series A and Series B investors watched them like a tennis-match, 
silent, eyes shining.

“Why don't you two get us a couple beers, OK?” Mac said to them. 
They mooched off petulantly. “LD, I hope you'll stay on because you 
have a significant stake in this company. We don't want to dismantle 
CogRad, we want to \emph{grow} it, turn it into something big and 
important. We need your help still, to make this work. It's your show, 
we know it. We can't make the transition without you.”

He knew that Mac was blowing smoke, massaging his ego. But he was good 
at suppressing his ego for the good of his company. \emph{The} company. 
Not his company, not anymore. “What's in it for me?”

Mac made a face and leaned in close, whispering. “Look, \emph{they} 
don't want you to stay. I had to fight, hard, to keep you in, even 
during the transition. I put a lot on the line for you. \emph{I} know 
that your job can't just be filled by a warm body, \emph{I} know that 
you're the only one who can train your successor. We can make this 
company \emph{really} big -- you'll be able to retire on your share in 
18 months if we go according to plan. We're going to franchise out, 
start new busses in Latinamerica, Asia and Africa in four months, turn 
it all into a turnkey solution, something that scales. We'll raise 10 
billion on IPO if we raise a cent, you just watch. I've been through 
this, LD, and I know what a success smells like. This will be a success 
-- your success -- if you play along. If you don't, well, we could all 
end up in the shitter. Canada was the last straw for them. We either go 
on without you or we don't go on at all, do you understand?”

The Series A and Series B men returned with a couple of novelty beers 
in aerosol cans. Mac and Lee-Daniel sprayed their throats with the brew 
and swallowed, making faces. This was high style in the circles the 
Series A and Series B men traveled.

“I see,” Lee-Daniel said. “So I either walk out of here as 
interim CEO, knowing that I'm gone in a couple weeks, or I walk out of 
here fired. Is that the deal?”

“That's the deal,” the Series A man said. “And I don't see anyone 
offering anything better.”

MacDiarmid gave him a shut-up-asshole look, then spread his hands out.

“When I raised money from you, we did it over the course of several 
weeks. We talked to lawyers. They exchanged documents. I don't think 
it's reasonable for you to expect me to sign anything now without at 
least consulting a lawyer.”

“You want several weeks?” the Series A man said, with mock 
incredulity.

“Half an hour,” Lee-Daniel said. “I don't think that's too much 
to ask.”

\tb

They settled for Warrior escorts for the antennamen and the surveyors. 
The Warriors were resentful at first, but they came around.

Lee-Daniel went out with a crew that Elaine was leading, up on the 
northern border of the sovereign. She had two junior surveyors with 
her, all of them loaded with positioning gear that tied into Galileo, 
the European GPS network -- the Galileo gear cost a fortune, but they'd 
found that their American GPS kit often mysteriously stopped working 
when they were working on projects in the territorial USA. They'd 
ordered the Euro stuff from a bunch of anti-globalization activists 
who'd found that the same thing happened in any city hosting an 
economic summit. Europeans were more likely to treat infrastructure as 
sacrosanct, while the US was only too happy to monkey with GPS for 
tactical reasons.

The surveyors and the Warriors kept their distance as they set out, one 
Warrior leading and one bringing up the rear. Elaine would call for a 
break every five or ten minutes and do magic with her many devices, 
chattering into her cellphone to communicate with the other crews, make 
sure they weren't overlapping or diverging too widely.

The woods had a high canopy, which was good news. When they started 
out, they'd focused on getting above the leaf-line, since leaves badly 
scattered RF signals, but they'd ended up with networks that were only 
reachable by people who were twenty feet off the ground. They'd blown a 
fortune downlinking the relays to ground-level stations with 
omnidirectional yagi antennae.

But then Lee-Daniel had had a brainstorm -- build the network 
\emph{below} the leaf-line. Heavy canopy starved out any foliage that 
grew below the tree-tops, leaving a clear line-of-sight (modulo the 
treetrunks, which were largely RF transparent) on the forest floor. 
That pushed CogRad from a theoretical project to a real success, and 
Akwesahsne was just the sort of woods that the CogRad gear thrived in. 
Within a week, the entire Rez would be unwired at 500 megabits/second, 
enough connectivity to move whatever data they could find a use for.

The frontmost Warrior, a girl of about 16, started off treating 
Elaine's halts as a nuisance, but after the fifth one, when Elaine 
unshipped some especially tasty laser-based theodolites and a 
high-sensitivity digital altimeter, the girl's curiosity overcame her, 
and she crowded in close to watch Elaine work. She didn't say anything, 
but thereafter, it was clear that she was fascinated by Elaine and her 
masterful use of all her toys, bangles and bobs.

Elaine noticed it, of course. She was like a magnet for teenaged girls 
-- competent, beautiful, in charge. At the next stop, she handed the 
girl a can of pink spraychalk and directed her to mark the sight-lines. 
The girl almost dropped the can, but then recovered and puffed up a 
bit, marching off to lay down the hot pink lines. The Warrior at the 
rear, a man of indeterminate age who wore a camou balaclava, rolled his 
eyes, but that was OK; Lee-Daniel was figuring out a way to get him 
engaged, too.

At the next stop, a bare ridge that overlooked the woods on one side 
and the public highway on the other, Lee-Daniel tapped the other 
Warrior on his shoulder, then gestured at a travois on which Elaine's 
juniors had been hauling their satellite tester. He cocked his head, 
then bent down to take one end, and the other Warrior fell in at the 
other end. The two juniors looked relieved and hitched up their packs, 
breaking out protein bars from their belt-pouches.

And so it went. By the time they reached the next ridge, the girl 
(“Mermaid”) had introduced herself, and the man (“Cobra”) had 
done likewise, removing his balaclava to reveal a middle-aged face 
handsome but for the deep acne scars.

And so it went, for all the CogRad crews, who'd never had explicit 
training in making friends with the locals on a gig, but who had 
learned from the example set by Lee-Daniel and by the middle-managers 
who'd learned it from him.

Elaine gave Mermaid a cheap theodolite with an integrated compass, GPS 
and altimeter, and a little booklet on how to use it, and the next time 
Lee-Daniel saw her, she was leading a group of even younger girls on a 
series of surveying missions around the Indian School.

\tb

“Privacy, please,” he said.

“We're standing all the way over here,” the Series B man said, from 
the across the little table. “How much more privacy do you want?”

Lee-Daniel shook his head in exaggerated disbelief and then MacDiarmid 
led them back out to the communal area.

Once they were gone, he flipped open his phone and called Joey Riel.

“What?” he said.

“You're fired,” Lee-Daniel said. “I hate to do it, kid, but it's 
coming down from the investors. You, Elaine, all the customer service 
reps, all but the two most senior antennamen and switchgirls.”

“What the \emph{fuck}?” Joey Riel said.

“Keep it down,” Lee-Daniel said. “Just keep it easy. They're 
coming out now to do a head-count and sort out the order for the 
firings -- they're going to do it in small groups. I wanted to let you 
know, because you've only just gotten your promotion, so you might get 
severance at the old salary-level, so I thought I'd give you a little 
extra time to make some calls and line up some money before the 
roadhouse cuts off your credit. You're going to need a ride, too.”

“What?” Joey Riel said. He sounded \emph{purple}, ready to bust. He 
cursed in three languages.

“You heard me. Keep it to yourself, OK? I gotta go.” He snapped his 
phone shut, wondering how long it would take before everyone in the 
company knew: five minutes? Ten?

He called his lawyer.

\tb

It wasn't his idea to bring the investors along on the perimeter walk. 
This was a purely ceremonial event, only initiated once the real 
post-install survey had been completed and he was sure that there was 
network integrity. But networks must not only be integrated, they must 
be \emph{seen} to be etc, so they split into four crews and walked the 
perimeter.

They used ruggedized videoconferencing tablets as they went, digital 
clipboards whose screen was divided into a two-by-two grid, each square 
with the feed from one of the crews. The data went over the localnet, 
and streamed out over the uplinks to residents of any other unwired 
sovereign that wanted to welcome the newest Rez to the party.

The four parties each took a direction and hiked out to the 
most-distant corner of the Rez and then began walking 
counter-clockwise, keeping in constant communication. A little 
blinkenlight in each quadrant mapped the throughput to and from that 
host, five bars all the way and not a single frame dropped if all went 
according to plan.

The investors were with Northeast party, along with Joey Riel, Meatloaf 
and Mermaid. Not Mac, he was on the bus, where he usually spent the 
dusks and dawns, in air-conditioned gloom out of the mosquitos' range. 
But the Series A and Series B men went Northeast, while Lee-Daniel took 
the opposite corner, Southwest, with Elaine and the hard-line girl from 
the gate on the first day and Cobra, who'd taken to watching the 
sunsets with him and sharing a pint of bourbon, not saying anything.

They reached the perimeter and began to pace it off. Over the audio on 
the videoconferencing tablets, he heard the investors' labored 
breathing, the slipping of their impractical Oxfords on the slippery 
humus that carpeted the forest.

It was a nice early-fall day, with bloody streaks of sunset on the 
horizon and the crisp smell of damp and wind and sap dripping from the 
maples. Lee-Daniel loved an autumn walk in the woods, hell, who didn't, 
and so he was pretty relaxed by the time he got half-way around the 
Rez, an hour later, in the growing gloom.

It was then that bright beams of light stabbed at them from all sides. 
Behind him, he heard Cobra curse and then he was shoved aside and down 
as Cobra and the girl took up back-to-back positions with their weapons 
-- a gas-fogger for her, a hunting rifle for him -- at ready.

“Surete,” Cobra hissed. Surete du Quebec -- the Provincial cops.

He'd done the research, knew that the SQ and the Warriors hated each 
other. The Mohawk Warriors Society had been fired in a kiln bricked 
with SQ beatings, shoot-outs and gassings. But the Akwesahsne Rez had 
been at peace for almost three years! Why the hell couldn't this have 
happened \emph{tomorrow}, when they were on the road?

Lee-Daniel knelt down and dialed down the screen brightness on his 
tablet, then peered at it. His own quadrant showed his long, narrow 
face, uplit like a Jack-O-Lantern by the screen, eyesockets black and 
deep, cheeks hollow and stippled with patchy three-day beard. Two of 
the other quadrants were black -- the tablets were offline or broken. 
The final one showed the Northeast party, skinny Joey Riel holding a 
thick branch in one hand and a rock in the other, ridiculous alongside 
Meatloaf and Mermaid, who had already fitted their masks and goggles 
and drawn their sidearms, crouching back to back against each other.

The investors hove into view, whey-faced, lips skinned back from their 
teeth, eyes crazy-white.

“Get down,” Lee-Daniel said, leaning into the mic. “Head to the 
bus.”

“It's dark,” the Series A man said, jinking from foot to foot, 
making the camera sway seasick.

“The bus,” Lee-Daniel said. “Get in the bus. Get everyone to the 
bus. This isn't our fight.” He looked around for Elaine, but he 
didn't see her. Headed for the bus, that's what you did in an 
emergency. Fuck.

It was an emergency. There was an even tramping of feet ahead of him, 
behind him, to his left and right. He stood, slowly, and put his hands 
in the air.

“I'm not a combatant,” he said, loudly, but in a steady voice.

He walked toward the bus, hands still in the air. “I am not a 
combatant,” he said again. A laser dot climbed his toe, his leg, 
centered on his gut. He looked down at it.

“They will shoot you, you know,” Cobra said. “They shoot. They 
think they're playing cowboys and indians.” He sounded very calm.

“I am not a combatant,” he said again, taking another step forward. 
A second red dot joined the first, climbing his leg and resting within 
inches of the first, dancing and bobbing like a firefly. From the 
woods, someone barked in French.

“I surrender,” Lee-Daniel said.

“They don't speak English when they don't want to,” Cobra said. 
“If I were you, I'd get down and stay down.” Then he yelled 
something defiant in French. The girl behind him tittered nervously.

“Cobra's making them mad,” she said, giggling again.

Lee-Daniel turned around slowly, getting away from the harsh white 
light. Green blobs swam in his vision. He began, very gently, to sink 
to his knees, when out of the corner of his eye, he spotted Elaine and 
two of her crew, in silhouette, up in the boughs of a maple that they 
must have climbed as soon as the SQ arrived on the scene. More steps 
from the brush, the light coming closer.

Cobra called out more French, three lights on him, his rifle at his 
shoulder. Two laser-dots danced on him, and Lee-Daniel had an 
irrational urge to slap them away, like horseflies.

The young girl hit her fogger, spraying a thick, opaque cloud of gas. 
“Cover your eyes,” she said, and giggled again. Lee-Daniel pulled 
his shirt up over his face and dropped. He belly-crawled blindly, 
towards where he thought Elaine and her crew had been treed.

He knocked his head on a tree-trunk and gasped involuntarily, getting a 
lungful of the gas, which made him retch into the depths of his shirt, 
bringing on more gasps and more retching. He rolled for the clearing's 
edge, hit another tree and got to his knees, heaving like a dog. He 
still had hold of the tablet, and when he could open his eyes again, he 
looked into it, saw the investors still staring at him, wide-eyed.

“Go!” he hissed. “Jesus, get to the goddamned bus.”

“Are you all right?” they said.

“I'm fine,” he said. “Go go go!”

\tb

The CogRad drunk-ons were legendary. When you spent weeks at a time in 
the deep bush on dry reservations, lugging gear and fighting with bitch 
physics, you needed to unwind. On the off-days, it was traditional for 
a drunken riot to ensue. Lee-Daniel occasionally partook, enough to be 
friendly, but never so much that he lost control. He set a sane 
example, and the crew followed it, and so the most harm that a big 
booze-on would cause was a gang-wide neolithic hangover, swampy and hot 
and damp.

But the drunk-on that was proceeding when Lee-Daniel stumbled out of 
the dining-room was like a heavily sponsored Bosch painting. Elaine was 
alternately necking with and slapping Joey Riel; Mortimer was collapsed 
on a heap of still-steaming rum-toddy cartons; the Customer Service 
Reps were playing kick-the-can with their ringing cellphones. The 
aerostats and the advertorial screens had automatically adjusted to 
overcome the ambient noise level, and were consequently pitching their 
jingles and come-ons at megaphone levels.

The Series A and Series B men were huddled together out front of the 
roadhouse, along with MacDiarmid.

“That's some scene, huh, boss?” Lee-Daniel remarked as he stepped 
into the cool night, sucking up the fresh air and the moonlight.

“What'd you tell them?” the Series A man said.

“Tell them?” he said.

“They think they're all fired,” MacDiarmid said. “Why do they 
think that?”

“Just road-crazies. Like when they thought they all had West Nile. 
They get worked up. Egomaniacs and social retards.” He was speaking 
in the grudging half-sentences that Cobra had preferred. Talking like 
that made him feel crazy and brave and alien.

“What are you going to do about it?” the Series B man said.

“Nothing,” Lee-Daniel said.

There was a crash from the bus. One of the surveyors had beaten in the 
safety-glass skylight and dropped down inside. They watched a 
headlamp's beam jump crazily around the bus's interior, high, low, 
left, right.

“What's he doing?” the Series A man said. Lee-Daniel knew that the 
surveyor in the bus snored, that he did tricks with a butterfly knife, 
that he sent money home to his little sister in Muncie. He couldn't 
remember his names. He was no good with names. But he knew his people.

“What you do, when you get fired. Stealing office supplies.”

The surveyor crawled back out of the sunroof with a pillowcase stuffed 
with schwag, then lit out down the freeway's shoulder in the direction 
of Buffalo.

“He won't be the last,” Lee-Daniel said.

“No,” said MacDiarmid. “I don't expect he will.”

“But they're \emph{not fired},” the Series B man said. Lee-Daniel 
didn't know either of the investors' names. Fucking spear-carriers, 
fronts for unimaginable, implacable wealth, charged with returning 400 
percent over three years on a national-budget-sized fund.

“Tell them,” Lee-Daniel said. “They don't listen when they're 
like this.” They don't listen to people like you, not ever.

\tb

“Get down,” he said to Elaine. She was wedged into a crook and tied 
off with an improvised harness made out of nylon rope and carabiners 
from her vest. “We've got to get back to the bus!”

“They'll shoot us,” Elaine said.

“They can't see us,” he said. Laser sights danced in the fog. He 
heard the crack of Cobra's rifle.

“He's scared,” Elaine said. Next to her, also tied off -- where did 
Elaine keep that many carabiners? -- was a young surveyor, one they'd 
just picked up in Montana, a kid with a shaved head who had shyly asked 
him for a job after meeting Elaine at the local Army-Navy store and 
getting a lecture on which gear to buy and why. He was wrapped around 
the branch like a serpent, locked at the ankles, thighs and wrists.

“So am I,” Lee-Daniel said. “They're shooting. It's natural. Get 
him down. Push him off the branch if you have to.”

“What about him?” she said, gesturing at the branch below her. 
There was another surveyor, a forty-something lunk who didn't wash 
enough and farted too much and blamed it on everyone. He was balding 
and his comb-over hung limply at one side of his head as he hugged the 
trunk.

“Push him too,” Lee-Daniel said.

The tablet, stuck in his waistband, spoke. It was the Series B man. 
“Don't give them any more advice. You shouldn't be liable for what 
they do in this situation. Return to the bus.”

Lee-Daniel shrugged up at her, caught a whiff of gas that set his eyes 
to watering and looked back at the clearing. Cobra was lying on his 
side, face away from them. The girl was holding his hand, face covered 
by a placid mask, but he heard her sob.

“Now! Back to the bus!”

Lee-Daniel climbed the tree. He got up to the first surveyor's branch, 
Ole Stinky, and he gave the man a shove. He fell like a stone. He 
stepped on Stinky's branch, grabbed the kid by an arm and yanked, hard. 
The kid dropped, too. “Down,” he said to Elaine, and dropped, 
landing on the kid.

“Leave them,” the Series A man said. “We aren't insured --”

He helped the kid to his feet, then Stinky. In the clearing, the Surete 
had surrounded the girl. Her hands were up, glistening with blood. One 
turned towards them and shouted something in French, raising his (her? 
hard to tell with the martian armor) sidearm. Lee-Daniel froze, and 
then a red dot appeared on the SQ's leg, traveling up to his (her?) 
crotch. One of the other SQs pointed and the SQ with the gun looked 
down, then dropped his (her?) arm and leapt back.

Elaine jumped down, holding her laserpointer in her hand.

“Run!” Lee-Daniel said, shoving at the two surveyors, then taking 
off, running blindly in the moonlight, whacking into tree-trunks, 
tripping.

The bus was crowded with CogRads and he vaulted up the steps and 
slammed into the driver's seat, authenticating on the palm-reader and 
putting the bus through its warm-up/lock-down urban defense checklist. 
He was vaguely aware of more bodies coming in, then he slammed the 
door-close button as the shutters unrolled over the windows.

He hit the internal lights, stood and turned around. “Count off,” 
he shouted, over the buzz of conversation.

Automatically, MacDiarmid said, “One,” as he had at every count 
since he'd signed on and gotten the first investor's chair.

He turned to the Series A man expectantly. “Fuck that,” he said, in 
an uncharacteristic burst of blue talk. “Drive, goddamit! Now!”

Lee-Daniel pointed at him and said, “Two,” then pointed at the 
Series B man and said, “Three.” He pointed at the middle-managers 
behind him and they began to count off methodically.

The Series A man turned around and shouted, “Shut the fuck up or I'll 
throw your ass out of the fucking bus. \emph{Drive}! ”

Lee-Daniel looked at Mac. “Better drive, LD,” he said.

They were already 20 miles off from the Rez when Elaine told him that 
Stinky and the kid never got on the bus.

\tb

Three more raiders looted equipment off the bus. Lee-Daniel played a 
little game with himself, betting on which one would be the next to 
turn raccoon. He was two for three within twenty minutes.

“Stop them!” the Series A man said.

“Why?” Lee-Daniel said.

“They'll ruin us!” the Series B man said.

“They'll ruin \emph{you}.” Lee-Daniel said. “There go the 
satellite uplinks. Wow. Good retirement benefit.”

“What do you want?” the Series A man asked.

“Sell me your stake. He sells Mac his stake. The paperwork's inside, 
on the fax machine.”

“You're joking,” the Series A man said.

“That's the deal,” Lee-Daniel said. “And I don't see anyone 
offering anything better.”

There was a tinkle from the bus and Joey Riel sped away with a stack of 
optical switches, off into the night, staggering slightly from the 
burden and the booze.

“I need to call the partners,” the Series A man said.

“You have fifteen minutes. Better not take much longer,” Lee-Daniel 
said, as an expensive crash resounded from the direction whence Joey 
Riel had gone.

“Fifteen minutes?” the Series B man said, whitening.

“You want several weeks?” Lee-Daniel said. He wiped his palms on 
his thighs and kept his grin off his face. It'd be good to get back in 
the driver's seat.

\section{Afterword:}

I wrote this for Bruce Sterling's Turkey City workshop in Austin, TX. I 
was nervous and thrilled to be invited. Bruce is one of my idols -- and 
he's now a friend and colleague, and my daughter's godfather, besides. 
We'd corresponded, sat on panels together, but this, this was levelling 
up. It was a hell of a workshop, and it was also where I met Raph 
Koster, now also a good friend (as well as an astute and inspiring game 
designer and theorist).

I'd admired a play by Dewayne Hendricks to use Indian land in the USA 
to test out cognitive radio applications, on the basis that these 
sovereign territories were not under FCC jurisdiction. He'd found 
various tribal leaders who were excited by the idea. Cognitive radio 
may just be the most radical, game-changing technology on our immediate 
horizon -- if it works.

In the meantime, I couldn't shake my memories of the brutal standoff at 
Oka, in Quebec. This is a Mohawk reservation whose sacred burial ground 
the local town decided to expropriate for a golf-course. The situation 
escalated and soon there were heavily armed Mohawks and heavily armed 
cops and heavily armed soldiers all pointing automatic weapons at each 
other. I never forgot the desperate bravery of the Mohawk Warrior 
Society.
\end{document}
