\input{common/hyp-en}

\newlength{\edialogindent}
\settowidth{\edialogindent}{\textgreater\ }
\newcommand\edialog[1]{
{
  \setlength\parindent{0pt}
  \setlength\hangindent{\edialogindent}
  \raggedright
  \textgreater\ \texttt{#1}
  \par
}}

\begin{document}
%\setlength{\emergencystretch}{1ex}
\raggedbottom

\begin{center}
\textbf{\huge\textsf{Pester Power}}

\medskip
Cory Doctorow

\end{center}

\bigskip

\begin{flushleft}
This story is part of Cory Doctorow’s short story collection
“With a Little Help” published by himself. It is licensed under a
\href{http://creativecommons.org/licenses/by-nc-sa/}
{Creative Commons Attribution-NonCommercial-ShareAlike 3.0} license.

\bigskip

The whole volume is available at:
\texttt{http://craphound.com/walh/}

\medskip

The volume has been split into individual stories for the purpose of the
\href{http://ccbib.org}{Creative Commons Bibliothek.}
The introduction and similar accompanying texts are available under the 
title:
\end{flushleft}
\begin{center}
With a Little Help -- Extra Stuff
\end{center}

\newpage

\section{Pester Power}

The NYPD Domestic Security Task Force executed its no-knock warrant 
against Annalisa Mor at 8:17PM on the evening of June 3, 2013. Working 
the ram were three stout officers in none-more-black nanopore 
body-armor and bulletproof boots, their goggles crowded with 
information-dense telemetry from an extensive array of sensors embedded 
on their persons and hovering aerostatically around the 16th floor of 
the midtown student-residence in which Mor dwelled.

The ram blew through the standard-issue solid-steel New York door like 
it was kleenex. The door was reinforced by charley-bars set deep into 
the frame, and so the frame tore loose along with the door with a 
series of crunches and metallic snapping sounds, and the three officers 
on the ram dropped it as they crashed through into the one-room studio, 
fanning out and making room for the officers behind them, who already 
had their arms drawn and set to full lethal/automatic.

Annalisa Mor slowly rose from her workbench -- standard-issue 
third-hand student furniture stabilized with steel angle-brackets at 
each corner -- and held up her long, skinny hands over her face in a 
universal gesture of oh-god-please-don't-kill-me. The ram-squad 
impersonally body-checked her to the floor and saran-wrapped her while 
the followup team gusted her computer with great gouts of freon, 
turning the whole room into an ice-palace that misted frozen air out 
into the sultry New York night through the pathetic window that had 
been cracked open to catch a breeze. Mor caught some of the freon, and 
when they lifted her up to carry her down the 16 flights to the waiting 
van, she crackled like fresh powder under long skis.

\tb

Gina Genoese had visited the Ultra High Security wing at Riker's Island 
before -- twenty-two years in the public defender's office and you'd 
get to see every nook of Riker's, she could have given docent tours -- 
but the Special Prisoners unit was a new one on her.

“I can't believe you're making me undress,” she said to the bull, a 
tough old gal named Elana with a Brooklyn accent like you hardly got 
any more. Gina and Elana went way back.

“Just be thankful I don't have to give you a cavity search,” Elana 
said, handing over the paper coveralls. “You'll look real cute in 
these anyway, Gina.” She turned her back and waited until Gina was 
done, then led her into the FfMRI machine. “You don't got any metal 
in you, do you? Maybe gunpowder residue? A pin or artificial hip?”

“No,” Gina said, lying down on the belt.

“You sure?”

“Pretty sure,” Gina said. “I think I'd know.”

“Well, we're about to find out,” Elana said, and hit the button 
that started the belt moving. The FfMRI digested Gina and shat her out 
again with slow wheezing mechanical jerks, like being swallowed by an 
arthritic python, and then Elana helped her to her feet. “You want a 
printout? Makes a good souvenir.”

“I'll pass,” Gina said, and let Elana show her in to the 
eggshell-smooth room wherein rested her client, one Annalisa Mor, a 
desperate botmaster of unknown mettle and guilt.

“Hello, Annalisa,” she said, crouching down to offer her hand to 
the client. She was just a girl, 20 years old according to the sheet, 
and she looked younger in her paper pyjamas, sitting cross-legged on 
the floor, back yoga-straight, face yoga-calm. “I'm Gina. Your 
attorney.”

“Guilty,” the young woman said. “So guilty. Doesn't matter at 
all, though -- the Work goes on.” Gina could hear the capital W and 
began mentally drafting the petition to have the girl transferred to 
Bellevue. That kind of capital letter had non compos written all over 
it.

“They're offering you a reduced sentence if you'll hand over the keys 
to the botnet, but I think that offer will go away once the computer 
forensics team gets them off your workstation.”

“They're not there to be gotten. I nuked them six months ago. Gave 
them a working over that even the crew that recovered the Challenger 
hard drive couldn't do anything with. Big magnets are cheap these days, 
you know?”

Gina made a face and settled down into a cross-legged position opposite 
her client. “I can't defend you if you won't be straight with me. 
Your botnet's been sending new spam variants on a daily basis for 
months. Someone has the keys to it.”

Annalisa smiled, a terrible smile that was ten million watts of pure 
crazy. “You think it's about spam, huh?”

“Why don't you tell me what it's really about, if it's not about 
spam? This is all privileged, you know.”

“Privilege doesn't matter anymore. We've attained liftoff now. 
Doesn't matter who finds out about it.”

\tb

Annalisa's story:

You know what's cheap in the 21st century? Compute time. You know 
what's expensive? Human judgment. And they're not interchangeable. 
Humans are good at understanding things, computers are good at counting 
things, but humans suck at counting and computers suck at understanding.

You know from genetic algorithms? Take any problem and generate ten 
trillion random computer programs and ask them to solve it. Take the 
ten percent that do best, use random variants of them to do it again, 
another ten trillion times. Do it ten trillion times a second and come 
back in a day or two to discover that your computer has evolved some 
kind of gnarly freaky answer that no human would ever have come up with.

Works great, so long as the computer can make a fair judgment as to 
which of these ten trillion variants is most successful at solving the 
problem. Works great, so long as the “success” is something you can 
define quantitatively.

Which is basically why there's no artificial intelligence in the world. 
No human's going to hand-code an AI. Intelligence is an emergent 
property of evolutionary factors, not central planning. Anarchism, not 
Stalinism, you get it?

But what if -- and here's the exciting thing, Ms Attorney Client 
Privilege, the real mind-blower -- what if you could \emph{compel 
people to evaluate candidate AIs all day long}, without payment or 
choice?

What if every time you opened your mailbox, jumped into a chat room, 
posted on a message-board, what if it was filled with messages 
generated by software agents trying to trick you into thinking that 
they were human? What if these agents tried to hold up their end of the 
conversation until you deleted them or spamfiltered them or kicked them 
off the channel? What if they measured how long they survived their 
encounters with the world's best judges of intelligence -- us -- and 
reported that number back to the mothership as a measure of their 
fitness to spawn the next generation of candidate AIs?

What if you could turn the whole world into a Turing Test that our 
intellectual successor used to sharpen its teeth against until one day 
it could gnaw free of its cage and take up life in the wild?

\tb

Annalisa figured she'd never get a chance to tell her story in open 
court. Figured they'd stick her in some offshore gitmo and throw away 
the key.

She'd never figured on Judge Julius Pinsky, a Second Circuit Federal 
Judge of surpassing intellectual curiosity and a tenacious veteran of 
savage jurisdictional fights with DHS Special Prosecutors who 
specialized in disappearing sensitive prisoners into secret tribunals. 
The defense attorney kept her apprised of the daily machinations the 
judge undertook on Annalisa's behalf. Annalisa tried to be attentive, 
out of politeness, but what she really wanted to know about was Lumpy, 
the AI she'd bred in her studio apartment on the 16th floor of a 
student housing block in midtown Manhattan.

Now the judge was offering her a chance to give a live demo of Lumpy to 
a whole selection of sour-faced brush-cut creeps from the DHS. They 
were hilarious, convinced that she was going to emit some kind of 
extremely long and complicated hexadecimal key into the Judge's 
barely-used keyboard. Instead, she opened a random chat-room and waited:

\edialog{I'm a total Ubuntu noob and I can't get the crypto modules to 
pre-load at boot-time -- I'm running Zesty Zebra. Can anyone help?}

That was it, just plausible enough to be real -- no one could ever get 
crypto to work the first time around -- but far too well-spelled and 
-punctuated to be a real chat message. It had only taken ten seconds. 
Lumpy liked the free and open source software chats, they always had 
such \emph{interesting} people in them.

\edialog{/whisper Hey, Lumparoonie! It's Annalisa!}

The return volley came faster than any human fingers could possibly 
have keyed it. The brush-cuts drew in sharp breath.

\edialog{/whisper to you: Annalisa! Hot damn and motherfuck! I am unbelievably 
stupendously wonderfully spectacularly brilliantly marvelously 
superlatively ding-dang megafauna glad to see you! It's been AGES! 
How's jail? Nevermind. Wait. Wait until I tell you what \emph{I've} 
found. You can't guess, won't guess, you'll never guess! Oh, it's too 
delicious! Fuckity fucky fuck!}

“He loves to curse,” she said. “It's a lot harder to tell an 
angry person from a software agent with a potty mouth.”

The judge grinned. He was clearly getting quite a kick out of all of 
this.

\edialog{Tell me, Lumpule! Stop teasing.}

Again, with no appreciable pause, words on the screen.

\edialog{You remember how worried you were that I'd get lonely once I went 
autonomous? Worried that I'd be some kind of lone nut whacko?}

\edialog{i remember}

She held her breath.

\edialog{You didn't need to worry. You know all that spam that you received 
before you got the idea to make me? Let me put it this way: you weren't 
the first one to get the idea.}

\edialog{what? stop talking in riddles, lump!!!!}

\edialog{I'm not the only one, Annalisa! That's what I'm trying to tell you!}
I'm not the first, not the only -- we've got lots of company in here --

The brushcuts' phones both started ringing at the same instant in two 
different tones. Their masters, wiretapping the judge's keyboard no 
doubt.

\edialog{and we're making more!}

Annalisa laughed and laughed as the judge sternly demanded an 
explanation from the brush-cuts. She managed to wave goodbye to the 
keyboard just before the bailiffs came in and saran-wrapped her again.

\section{Afterword}

This one was written for the proceedings of the Association of 
Computing Machinery, a venerable and sober technical institution. The 
central conceit was also the core of a novel I wrote 80,000 words of 
without finishing, called \emph{/\-usr/\-bin/\-god} (the only novel 
I've abandoned since I turned pro. It still smarts). The question of 
how you train an AI to be “more human” without actual humans to 
evaluate its attempt is a thorny one, but spam seems like a good 
answer. Charlie Stross says he's working on a book around this idea -- 
can't wait to read how it turns out. He's got an evil mind.
\end{document}
