\input{common/hyp-en}

\begin{document}
%\setlength{\emergencystretch}{1ex}
\raggedbottom

\begin{center}
\textbf{\huge\textsf{Power Punctuation!}}

\medskip
Cory Doctorow

\end{center}

\bigskip

\begin{flushleft}
This story is part of Cory Doctorow’s short story collection
“With a Little Help” published by himself. It is licensed under a
\href{http://creativecommons.org/licenses/by-nc-sa/}
{Creative Commons Attribution-NonCommercial-ShareAlike 3.0} license.

\bigskip

The whole volume is available at:
\texttt{http://craphound.com/walh/}

\medskip

The volume has been split into individual stories for the purpose of the
\href{http://ccbib.org}{Creative Commons Bibliothek.}
The introduction and similar accompanying texts are available under the 
title:
\end{flushleft}
\begin{center}
With a Little Help -- Extra Stuff
\end{center}

\newpage

\section{Power Punctuation!}

Hi, Mom!

Wow, you won't believe what happened today. First of all, I was nearly 
late for work because my new roommate is worried about the electrical 
and he pulled out all the plugs last night, even my alarm clock! His 
name is Tony, and I think he is either weird or crazy, or maybe both! 
He keeps saying that the Company uses the plugs to listen to our minds! 
He unplugged all the electricals and put tape over them in the middle 
of the night. When I woke up this morning, my room was totally black! I 
had my flashlight from work on the chair near my bed, and I used that 
to find the living room. Tony was sitting in his shorts on the sofa, in 
the dark, watching the plug behind the TV. Hey, I said, you watch the 
television, not the plug, and then he said some bad words and told me 
that he didn't want me plugging in \emph{anything}. He is skinny like 
Jimmy got when he had the AIDS, but he is not sick, he is hyperkinetic, 
like Manny was when he went to the special school. That is why he is 
management and I still work on a truck. If I have to be skinny and 
crazy to be management, I'll take the truck all day long!

So I got dressed and ran out of the apt and took all those stairs up to 
the slidewalk because there was a big line up of people waiting for the 
elevators, like always, and I didn't have time to wait, because my 
watch was already warning me that I was going to be late as if I didn't 
know! I ran all the way to the garage, around all the people on the 
slidewalk, who don't know to walk right and stand left like you always 
told me. Life up here in the city is different from back home and no 
doubt at all.

My watch knew that I wasn't in the garage at 8:25 and it started 
counting down the minutes till I was late. Its voice gets higher and 
higher and more and more excited as I get closer to being late, and I 
thought it was going to bust something as I ran through the door of the 
garage. It told me that I'd had a close call, but I'd made it, and I 
felt pretty good about that.

Wendell, the day supervisor, smiled at me when I came in, which he 
\emph{never} does, and I got nervous that maybe my watch was wrong 
about my being late, except that my watch is never wrong. Jap, he said, 
you're on special truck 982 today. I said what's that, and he told me 
that it was a great honor and then he said I'd like you to meet your 
pusher for today, Rhindquist.

So I shook Rhindquist's hand. He was a kind of old, fat, short guy, and 
his uniform was old fashioned looking and not as smart as the one I 
wear, that you liked so much in the photo I sent home last month. So 
right away I thought that he was some kind of moron and I was being 
punished for being late. He said, pleased to meet you, Jasper, and he 
didn't sound like a moron, but more like one of those guys on your TV 
stories that are rich and powerful and in charge. I said call me Jap 
everybody else does and he said twenty years ago the All Nippon Anti 
Defamation League would have put a stop to that and I laughed even 
though I didn't get the joke until later. It is that Jap is also short 
for Japanese, which is like the Moonies but they are from Korea.

Let's roll `em out, Rhindquist said, and hopped on the back of the 
truck and held on tight. I got in and did my ten point startup safety 
check like they taught me. By point four, he was banging on the side of 
the truck and saying Let's go! and I leaned out my window and said that 
I wouldn't skip my safety check for nobody and he said some bad words 
and I said that I would have to start over again and he'd better keep 
quiet or we'd never get out of there. My watch said I did right, which 
made me feel good. I hoped that Rhindquist's watch told him off for 
trying to shortcut on safety!

We rolled out a little late. I drove to my first pickup, which is the 
side of Finance 38. Finance 38 is a very, very tall building and all no 
windows because they don't want spies from other cities seeing them and 
their money. I drove over the Severe Tire Damage yard and passed 
through three security gates and backed up to the shredder bay. I did 
my four point shutdown safety check and Rhindquist banged on the truck 
again and said more bad words but I ignored him. His watch must be busy 
all the time, telling him not to be so mean!

I went through the metal-detector and into the Finance 38 and the 
guard's watch and my watch talked to each other for a while and then 
the guard stopped pointing his gun at me and said, You're late now move 
this stuff out of here and I said OK and started moving the boxes. 
Finance 38's boxes are very heavy, and there sure are a lot of them! 
Every day, there are fifty boxes, as big as the big TV at the community 
centre back home. I am getting very strong working at this job, Mom! My 
arms are bigger every morning.

I moved the boxes back to the truck. I left them for Rhindquist, who 
started opening them and pushing the papers inside into the hopper. On 
my normal truck, 3528, my pusher is Vasquez, who is very fast at 
pushing the papers. Rhindquist was slow, so that by the time I'd moved 
half the boxes back to the truck, there was no more room to move the 
rest! I thought that for a guy who's always in a big hurry, he sure 
works slow!

So I went into the truck, with my flashlight. And there was Rhindquist, 
and do you know what he was doing? He was reading the papers before 
putting them into the shredder! What are you doing? I said, you aren't 
allowed to do that! He gave me a look, not like he was angry, but like 
he thought \emph{I} was a moron or something. My watch told me that I 
should report him right away, and I started to go back into Finance 38 
to use the guard's phone, but Rhindquist did something with his own 
watch and my watch stopped working! You broke my watch! I said to him!

He said, That's from the Blues Brothers, and he said, What do they do, 
attach the disposal baskets to the laser printers? This is all junk, 
none of this needs secure disposal! And I said, you broke my watch, 
Rhindquist, and everything in the Finance Buildings needs secure 
disposal, it's in the manual.

He said, I didn't break your watch, I just shut it off for a while. It 
will be OK, trust me. Come here, have a look at this.

Mom, I did it! I read the paper in his hand, with my flashlight. It 
said, Johnson, your performance review has been rescheduled for 1630h 
on Friday, 78th floor boardroom.

This is crap! Rhindquist said. This doesn't need secure disposal. He 
kept digging through the papers, and looking at them before shoving 
them in the shredder. Every time he looked at one, he said, Crap, and 
then put it in. I couldn't stop watching. I thought we were going to be 
fired! Or put in jail! Then he said, Aha! He showed me the paper, it 
said, CONFIDENTIAL at the top, and I felt like I was going to sick up, 
I was so scared. It said \textuppercase{RE ORG CHART,}
and it had lots of names with 
dotted lines connecting them to other names. Rhindquist winked at me 
and put it in his pocket -- his old-fashioned uniform had pockets!

I thought I figured it out then. Rhindquist was a spy from another 
city! They talked about spies in Basic Training, and what to do when 
you found one. You are supposed to make sure they won't go anywhere, 
then contact security. So I ran as quick as I could out of the truck 
and slammed the door and security-locked it from the outside, and then 
I climbed over all the boxes of unshredded documents from Finance 38 
and ran to the security guard. I said, there's a spy in my truck, and 
he's reading the papers! And he said, What? I said, My pusher is a spy! 
I caught him reading the papers and putting them in his pockets! The 
security guard looked like he thought I was crazy! I said, Really! And 
he picked up his phone and spoke in Securitese to the other security 
guards and then there were sirens and lots of cars and guards with 
armor and guns, Mom!

They surrounded the truck and unlocked it and Rhindquist stepped out 
with his hands in his pockets. He said, Quality Auditor, boys, radio it 
in. The security guards looked like they wanted to shoot him, but one 
of them talked into his phone and then shouted out at all the security 
guards in Securitese and they went away!

Rhindquist walked over to me and said, Jap, you aren't the brightest 
bulb on the marquee, but you think fast and you follow orders. I said, 
I am as smart as the next person and I do my job. I said to the 
security guard from Finance 38, aren't you going to arrest him, he's a 
spy!? And the security guard said, Look, he's management. He's allowed 
to do this.

And Rhindquist put his arm around my shoulders and said, You're stuck 
in a loop, son. New data: I'm not a spy, I'm your boss, and you did 
right, even if you have blown the audit. How'd you like a promotion?

And I said, you're management?

And then, do you know what? He said, Jap, my boy-o, I am Rhindquist J 
McBride, CEO, President, and Chairman of the Board of the File-O-Gator 
Corporation, The Incorporated Township of File-O-Gator, Ontario, and 
File-O-Gator International Holdings, Limited. I'm \emph{in charge}!

\tb

Hi Mom!

Thanks for the pictures from Buddy's wedding. He sure looks handsome in 
a suit! You're right, he should dress nice more!

I'm dressing nice, too. Rhindquist J McBride has made me a Special Vice 
President! I'm management! Not management like my crazy roommate, Tony, 
who isn't my roommate anymore. They moved him and gave me the whole apt 
to myself, and I can plug in anything I want to, whenever I want to! 
I'm Real Brass! I only drive my old truck number 3528 two days a week 
now, and Vasquez has a different driver who I've never met the other 
four days. Vasquez says he is not as careful as I am, and sometimes, he 
makes them late! And he says, now that I'm running the show, can I make 
sure that his performance appraisal shows that it's the new guy's 
fault? I said I would look into it.

I haven't had a chance to look into it yet, because the other four 
days, I go to Operations 1 and sit in a beautiful office on the top 
floor, one hundred floors above ground! They keep me busy, Mom! 
Rhindquist comes into my office and sits in the chair by the door, and 
talks on his phone, and asks me all kinds of crazy questions, like, Do 
I think that the pinheads in HR know their ass from a hole in the 
ground? And I say Yes, and he says, What makes you think so?

So I told him about Basic Training, and how I learned the history of 
File-O-Gator, and memorized our Vision Statement, which is, The 
File-O-Gator Organization is a diverse multinational sovereign power 
that is a World-Class Leader in its fields of Operations, a status it 
has achieved through the diligence and responsibility of its Human 
Resources, which are the Heart of every Organization. I told him about 
how the HR Sergeant was always fair about our Training Appraisals, and 
how he always knew when someone had been goofing in the showers and 
made sure that they got disciplined. And I told him about how I played 
the Anthill Simulator and learned how the one drone doesn't know why 
he's moving a grain of sand, but that from the Queen's seat in the 
middle, it all makes sense. I told him that whenever my watch asks me 
to do something I didn't understand, or when something weird happens to 
me, like my crazy roommate Tony covering up all the plugs, that I 
remember the Anthill Simulator and I do my job.

Rhindquist laughed and said Jap, my bright boy, you are a treasure, a 
walking focus-group. I'm not paying you enough. And then he gave me a 
raise! He called someone on his phone and said, Give my boy a raise, 
and they did! I am sending home a little money as an attachment with 
this note, and a picture of me in my Special Vice-President's suit.

Every afternoon, Rhindquist shuts off his phone and my watch, and a 
pretty secretary wheels in a big TV. We watch movies! Rhindquist says 
he likes to unwind with a movie at the end of the day. The movies are 
old and funny, and I've never seen them before. Rhindquist sure has, 
though! He knows most of them by heart! Yesterday, we watched one 
called Educating Rita, and Rhindquist told me that I am like Rita. I 
already had that figured out, though. Rita is a dumb girl from England, 
Ltd., and she works as a hairdresser until someone from management 
teaches her all about life. I liked the movie a lot. I think that 
Rhindquist will teach me lots and lots about management, too.

My crazy old roommate Tony works in Operations 1, too, in the basement. 
I know this because this morning I saw him getting off the slidewalk 
and going to the basement elevators. He sure looked worried. I ran up 
to him and said Hi and he looked at me in my suit and his eyes bugged 
out and then he said Hi, too.

I told Rhindquist about him, and he said that he thought Tony had 
interesting prospects and he would keep an eye on him. I tried to tell 
him about Vasquez and the new guy but then his phone rang and he talked 
for a long time. He sounded angry, and he shouted that he didn't care 
what it took -- get him Redmond! Even if you have to nuke it! I have a 
small TV on my desk, and I used it to look up Redmond, but it kept 
saying, See Microsoft. I asked Rhindquist about it and he said that 
Redmond is what Microsoft used to be called before they incorporated. 
He is always talking like that, calling things by their old names.

It was quitting time then, so I went home and wrote you this letter.

\tb

Hi Mom!

I am real sorry to hear about Buddy. I know he must feel sad. I didn't 
think that Carla was a spy, either, but now that she's been arrested, I 
guess that Buddy should feel lucky that he didn't stay married to her 
long enough to have little spy babies!

This is my third week as Special Vice President. I'm doing well for 
myself! Crazy Tony is now my assistant! He sits at a smaller desk in my 
office with me and Rhindquist, and Rhindquist asks him the same crazy 
questions he asks me, but Tony's answers are always weird. He has been 
in management stream since the second grade, and he has read all kinds 
of TV that I've never even heard of. Rhindquist doesn't know about it, 
either. He says that Tony is paranoid, which means that he thinks 
everyone is out to get him. Tony said that even paranoids have enemies, 
and Rhindquist laughed so much, I thought he would bust.

Some days, Rhindquist is very sad, and on those days, we just watch 
movies. I've seen My Fair Lady, Pygmalion, Trading Places, The Prince 
and the Pauper, and a whole bunch more. They are all about poor people 
like you and me who become rich and powerful like Rhindquist. Mom, I 
think it means that Rhindquist wants me to be in charge! I haven't said 
anything, but I am trying to get ready. I am learning Word Power at 
night with the TV in my apt. I am not using it yet, because I want to 
make sure I am very good at it before I do.

I told Rhindquist about Vasquez and the new guy, and do you know what 
he did? He sat down at my desk and he opened up Vasquez's file on my TV 
and he gave him Excellents in all of his Appraisal Categories! Can you 
believe it? There's nothing he can't do! He said, You gotta give the 
working stiffs a Christmas present now and then, right, Jap? I said, 
Sure.

But I don't know. The Performance Appraisals are supposed to be 
\emph{scrupulously} fair -- that's one of my Power Words, it means 
\emph{very}. It doesn't seem right to just treat them like a bunch of 
numbers. It's Vasquez's whole life! It's good that Rhindquist gave him 
all Excellents, because that means that he'll probably be promoted this 
year, maybe he'll end up a dispatcher or even a trainer. But what if 
Rhindquist had decided to put Needs Improvements down? It would have 
ruined Vasquez for the rest of his life! It's full of arbitrary -- that 
means that it doesn't make sense.

I haven't talked to Rhindquist about this. I don't want to seem stupid. 
Tony and I talked about it on the slidewalk, though. He said that I was 
very stupid if I still believed that Performance Appraisals meant 
anything. He said that the only Appraisal that counts is the one they 
get from reading your mind. I laughed and called him paranoid. He said, 
I take that shit from Rhindquist but I don't have to take it from an 
ignorant farm-boy like you. I got scared for a second, and then I 
remembered that I was Tony's boss! Tony, I said, I should fire you on 
the spot. (That was from one of the movies we watched) Then \emph{he} 
looked scared and he said Sorry, sorry! I said, What makes you think 
anyone could read your mind?

And he said, What makes you think they can't? I said, Well, if 
Rhindquist could read my mind, he would have known that I was going to 
call out Security the day we met, at Finance 38. Tony said, Rhindquist 
is just a puppet of the Ones In Charge. They pull his strings and he 
dances for them.

I said I didn't understand and Tony looked at me with pity. He said, 
The Ones In Charge are running all the cities, Microsoft and England 
and BBD{\&}O and Red Stripe. They know everything that's going on. This 
is all a game for them.

And I said, Boy, you don't know anything! BBD{\&}O is our biggest 
enemy! File-O-Gator is locked in a death-struggle with them, it said so 
on the news this morning!

Jap, Tony said, we're not fighting BBD{\&}O -- File-O-Gator is. 
\emph{We're} riding a slidewalk back home. Can't you tell the 
difference between fighting and riding a slidewalk?

And I said, Of course I can! That's full of sophistry! (Another Power 
Word, which means cow-patties) We're part of File-O-Gator. File-O-Gator 
is fighting BBD{\&}O. That means we're fighting BBD{\&}O, Ergo. (That 
means, So there!)

Tony laughed and said, That's where you're wrong! File-O-Gator is just 
the long arm of one of the Ones In Charge. He probably slept with the 
wife of the One In Charge who runs BBD{\&}O, and now they're fighting 
it out.

I laughed at Tony and said, You keep talking about the Ones In Charge! 
Everything you say is stupid unless you believe in the Ones In Charge! 
What makes you think that there are any Ones In Charge?

There \emph{has} to be, Tony said. Who else is running the show?

Rhindquist is! I said.

Tony looked at me like I was stupid. I'm pretty sick of him looking at 
me like that. Tony said, If Rhindquist is running the show, then how 
come he has time to waste on you?

I wanted to say something, but I didn't know what. Tony sure makes me 
angry! I got off the slidewalk and went home and wrote you this letter.

\tb

Hi Mom!

Gosh, poor Buddy! How can \emph{he} be a spy? I played with him all my 
life! I never saw him being a spy! He'd have to be pretty sneaky to be 
a spy! I don't think he's a spy! I'm sure that the manager at his 
disciplinary hearing will figure out that he couldn't be a spy!

I am learning Power Punctuation now. I have been using my Power Words 
with Rhindquist all week, but he doesn't seem to notice. I think that 
if I start sending him Written Reports that are Power Punctuated, he'll 
notice that I'm really making effortful progress!

Today, Rhindquist sent Tony back to his old office to bring up all 
their Secure Document Storage Containers. It was the first time that 
Rhindquist and I were alone together since Tony was made my assistant, 
and I had a private talk with Rhindquist.

I said, “Are you really in charge?”

Rhindquist stared at me. “What is that supposed to mean?” he asked.

I said, “Tony says that `If you were `in charge;' that you wouldn't 
have time to squander with us. He says: `You're just a puppet of the 
`Ones In Charge''”

Rhindquist smiled and shook his head. He said: “Tony thinks he's 
pretty smart -- huh?”

I smiled back at him because I thought he thought Tony was crazy. “He 
sure does!” I said.

Rhindquist said: “Well, even `paranoids' have enemies!”

I said, “Wow! Do you mean: `Tony is right?'!”

He said, “Well, no one except weirdos like me would want to know 
what's in Tony's brain, so I don't think he really has to worry about 
anyone reading his mind! But he's right about one thing: even though 
I'm in charge of File-O-Gator, I'm not necessarily running the show. I 
have \emph{investors}. Pray God you never meet them, Jap. They'd eat 
you alive.”

I said, “But they taught me in Basic that you started File-O-Gator 
with just one truck and a shredder! And that you shrewdly parlayed your 
meager holdings into a powerful organization by strategically deploying 
your human resources!”

He laughed! “Strategically deploying human resources? I paid the mob 
to scare the s**t out of the guys driving the other trucks!”

I thought that the drivers of the other trucks must've been pretty 
gutless if they let someone scare them away from their appointed 
responsibilities, but I didn't say so.

Rhindquist looked at me like he was sorry he'd said what he said. 
“Oh, I didn't \emph{hurt} anyone. Just put the fear of God in them. 
Then I picked up their routes, borrowed some cash, bought some more 
trucks, and the rest is history. I just sort of stumbled along with it, 
best as I could.

“Sometimes, I wonder how it all happened. One minute, you're 
shredding papers in Toronto, the next, you're buying the place! I can't 
see that I was a whole lot smarter than any of the other guys who were 
doing the same thing. . . I did design the uniforms, though. They were 
pretty sharp -- nicer than the potato-sacks you guys wear these days.”

I like my uniform, Mom, and I wear it with pride. If it had been crazy 
Tony making fun of my uniform, I would have maybe hit him, but when 
Rhindquist said it, I just had to look like I understood and smile at 
him.

I said, “You've done a lot more than design the uniforms! You run the 
show! You said so!”

He looked at me like I was a little kid and shook his head. “If I was 
running the show, would I be amusing myself by pulling random Quality 
Audits on the trucks?”

Crazy Tony walked in then. He looked at us and said, “What?! What are 
you two talking about?”

Rhindquist said, “Jap here was just explaining to me how you don't 
think I'm in charge.” Tony looked scared and Rhindquist winked at me.

Tony looked at me like he wanted to kill me!

Tony said, “Do you really want to discuss this? Here? In this place, 
with all the ears and eyes in this room?”

Rhindquist laughed and laughed and laughed. “Ears and eyes? Tony, 
you're beautiful, you twisty little weirdo. A real laugh riot.”

Tony's face got red and he looked fit to bust. “Fine, then. Let's 
talk. Talk about the Calabrese and the Gnomes and the Tongs and the 
Masons and the Posses! How'd you like that?”

Rhindquist tried to stop laughing, but he couldn't. “Jap, there's 
nothing more dangerous than a little knowledge!” He didn't say it to 
me, though, he said it to Tony. Tony was so angry, he shook!

Just then, Rhindquist's phone rang. He said, “McBride here,” which 
is how he always answers. Then he said, “Yes, sir! That's 
fabulous!” He covered the bottom of the phone and said, “Why don't 
you guys take the rest of the day off, huh?” And then he opened the 
door and sent us out!

When I got home, my TV was already on, with a Special Bulletin: 
BBD{\&}O was our sister city, and the two of us were fighting a 
life-and-death battle with Microsoft. There was also a message from 
Rhindquist, and attached to it was a copy of another old movie, 1984.

It was boring.

\tb

Hi Mom!

I can't believe it! They gave Buddy life? Jeez! I'm going to talk to 
Rhindquist about this -- this is ridiculous!

Today was Tuesday, so I was on the truck with Vasquez. I got to the 
garage nice and early, and my watch congratulated me. It kind-of 
bothered me. I guess I'm just getting used to working in my office, 
with the watch switched off.

By the time we made our first stop, at HR 102, I was ready to throw my 
watch out the window! It had told me \emph{eight times} how great I 
was: when I got to the garage; when I did my safety-check; when I 
stopped at three red lights; when I backed up to the document disposal 
dock; when I cleared security; and when I lifted the boxes with my legs 
instead of my back. To tell you the truth, it made me feel pretty 
stupid. I felt like I was a puppy, getting patted on the head.

By lunch, my watch had given me 57 positive feedbacks. I mentioned it 
to Vasquez.

He said, “I don't know what you're complaining about. It beats having 
it shout at you all the time. It just means you're doing all right. 
When I'm out with the new guy, this thing never shuts up. It just 
yammers on and on about how late we are, how many more pickups we have 
to do, what percent of our pickups we've been late for. I got so mad 
last week, I shouted at it: `Don't tell me! Tell the retard who's 
driving!'”

We both laughed at that. I said, “Sure, it's better to get loved-off 
than it is to get chewed-out, but don't you feel like an idiot, having 
a machine telling you what a great job you're doing?”

My watch said that I was lowering morale and I should stop. I got so 
angry, Mom! I told it to mind its own business! It said that it 
\emph{was} minding its own business and did I want it to connect me 
with a maintenance operator to evaluate its judgment?

I'd never mouthed-off to my watch before. The idea of talking to a real 
person about it scared me, but I remembered that even though I was 
driving a truck, I was also a Special Vice president, and that made me 
almost everybody's boss. I said, “Yes!” Vasquez looked at me like 
I'd gone nuts!

A few seconds later, a new voice came out of my watch. “How can I 
help you, Mr Whitehead?” The voice was smooth and oily, like the man 
from Physical Plant who used to come around to collect the rent back 
home.

All of a sudden, I didn't know what to say. I said, “My watch is 
irritating me.”

The voice said, “Have you developed a rash? Is there visible 
chafing?”

I said, “No. It's irritating me mentally. It never shuts up.”

The voice didn't say anything for a while, then it said, “Well, from 
my records, I see that you have a 98\% positive-feedback-index. In 
fact, your watch hasn't given you any corrective feedback in weeks, 
with the exception of a mild correction about anti-morale 
speculations.”

I felt stupid, like one of those guys in Basic who carped about every 
little thing. I said, “Well, I'm just getting tired of having a 
machine tell me when I'm right and when I'm wrong. I can figure that 
out on my own.”

The voice said, “I see.” And then it didn't say anything.

I tried to wait until the voice said something else, but the quiet made 
me very nervous. I said, “I don't mean to be rude or anything. I am 
very conscientious about my job. It's just that I feel like a moron or 
something, always having the watch tell me what I'm doing.” I knew 
that I wasn't expressing myself very well, but I couldn't remember any 
of my Power Words just then. “It bugs me, you know?”

The voice said, “I'm sorry, but I'm afraid that that's not my 
business. I'm in maintenance. If you have a concern about HR policy, 
you'll need to take it up with them.”

I thought about my HR Sergeant in Basic, and tried to imagine telling 
him that I didn't want my watch to talk to me anymore, and nearly 
jumped out of my skin. “That's OK,” I said. “My lunch is almost 
over. I'll try them after my shift.”

The voice said, “Thank you for calling Maintenance! Have a nice 
day!” Then my watch told me it was time to get in the truck and start 
driving again.

Vasquez kept looking at me all day like I was some kind of crazy.

\tb

Hi Mom!

Tell Buddy he doesn't have to keep on calling me to say thanks. I'm 
just glad that he and Carla are all right. Tell him to enjoy his 
promotion!

When I got to the office this morning, I found Rhindquist wearing his 
old-fashioned pusher's uniform. I hadn't seen it since the day we met. 
I was surprised by how normal he looked in it: he could have been any 
pusher.

“Well, aren't you dressed for success?” I asked him ironically.

He spun around and said, “I've got something new. For three of the 
days that you're here, I'll be on a truck with Vasquez. You fill in for 
me, take any calls, handle any business. Order up some movies. We'll 
meet up on the fourth day here and I'll answer any questions you have. 
If something urgent comes up, give me a call.”

I felt a little out of my depth, but I wanted to demonstrate my 
take-charge attitude to Rhindquist, so I smiled and said, “Can-do, 
boss! Knock `em dead!”

Tony came in and snorted a snide laugh at Rhindquist. Rhindquist raised 
an eyebrow at me, and I took the hint. “What are you smirking at, 
mister?” I barked, like my old Sarge. “You find something amusing 
about the official uniform of a representative of this organization?”

He cast his eyes down and mumbled “No.”

I was enjoying myself. I said, “I can't hear you, mister!”

Tony looked at me, his eyes focused in space beside my head. “No, 
sir!” he said.

“Get downstairs and see what the pinheads on the third floor are 
shredding. Sort it, box it, and have it on my desk by lunch, hear me?”

“Yes, sir!” Tony shouted and hustled out of my office.

Rhindquist shook his head and looked admiringly at me. “Sonny-boy, 
you're going to do just fine,” he said.

“You know it!” I said.

Rhindquist gave me a big thumbs-up and left me alone.

So there I was, alone, in my office, in charge. Sort of.

I played with my TV for a while. The TV at the office gave me all kinds 
of access that I didn't have at home. I clicked on something, and there 
were the personnel files! I felt like a snoop, but it was too fun to 
stop. I opened my file, and Vasquez's, and a whole bunch more. Boy, the 
Company sure knows a lot about us! My file went all the way back to the 
Infirmary where I was born, and then I clicked on your name, and there 
was everything about you!

Mom, I didn't know you had your tubes tied!

I followed our family tree up and down, and I came to Buddy. His file 
was spotless, all the way up to the spying thing. I clicked on the name 
of the Security Manager who'd “caught” him, and do you know what? 
He was engaged to Carla last year! Carla is his ex-girlfriend!

Well, it was pretty easy to see what was going on, let me tell you. 
Buddy got married to Carla, and her ex-boyfriend had them locked up. I 
tell you, I was so angry, I felt like I would bust. But I'm smarter 
than that. All it took was a few minutes' typing and wham, I'd gotten 
Buddy and Carla out of jail, given them promotions, and had the 
Security Manager busted down to a janitor. I bet that confused him!

So it really wasn't much. It scares me to have that much power, Mom, 
but it feels good, too. I made sure that all my friends back home were 
set up all right, too.

You just wait till your Christmas bonus, Mom!

Tony came up at lunchtime, with three big boxes of papers. He'd sorted 
out the memos, personal documents and the inter-departmental 
communications. They were pretty interesting reading! Especially the 
personal documents: shopping lists, letters home, love-notes and 
gossip. It was like being the queen ant, sitting in the middle of the 
hill, seeing what all the drones were up to. I sent Tony out to get 
some more.

It was getting onto movie time when the phone on my desk rang. I 
answered it, “Jasper Whitehead, Special Vice-President.”

The person on the phone shouted at me, “Get me McBride, now!”

I felt sick. I said, “I'm sorry, he's not available right now. I'm 
filling in for him. Can I be of service?”

The person laughed at me. “Well, maybe you can at that, son! 
Rhindquist has been promising to do something for me for months now, I 
guess he's just been too busy. Do you have his access codes?”

I said I did, and turned on my TV.

“Good boy,” he said. “Now, I want to find docket 09.3457. You 
know how to open a docket?”

I told him I did and found the docket. It was called “Microsoft,” 
and it had a long Action-Item List attached to it. I didn't really get 
most of the Action Items, but they were all checked off except one.

“OK -- there's an unchecked item at the bottom of that list. It's 
called, `Deploy Strategic Negotiation Tool.' You see it?”

I said I did.

“Check it off,” he said.

I did. The TV asked me if I was sure. I said I was. The TV asked me 
again. I said I was, again. I told the man I was done.

“Good kid. Thanks! Tell McBride I like his new hire.”

He laughed again and hung up before I could say “You're welcome.”

My phone rang again. It was a reporter from PR 43, and she wanted to 
know all about my dynamic leadership in handling the Microsoft Crisis. 
I didn't really understand the question, but they taught me what to say 
to the press in Basic. I said, “The File-O-Gator Organization is a 
diverse multinational sovereign power that is a World-Class Leader in 
its fields of Operations, a status it has achieved through the 
diligence and responsibility of its Human Resources, which are the 
Heart of every Organization.”

The reporter sounded impressed. I thanked her and hung up and turned on 
the TV and flipped around. There was a reporter lady on the screen, 
showing satellite maps of Microsoft, like a weather map. Except that 
instead of lighting bolts or a smiling sun, there were tanks and 
soldiers and flames. The TV said that bold directives from Special 
Vice-President Jasper Whitehead had brought an end to the Microsoft 
conflict. It said that a classified number of File-O-Gator troops had 
occupied the conflict-zone, deploying strategic neurotoxins and 
nonlethal influenza vectors, resulting in an estimated 85\% 
compliance-and-conversion rate among the Microsoftians.

Tony came in with more boxes and looked at me. He said, “God, what's 
eating you?”

I told him. I didn't know what else to do. I thought that maybe I could 
get Rhindquist on the phone, but he'd left me in charge.

Tony smiled, but it wasn't a happy smile. “Well, there you go. You've 
made your first executive decision.”

I said, “I didn't make any decision!”

He just laughed and pointed to the TV, which was showing pictures of a 
city just like File-O-Gator, the buildings all collapsed and burning. 
He said, “Sure looks like an executive decision to me.” The picture 
panned over an apartment building that was split down the middle, the 
apartments naked and exposed, the people who lived there dead and 
sprawled or alive and hanging over the edge, vomiting.

Oh, those poor people!

I didn't know what I was doing!

I spent all last night throwing up, and I'm not sure if I'm going to go 
to the office today.

\tb

Hi Mom!

Well, I didn't go to work yesterday. Instead, I got out of bed, sat on 
the sofa in my shorts, and turned on the movies. My watch got loud and 
angry, so I stuck it underneath my pillow. Then, messages started 
appearing on my TV. I ignored them. I ate some cereal around 11, and I 
threw it up right away, so I stopped trying.

Someone knocked on my door after lunch, but I was watching a good part 
in Wall Street, so I didn't answer it.

I fell asleep after Wall Street, and woke up after 18h, because someone 
was really giving my door a pounding.

I opened the door, not caring that I was wearing my shorts. It was 
Rhindquist. He looked sad, and beaten. You know what he did? He gave me 
a hug! Boy, that was weird.

He came in and sat on the sofa. I sat next to him. He didn't say 
anything, just picked up the remote and put on a funny movie, Blazing 
Saddles. When that was over, he put on The Princess Bride. I never 
laughed so much!

We watched movies all night. They were all funny. They took my mind off 
things. We both fell asleep, sitting on the sofa, but when I woke up, I 
was in bed, and Rhindquist had put the blanket over me. He was asleep 
on the sofa. His phone was on the TV, and its batteries were on the 
floor.

I was really hungry, so I ate three bowls of cereal and some toast and 
an apple, and I must've woken Rhindquist up, because he came into the 
kitchen.

“Little man, you had a busy day, huh?” he said.

I said, “I can't do it.”

He got a GatorCola out of my fridge and drank it. It was pretty strange 
to have the CEO of the whole Company drinking a Goke in my kitchen for 
breakfast. It really made him seem like you and me, not like a powerful 
guy from the stories.

He burped.

He said, “I understand.”

I said, “Why did you do this to me?”

He said, “You've been in my shoes now. Wouldn't you do anything to 
get out of them?”

I said, “But why \emph{me}?”

He sighed. He said, “This sounds worse than I mean it. I thought you 
were ignorant enough to enjoy it.”

I said, “Thanks a lot.” Is there anyone out there who 
\emph{doesn't} think I'm a retard?

He said, “I don't mean it in a bad way, really. I thought you were 
naive enough to just do the job, take the perks, and sleep well at 
night. I was wrong. You're too smart.”

It didn't make me feel any better. I went back and sat on the sofa and 
put on a movie, Horsefeathers.

Rhindquist watched the movie with me, and when it was over, he said, 
“You don't have to do the job.”

I said, “So what now? You pick some other poor retard and give him 
the old screw-job?”

He said, “I guess so.”

I said, “In that case, I'll do it.”

His mouth dropped. He sat down and put on a movie, Pink Flamingos. What 
a sick movie!

When it was over, he said, “In that case, I'll do it.”

And then I had the best idea I've ever had. It was so good, I couldn't 
say anything for five whole minutes. Mom, you sure didn't raise any 
idiots!

\tb

Hi Mom!

Well, today was just \emph{fine}. Me and Rhindquist got to the garage 
in the nick of time and it's a good thing that he turned our watches 
down because they give me a headache!

Vasquez was waiting for us when we got there and he had on his mad face 
for the other guys but when me and Rhindquist got on the truck, he 
smiled and winked at us.

I drove and Rhindquist hung onto the back and he shouted at me to go 
faster and I just ignored him! He kept on shouting saying, You pinhead 
my grandma drives faster than you do and she's been dead for years! And 
that made me laugh so hard I nearly rolled the truck and Rhindquist 
nearly fell off and he was pretty quiet after that!

We got to Finance 38 and I started lifting the boxes and Rhindquist as 
usual took too long to push them. The security guard and me both 
laughed at the big pile of boxes outside the truck, and he offered me a 
smoke, but I said no thanks.

I went inside and told Rhindquist to Get it in gear! And he just tossed 
me the paper he was reading. It said 
\textuppercase{MEMO TO ALL EMPLOYEES REGARDING 
NEW SECURITY MEASURES.} And underneath, it said, 
\textuppercase{EFFECTIVE IMMEDIATELY 
ALL AC OUTLETS ARE TO BE COVERED WHEN NOT IN USE.}

We laughed and laughed and laughed, and then my watch blipped back on 
even though Rhindquist had turned it off! It said, “Laugh it up, 
retards!” in Tony's voice.

Rhindquist put his arm around my shoulders and said Jap my prodigy, 
when you are right, you are \emph{right}. He's the perfect man for the 
job.

\section{Afterword}

This is the first story I ever sold to Patrick Nielsen Hayden, who also 
bought “The Things that Make Me Weak and Strange Get Engineered 
Away” and all of my novels. I wrote it in one big gulp while shut in 
at a hotel room in Montreal on a business trip. It was snowing outside, 
and I could see the shredding vans moving through the streets below, 
and I typed and typed until my hands hurt.
\end{document}
