\input{common/hyp-en}

\begin{document}
%\setlength{\emergencystretch}{1ex}
\raggedbottom

\begin{center}
\textbf{\huge\textsf{Visit the Sins}}

\medskip
Cory Doctorow

\end{center}

\bigskip

\begin{flushleft}
This story is part of Cory Doctorow’s short story collection
“With a Little Help” published by himself. It is licensed under a
\href{http://creativecommons.org/licenses/by-nc-sa/}
{Creative Commons Attribution-NonCommercial-ShareAlike 3.0} license.

\bigskip

The whole volume is available at:
\texttt{http://craphound.com/walh/}

\medskip

The volume has been split into individual stories for the purpose of the
\href{http://ccbib.org}{Creative Commons Bibliothek.}
The introduction and similar accompanying texts are available under the 
title:
\end{flushleft}
\begin{center}
With a Little Help -- Extra Stuff
\end{center}

\newpage

\section{Visit the Sins}

Sean had a way of getting his way -- a way of delivering argument that 
implied that everyone in earshot was savvy and bold, and that the 
diatribe-du-jour was directed at the Enemies of Art ranged without. His 
thesis advisor bought it every time. Sean turned in his due-diligence, 
a bunch of theses written in the last century: collected memoirs of the 
survivors of electroshock, lobotomies, thalidomide. His advisor signed 
off and within twenty-four hours, he was debarking in Orlando and 
renting a car to take him to the Home.

He didn't tell his father. He'd have to, eventually, before he could 
finish the thesis. But for now, it was just him and Grampa, 
head-to-head.

\tb

Grampa was switched off when Sean found him on the ward, which throbbed 
with a coleslaw of laser-light and videogames and fuck-pix and 
explosions and car-wrecks and fractals and atrocities.

Sean remembered visits before the old man was committed, he and his 
dutiful father visiting the impeccable apartment in the slate house in 
Kingston, Ontario. Grampa made tea and conversation, both perfectly 
executed and without soul. It drove Sean's father bugfuck, and he'd 
inevitably have a displaced tantrum at Sean in the car on the way home. 
The first time Grampa had switched on in Sean's presence -- it was when 
Sean was trying out a prototype of Enemies of Art against his father's 
own As All Right-Thinking People Know -- it had scared Sean stupid.

Grampa had been in maintenance mode, running through a series of 
isometric stretching exercises in one corner while Sean and his father 
had it out. Then, suddenly, Grampa was between them, arguing both sides 
with machinegun passion and lucidity, running an intellect so furious 
it appeared to be steam-driven. Sean's tongue died in his mouth. He was 
made wordless by this vibrant, violent intellect that hid inside 
Grampa. Grampa and his father had traded extemporaneous barbs until 
Grampa abruptly switched back off during one of Sean's father's 
rebuttals, conceding the point in an unconvincing, mechanical tone. 
Sean's father stalked out of the house and roared out of the driveway 
then, moving with such speed that if Sean hadn't been right on his 
heels, he wouldn't have been able to get in the car before his father 
took off.

And now, here was Grampa in maintenance mode. He was sitting at a 
table, flexing his muscles one-at-a-time from top to bottom. It was an 
anti-pressure-sore routine. Sean guessed that it was after-market, 
something the Home made available for low functioning patients like 
Grampa.

Sean sat down opposite him. Grampa smiled and nodded politely. Sean 
swallowed his gorge. The ones who had not had the surgery had been 
scattered, unable to focus. Then they had the operation, and suddenly 
it wasn't a problem anymore. Whenever their attention dropped below a 
certain threshold, they just switched off, until the world regained 
some excitement. It had been a miracle, until the kids stopped making 
the effort to keep their attention above the threshold, and started to 
slip away into oblivion.

“Hello, Grampa,” Sean said.

Grampa stared at him from dark eyes set in deep, wrinkled nests. Behind 
them, Sean could almost see the subroutines churning. “Sean,” 
Grampa said. Woodenly, he stood and came around the table, and gave 
Sean a precise hug and cheek-kiss. Sean didn't bother returning either.

He put the recorder on the table between them and switched it on.

\tb

Grampa was a moderately wealthy man. He'd achieved much of that wealth 
prior to his retirement, working as a machinist on really delicate, 
tricky stuff. The family assumed that he did this work switched off, 
letting the subroutines run the stultifying repetitions, but in his 
prelim research, Sean had talked to one of Grampa's co-workers, who 
said that Grampa had stayed switched on more often than not. Grampa had 
acquired the rest of the wealth shortly before Sean's father had sent 
him south, to the Home. The years-old class action suit brought by the 
guilty, horrified families of accidental zombies had finally ended with 
a settlement, and all the Survivors became instant millionaires in 
trust.

For all the good it did them.

“How are you?” Grampa asked, placidly.

“I'm working on my thesis, Grampa. I'm here to interview you -- I'll 
be around for the next couple weeks.”

“That's nice,” Grampa said. “How's your father?”

“He's fine. I didn't tell him I was coming down, though. You're a 
touchy subject for him.”

Grampa settled back into his chair. Sean was distantly aware of other 
Survivors on the ward, gabbling and twitching at videogames and smoking 
all at once. They were high-functioning -- they could be switched on 
with simple stim; Grampa only switched on for important occasions.

Sean said, “Dad wishes you'd die.”

That did it. It was easy to tell when Grampa was switched on; the 
rhythmic, methodical maintenance twitching was replaced with a 
restless, all-over fidget; and his eyes darted around the room. “Is 
he in some kind of financial trouble? He doesn't need to wait for a 
bequest -- I'll write to the trustees right now.”

Sean restrained himself from saying hello again, now that Grampa was 
switched on. He kept himself focused on the task of keeping Grampa 
switched on. “He wishes you'd die because he hates you and he hates 
himself for it. When you die, he can stop hating you and start mourning 
you. He knows it wasn't your fault. That's why I'm here. I want to 
collect your stories and make some sense out of them, before you 
die.” Sean took a deep breath. “Will you stay switched on?”

Grampa looked uncomfortable. “Your grandmother used to ask me that. 
I'd promise her I'd do it, every time, but then. . . It's not 
voluntary, Sean. It's reflex.”

“It's a learned reflex, Grampa. It's not breathing. You didn't ask to 
have the surgery, but you learned the reflex all on your own. You 
\emph{allow} your attention to drop below the threshold, you 
\emph{allow} the chip to switch you off. Some people do it less,” he 
jerked his head at the other old men and women, playing their twitch 
games and shouting argument at each other. “Some don't do it at 
all.”

“Bullshit!” Grampa said, leaning forward and planting his hands on 
his knees -- aggro Type-A body-language that Sean often found himself 
assuming. “Urban legend, kid. Everyone learned it. Once you had the 
surgery, you couldn't help it. You know what I'm talking about, or you 
wouldn't be here. Your father, too -- if he was ever honest enough to 
admit it. You've both got it as bad as me, but no one ever tried to 
\emph{cure} you.”

“I don't have it,” Sean said. “I just got off a three-hour 
plane-ride, and I was able to just look out the window the whole way. 
It didn't bother me. That's not coping mechanism, either -- I never 
even \emph{wanted} to watch the seat-back vid or chat up my 
neighbor.” It wasn't true, actually. He had fidgeted like crazy, 
splitting the screen-in-screen on the seat-back into sixteen quads and 
watching as many stations as he could. He'd tried to assemble his 
thoughts on his recorder, but he'd been too wound up. Eventually, 
somewhere over Georgia, he'd surrendered to the screen and to counting 
powers of two.

Grampa pierced him with his stare. “If your ego demands that you 
believe that, then go ahead.”

Sean restrained himself from squirming. He focused himself on directing 
the discussion. “What do you like best about the Home?”

Grampa considered the question for so long that Sean was afraid he'd 
switched off. “No one makes me feel guilty for switching off. No one 
tells me that I'm weak. Except your father, of course.”

“Dad's been here?” Sean said, shocked. “When?”

“Your father visits every month. He shouts at me until I switch on, 
then he leaves. He does it because the doctor told him that if I didn't 
switch on more often that they'd move me to the zero-function ward. 
Sounds fine to me, and I tell him so, but he's never thought much of 
his brain-damaged old man.”

“Where do you go when you're switched off?” Sean asked. It was a 
question that was supposed to come later in the interview, maybe on day 
two, but he was rattled.

“I don't know. Away.”

“Is it like sleep?” Sean said, forgetting the rule that you never 
ask the subject a simple yes/no question. His heart thudded in his 
chest, like he was giving the first interview of his life.

“No.”

“How is it different from sleep?” Sean asked.

“I usually switch on for sleep -- my subconscious is pretty good at 
entertaining me, actually. When I switch off, I just \ldots{} go away. I 
remember it later, like it was a book that got read directly into my 
brain, but I'm \emph{not there}. It's fucking great. You'd love it, 
Sean. You should get the surgery. I hear that there's a lot of 
black-market clinics where you can get it done: South-East Asia. The 
sex-trade, you know.”

Sean struggled to keep the discussion on-track. Grampa was often 
hostile when he was switched on, and his father always rose to the 
bait. Sean wasn't going to. “How do you know that you're not there? 
Maybe you're there the whole time, bored stupid, screaming in 
frustration, and you forget it all as soon as you switch on?”

Grampa raised an eyebrow at him. “Of course I am! But that's not the 
\emph{me} that's important -- \emph{I'm} the one that counts. And I get 
to fast-forward past all the slow parts. Which this is turning into, 
I'm afraid.”

Grampa's eye's stopped seeking out the ward's corners, and he slipped 
back into maintenance mode. The noise and lights of the ward closed in 
around Sean. He scooped up his recorder. “Thanks, Grampa,” he said, 
woodenly. “I'll see you tomorrow.”

“Bye, Sean,” Grampa said, and came around the table for another hug 
and kiss.

\tb

Sean checked into the first motel he found, the Lamplighter Inn, on a 
dreary strip populated with disused water-parks and crumbling plazas. 
He lay down on the bed, fed the Magic Fingers, and played back the 
recording.

It was junk. The noise of the ward masked nine words in ten, and what 
words made it through were empty, devoid of any kind of emotional 
freight. He tried to transcribe it longhand, filling in the blanks from 
memory, but couldn't keep his mind on it.

He took off his sweaty, wrinkled T-shirt and slacks, dumped out his 
suitcase on the chipped, cigarette-burned table, and found his bathing 
suit.

There was one other guest by the pool, an old, old woman in a one-piece 
with a skirt, wearing a sunhat tilted to shade her from the last of the 
pounding Florida sun. Sean gave her a perfunctory nod and jumped in.

The water was piss-warm, thickly chlorinated. It felt like swimming in 
pungent sweat. Sean managed one lap and then crawled out and sat in a 
sway-backed deck-chair.

“I wouldn't go swimming in that if I were you,” the old woman said, 
in a husky, nicotine-stained voice. She clattered a grin at him through 
her dentures. She was the color and texture of rawhide, not so much 
tanned as \emph{baked}.

“Now you tell me,” Sean said, squinting at her under his hand.

“Old Ross doesn't like dealing with the pool, so he just keeps on 
shoveling in the chlorine. Don't be surprised if you're blonde in the 
morning. My name's Adele. You here for long?”

“A couple weeks, at least,” Sean said.

Adele smiled and nodded. “That's good. That's fine. A good stretch of 
time to see the Parks. Don't miss Universal, either -- I think it's 
better than Disney. Most people don't bother with it, but for my money, 
it's better.”

“I don't think I'll get a chance to visit either,” Sean said. 
“I've got a lot of work to do down here.” He waited for her to ask 
him what kind of work, and mentally rehearsed the high-concept speech 
that he'd given a thousand times while working on the thesis proposal.

“What a shame,” she said. “Where did you come down from?”

“Toronto,” he said.

“Lord, not another snowbird!” she said, good-naturedly. “Seems 
like half of Canada's down here! They come here to get away from the 
winter, then they complain about the heat! What do they expect, that's 
what I want to know! Was your flight good?”

“It was fine,” Sean said, bemusedly. “A little dull, but fine.”

“So, you're here for a few weeks,” Adele said.

“Yes. Working,” Sean said.

“Nice work if you can get it!” Adele said, and clattered her 
dentures again. “I moved here, oh, five years ago. To be near my boy. 
In the hospital. I used to work, but I'm retired. Used to work at a 
dairy -- answering the phones! You tell people you used to work in a 
dairy, they think you were milking the cows! Old Ross, he gives me an 
annual rate for my room. It's better than living in one of those gated 
places! Lord! How much shuffleboard can a body stand?”

“Your son is sick?” Sean said.

“Not sick, no,” Adele said. “You wouldn't believe the roaches you 
get down here! Old Ross fumigates regular, but Florida roaches don't 
seem to care. I've lived in New York, and I've seen some pretty big 
roaches in my day, but not like these. Like cats! My boy, Ethan, he'd 
clean and clean our apartment in New York, quiet as you please, a good 
boy. Then he'd see a roach and whim-wham, he'd be talking, joking, 
skipping and running. Old Ross says there's nothing he can do -- he 
says, `Adele, this is \emph{Florida}, and the roaches were here long 
before us, and they'll be here long after, and nothing we do is going 
to keep them away.' That's all fine and good, but let me tell you, I've 
never seen a roach in the Home when I was visiting Ethan. \emph{They} 
know how to keep them out. Maybe it's all the shouting. Lord, but they 
do shout!”

A small lightbulb blinked in Sean's mind. “Is Ethan very 
high-functioning?” he asked, carefully.

Adele glanced sidelong at him and said, “The doctor says no. But I 
think he is. He's always walking around when I'm there, doing push-ups 
and sit-ups. He's not a young man, Ethan -- sixty this year! When his 
father was that age, he didn't do any push-ups, no sir! But the doctor, 
he says that Ethan's at zero function. Doctors! What do they know?”

“How old was Ethan when he had the surgery?” Sean asked.

“Just seven,” Adele said, without changing her light tone, but Sean 
saw knives of guilt in her eyes. “He was going to be held back in the 
first grade, or sent to a special school. They sent a doctor around to 
explain it. Ethan was smart as a whip, everyone knew that, but he just 
couldn't \emph{concentrate}. It made him miserable, and he'd pitch 
these hissyfits all the time. It didn't matter where he was: the 
classroom, home, out on the street -- in church! He'd scream and shout 
and kick and bite, you've never seen anything like it. The doctors, 
they told us that he'd just keep on getting worse unless we did 
something about it.

“It seemed like a miracle. In my day, they'd just drug you up.”

Sean knew the names of the old drugs: Ritalin, Cylert, Dexedrine. 
Anything that would keep you still and numb. Then came the surgery.

Adele brightened. “You should really try to at least visit Universal 
for an afternoon, you know. It's lovely.”

“They're going to move my grandfather to the zero-function ward, I 
think. If he doesn't spend more time switched on, they will.” Sean 
said. “I want to get his story before they do it.” And if not his 
stories, the \emph{reasons} -- reasons for who Sean was, who his father 
was.

“What a nice grandson you are! You know, it seems like no one cares 
about their grandparents anymore. Old Ross's grandchildren haven't 
visited once in the five years I've been here.”

\tb

Sean gave Adele a ride the next day. She wore the sunhat and a 
lightweight cotton dress and sandals, and looked frail and quaint.

Sean thought Adele would get off at a different floor, to visit Ethan, 
but she walked with him across Grampa's ward.

Grampa was sitting just where he had been the day before. His chin was 
shaved blue, and he was impeccable. He was methodically slicing and 
eating a hamburger.

“Grampa,” Sean said.

“Hello, Sean,” Grampa said. He laid his knife and fork in a precise 
X on his plate and pushed it aside.

“This is Adele. Her son is in the zero-function ward. She wanted to 
meet you. Adele, this is my grandfather, Brice Devick.”

“Pleased to meet you, Adele,” Grampa said, and shook her hand.

“Likewise,” she said. “Do you know my Ethan? I'm worried that he 
doesn't seem to have any friends here.”

“I haven't met him,” Grampa said.

“Well, would you do an old lady a big favor? Go and visit him. Your 
grandson tells me you're smart -- Ethan is as smart as a whip. You two 
should have lots to talk about.”

“I will,” Grampa said.

“I'm sure you two will get on very well. It was a pleasure to meet 
you. Excuse me, I'm sure Ethan's wondering where I am.”

Sean waited until she was out of earshot, then said, “Her son's a 
fucking vegetable. You're about eighty percent of the way there. You're 
spending so much time switched off, you might as well be dead.”

“What do you know about it?” Grampa said, fidgeting.

“I know plenty,” Sean said. “Plenty! You spent less than fifteen 
percent of the time switched off until you hit college. Then you 
switched off for months at a time. You used it for a study aid! I 
pulled your logfiles, when I was at Dad's -- he's had them ever since 
you were declared \emph{non compos}. You're a junkie, Grampa. You don't 
have the willpower to kick your habit, and it makes my Dad nuts. I 
never knew you, so it just makes me curious. Let's talk about the first 
time you remember switching off.”

Grampa snorted. “That's a \emph{stupid} question. You \emph{don't} 
remember switching off -- that's the whole point.”

Sean rolled his eyes. “You know what I mean. You may not remember 
switching off, but you'll remember switching on. Switching on 
\emph{has} to be memorable, doesn't it? Isn't that the whole point?”

“Fine. I switched on for about 20 minutes in a movie that I snuck out 
of school to see when I was twelve. It was in French, and it had made a 
lot of noise because it had a sex scene with a live pig. I saw that 
scene, and two others -- another sex scene and a scene where this woman 
cuts the pig's throat. I loved it. All my friends had done the same 
thing, but by the time the good parts had come around, they were too 
bored to enjoy them. I just caught the highlight reel.”

“How long until you next switched off?”

“I don't know. A while.”

“It was two days. I have the logfile, remember, Grampa? Don't jerk my 
chain. You switched off during Friday dinner. Did your parents 
notice?”

“Of course they noticed! They loved it! For once, I wasn't kicking 
the table-leg or arguing with my sisters or stuffing sprouts in my 
pocket. I cleaned my plate, then sat and waited until everyone else was 
done, then I did the dishes.”

“How'd you like it?”

“I loved it! I hated family dinners! I just got the highlight reel 
again -- dessert! I remember that fucking bowl of pudding like I was 
eating it right now. My mother couldn't cook for shit, but she sure 
opened a mean package of Jello Pudding.”

Sean found his mood matching Grampa's, aggressive and edgy. “How did 
you and Grandma end up getting married? I can't imagine that she was 
hot for a zombie like you.”

“Oh, but she \emph{was}, Sean, she \emph{was}!” Grampa waggled his 
eyebrows lasciviously. “Your Grandma didn't like people much. She 
knew she had to get married, her folks expected no less, but she mostly 
wanted to be off on her own, doing her own thing. I'd come home, switch 
off, clean the place, do any chores she had for me, then go to bed. She 
loved to have sex with me switched off -- it got so that if I 
accidentally switched on while we were doing it, I'd pretend I was 
still off, until she was done. It was the perfect arrangement.”

“But she divorced your sorry ass after ten years,” Sean said.

“You got a girlfriend, Sean?” Grampa said.

“No,” Sean said.

“You \emph{ever} had a girlfriend?”

“Yes,” Sean said, feeling slightly smug. Never ask a yes/no question.

“Why'd she leave you?” Grampa asked, his eyes sharp as razors.

“What makes you think \emph{she} left \emph{me}?” Sean asked.

“Did she?” Grampa fired back.

“Yes,” Sean said, as calmly as he could manage.

“And why did that happen?”

“We were growing in different directions,” Sean said, the words 
sounding prim even to him.

Grampa barked and slapped his palm on the table. The old men and women 
in the ward swiveled their heads to stare, momentarily distracted, then 
went back to arguing.

“You're full of shit, kid. What's that supposed to mean?”

“I was working on my thesis proposal. Lara was working on hers. 
Neither of us had time for a relationship. It was amicable.”

Lara had caught him watching television over her shoulder while she was 
delivering one of her dreaded Relationship Briefings, and had laid into 
him a little too hard. He'd come back at her with everything he had, an 
extended rant that ranged from her lame-ass thesis -- the cultural 
impact of some obscure TV show from before they'd been born -- to her 
backbiting, over-educated circle of friends. He'd moved onto her 
relationship with her mother; her insufferable whining about a suicidal 
uncle she'd been close to; and her pretentious way of sprinkling her 
speech with stupid pseudo-intellectual buzzwords. He crossed the line 
again and again and she kicked him out on his ass.

“Dad says that you never switched on during the divorce.”

“Your Dad has nothing to complain about. He got enough pity lavished 
on him to kill ten men. It was all your grandmother's family could do 
not to devour him whole.”

“But you stayed switched off,” Sean said.

“In the court, I was switched off. Ever been in a court, Sean?”

“You stayed switched off.”

“In the court-room.”

“And before, during the separation?”

“Same thing,” Grampa said.

“And after, during visitations?”

“Not then,” Grampa said, loudly. “Not during visitations.”

“I've got the logfiles, Grampa,” Sean said.

“What the hell do a twelve-year-old and a grown man have to talk 
about? I kept him fed. I took him out to the carny and to kiddee 
movies. I drove him to hockey.”

“You \emph{switched off}, Grampa,” Sean said. “The \emph{you that 
counted} wasn't there.”

“Sophistry,” Grampa said. “Bullshit. I remember all of it. I was 
there. Not many other parents were, let me tell you. Usually, it was 
just me and a few others in the stands, or kids running around loose 
like animals at the carny. Your father has \emph{nothing} to complain 
about.”

“Why, aren't you two looking excited?” Adele said, hobbling 
alongside of the table. She was leaning on Ethan, a vigorous old man 
with sinewy arms and dead eyes. His face was unlined, free from smile 
lines and frowning creases.

“Hi, Adele,” Sean said, trying to keep the exasperation out of his 
voice.

“Ethan, this is Sean and his grandfather, Brice.”

Ethan extended his hand and Sean shook it. “Very nice to meet you,” 
Ethan said. His hand was dry and papery, his eyes vacant. Sean shook 
it, and a frisson of shameful disgust sizzled up his abdomen. He had a 
sudden vision of Ethan's brain, desiccated in his skull, the gleaming 
edges of the chip poking free. He surreptitiously wiped his hand on his 
pants as Ethan turned to Grampa and shook his hand. “Very nice to 
meet you.”

“Do you mind if we sit down?” Adele said. “I'm afraid that I'm a 
little pooped. All those stairs!”

Sean offered his chair and went off to the lounge with Ethan to get two 
more. When they got back, Adele had her hand on Grampa's forearm. “ 
-- I worked in a dairy, answering the phones! You tell people you used 
to work in a dairy, they think you were milking the cows!” Adele 
laughed and Grampa shot Sean a hostile look.

Sean said, “Grampa was a machinist before he retired. You really 
liked doing that, huh, Grampa?”

Grampa nodded perfunctorily.

“I mean, the logfiles show that you almost never switched off at 
work. Must've been pretty engrossing. You should give workshops here. I 
bet it'd be good therapy.” Sean knew he was baiting the old man, but 
he couldn't stop himself.

“Your father's arriving tomorrow,” Grampa said. “He called last 
night. I didn't tell him you were here, I thought it would be a nice 
surprise.”

Adele clapped her hands. “Well isn't that \emph{nice}! Three 
generations, all together. Sean, you'll have to introduce Ethan and \erratum{I}{me} 
to your father. Ethan never had children, isn't that right?”

Ethan said, “Yes.”

“Always the bachelor, my boy. But it wasn't for lack of opportunity. 
You had to beat them off with a stick, didn't you, son?”

Ethan said, “Yes.”

“I always hoped for a grandchild to hold, but you have to let your 
children live their own lives, isn't that right, Brice?”

“Yes,” Grampa said, with a kind of horrified fascination.

“Ethan was always too busy for romance.”

“Yes,” Ethan said.

“Working and working and working for that transcription service. You 
must have typed a million words. Did you ever count them, Ethan?”

“Yes,” Ethan said. “I typed roughly fifteen million words.”

“Nowadays, of course, no one types. It's all talking to computers 
now. When I was a girl, they all said that you'd always have a job if 
you just learned to type. Times sure change, don't they?”

“Yes,” Ethan and Grampa said together. Grampa startled like he'd 
been shocked.

“Dad's coming tomorrow?” Sean said.

Grampa said, “Yes. He's catching the 6 AM. He'll be here by 10.”

“Isn't that \emph{nice},” Adele said.

\tb

They left Grampa and Ethan sitting at the table together. Sean looked 
back over his shoulder before they got on the elevator, and Grampa was 
still switched on, staring hard at him.

“You must be \emph{excited} about seeing your father again,” Adele 
said to him when they were sitting around the pool.

Sean was getting the hang of talking to Adele. “Ethan and my 
grandfather seem to be hitting it off.”

“Oh, I certainly \emph{hope} so! Ethan could use some friends at that 
place.”

Sean pictured the two of them, seated across from each other at the 
ward table, running maintenance routines at each other, saying, 
“Yes,” “Yes.” Unbidden, a grin came to Sean's face.

“Why did you put Ethan in the Home?” Sean asked, shifting to catch 
more sun on his face.

“He wanted to go,” she said. “The doctor came by and told him 
about it and asked him if he wanted to go, and he said `Yes.' That was 
it!”

Sean snuck a look at Adele. She was wincing into the light, following 
it like a sunflower. “Adele,” he said.

“Yes, Sean?”

“Ethan was in maintenance mode. He was switched off. He said `Yes,' 
because his subroutines didn't want to be any trouble. You know that, 
right?”

“Oh, that foolishness again! Ethan's a \emph{good boy}, is all. He 
remembers my birthday and Mother's Day, every year.”

“Subroutines, Adele,” Sean said, straining to keep an inexplicable 
anger out of his voice.

“Humph! Subroutines!”

“Adele, he's a robot. He's a walking coma. He's been switched off for 
so long, all you're talking to is a goddamn \emph{chip}, he's not a 
goddamn \emph{person} anymore. None of them are. My goddamn 
\emph{Grampa}'s spent three-quarters of his goddamn life \emph{away}. 
He's either an angry old bastard, or he's a goddamn \emph{zombie}. You 
\emph{know that}, right?”

“Sean, you're very upset,” Adele said. “Why don't you have a nice 
lie-down, and we'll talk in the morning. I can't wait to meet your 
father!”

Sean stalked off to his room and tried to record some field notes while 
flipping around in the weird, poky corners of the motel's cable system, 
Japanese game-shows and Hindu religious epics. He smoked half a 
cigarette, drank half a beer, tried to masturbate, and finally, slept.

\tb

Adele rang his room-phone at eight. “Rise and shine, sunshine!” she 
said. “Your father will be at the airport in an hour!”

Sean dressed, but didn't bother shaving or brushing his teeth. He 
staggered out to his rental and gave Adele a sheepish grin. Acid 
churned in his gut.

Adele waited by the passenger door, in a pair of slacks and a light 
blouse. She had hung a pair of sunglasses around her neck on a gold 
chain, and carried an enormous sisal handbag. Staggering in the 
horrible daylight, Sean opened the passenger door for her, and offered 
his arm while she got in.

He put the car onto the Bee Line Expressway and pointed it at the 
airport.

“Oh, won't this be \emph{fun}?” Adele said, as he ground the crap 
from the corners of his eyes and steered with his knees. “I'm sure 
your father is \emph{charming}. Maybe the five of us can go to 
Universal for an afternoon.”

“I don't think we can take them off the ward,” Sean grunted, 
changing lanes for the airport exit.

“You're probably right,” Adele said. “I was just thinking that 
Universal might be enough to keep them both switched on.”

Sean shot her a look and nearly missed his exit.

Adele rattled a laugh at him. “Don't look so surprised. I know which 
end is up!”

Sean pursed his lips and navigated the ramp-maze that guarded the 
airport. He pulled up to the loading zone at Air Canada arrivals and 
switched off the engine. He looked past Adele at the tourists jockeying 
for cabs. “I'm sorry about yesterday. I guess I'm a little wound 
up.”

“Yesterday?” Adele said. “Oh! By the pool!” She put a frail 
hand on his forearm. “Sean, you don't get to my age by holding 
grudges. Ethan's father -- \emph{he} held grudges, and it killed him. 
Heart attack. He never forgave the doctors. I'm just happy to have a 
chauffeur.”

Sean swallowed hard. “I'm sure that somewhere, Ethan knows that 
you're visiting him, that you love him. He's in there.” He said it 
with all the sincerity he could muster.

“Maybe he is, maybe he isn't,” Adele said. “But it makes me feel 
better. He's what I've got left. If you'd like, I'll wait with the car 
so you can go in and look for your father.”

“No,” Sean said. “That's all right. Dad'll come out for a cab. 
He's not the sort to dawdle.”

“I like a decisive man. That's why I talked to you by the pool -- you 
just jumped in, because you wanted a swim.”

“Adele, that was \emph{stupid}. It was like swimming in a urine 
sample.”

“Same difference. I like a man who can make up his mind. That's what 
Ethan's father was like: decisive.”

“You'll like my Dad,” Sean said. He drummed his fingers on the 
wheel, then lowered and raised his window. He whistled tunelessly 
through his teeth. Adele gave him a considering stare and he stopped, 
and started in on powers of two in his head.

“There he is,” Sean said, 224 later.

Sean had barely been in Florida for three days, but it was long enough 
that his father seemed as pale as freezer-burned ice cream. Sean 
checked the traffic in his rear-view, then pulled across the waiting 
area to where his father stood, acing out an irate cabbie for the spot.

Sean's father glared at the car and started to walk behind it to the 
taxi. Sean leaned on the horn and his father stooped and stared. His 
expression was bland and grim and affectless.

Sean powered down Adele's window. “Dad!”

“Sean?” his father said.

Sean popped the locks. “Get in, Dad, I'll give you a ride.”

\tb

Adele turned around as Sean's father was buckling in. “I'm Adele. 
Sean and I were thinking of taking you to Universal. Would you like 
that?”

Sean's father stared right through her, at Sean. “It's an obvious 
question, I know, but what are you doing here?”

“It's my thesis,” Sean said, and floored the rental, headed for the 
Home.

“Whee!” Adele said.

“How's Grampa?” Sean's father asked.

“Oh, he's delightful,” Adele said. “We introduced him to my Ethan 
yesterday, and they're getting along famously. Sean, introduce me to 
your charming father, please.”

“Dad,” Sean said, through grit teeth, “This is Adele. Adele, my 
father, Mitch. We were thinking of getting day-passes for Grampa and 
Ethan and taking them to Universal. You ever been to Universal, Dad? I 
hear you come down here a lot.” His normally fragmented attention was 
as focused as a laser, boring into his father through the rear-view.

His father's stern face refused to expose any of his confusion. “I 
don't think I want to go to Universal,” he said.

“Oh, but it's \emph{wonderful},” Adele gushed. “You shouldn't 
knock it until you've tried it.”

“I don't think so,” Sean's father repeated. “What's your 
thesis?”

Sean plunged headlong into the breech. “It's called `The 
Tri-Generational Deficit: What's My Father's Excuse?'”

Sean's father nodded curtly. “And how's it going?”

“Well, you have to understand, I'm just warming up to the subject 
with Grampa. And then I'll have to do an interview series with you, of 
course.”

“Did I miss something? When did I become the principle ogre in your 
pantheon? Are you angry at me?”

Sean barked a laugh and turned onto the Home's exit-ramp. “I guess I 
am, Dad. Grampa had the operation -- it was \emph{easy} for him to 
switch off. You needed to make a special effort.” The words flew from 
his mouth like crows, and Sean clamped his jaw shut. He tensed for the 
inevitable scathe of verbiage. None came. He risked a glance in his 
rear-view.

His father was staring morosely out at the Home. Adele patted Sean's 
hand and gave him a sympathetic look. Sean parked the car.

\tb

“Hi, Pop,” Sean's father said, when they came to the table where 
Grampa sat. Ethan sat across from him.

Grampa glared at them. “This guy won't leave me alone. He's a fucking 
vegetable,” he said, gesturing at Ethan. Adele pursed her lips at 
him. He patted her arm absently. “It needed to be said.”

Sean's father reached around the table and gave Grampa a stiff hug. 
“Good to see you, Pop.”

“Yeah, likewise. Sit down, Mitch. Sit down, Sean. Sit down, Adele.” 
They sat. “Ask your questions, Sean,” he ordered.

Sean found himself tongue-tied. He heaved a deep breath and closed his 
eyes for a moment. He thought about why he was here: not the reason 
he'd given his thesis advisor, but the \emph{real} goddamn reason. He 
wanted to \emph{understand} -- his father, himself. He wanted to 
reverse-engineer his father's childhood. He looked at Ethan, slack as 
Grampa had been whenever they'd visited. An inkling glimmered. “Does 
Ethan scare you, Grampa?”

Adele \emph{tsk}ed and scowled.

“Do I scare you, Mitch?” Grampa said, to Sean's father.

“Yes,” Sean's father said.

“Yes,” Grampa said. “Next question.”

“Do you think that switching off is a sign of weakness?” Sean said, 
sneaking a glance at his father, seeing his grandfather's features 
echoed in his father's face.

“Yes,” his father said.

“Of course,” his grandfather said.

“Then why?” Sean said.

“You know why,” Ethan said, his eyes glittering.

They all swiveled to look at him. “Because the alternative is the 
purest shit,” Ethan said, standing up, starting to pace, almost 
shouting to make himself heard over the din of the ward. “Because if 
you have to ask, you'll never understand. Because dessert is better 
than dinner, because the cherry on top is the best part of the sundae. 
Because strength is over-rated.”

Grampa applauded briefly, sardonically. “Because holding your nose 
and taking your medicine is awful. Because boredom is a suppurating 
wound on the mind. Because self-discipline is over-rated. You getting 
all this, Sean?”

But Sean was watching his father, who was staring in fascinated horror 
at Grampa. Nauseous regret suffused Sean, as he saw his father's 
composure crumble. How many times had he tried to shatter that deadly 
cool? And here he'd done it. He'd really done it.

Still looking at his father, Sean said, “Do you ever wonder how it 
feels to rank below oblivion in someone's book?”

Grampa spread hands on the table. “I can't help it if you take it 
personally.”

Sean's father reeled back, and Sean swallowed a throb of anger. “Of 
\emph{course} not, Grampa. I understand. It's a reflex. The world's 
full of sops who'll take offence at any little thing” -- Lara 
shriveling under the heat of his tongue, and him still watching the TV 
over her shoulder -- “but it's a \emph{reflex}. It's not conscious. 
It's no one's \emph{fault}.”

“Don't humor me,” Grampa snapped. “I know what you all think of 
me. I can feel your goddamn blame. I can't \emph{do} anything about 
it.”

“You could apologize,” Ethan said. Adele took his hand and wiped at 
her tears with its back.

“Fuck off, zombie,” Grampa said, glaring at him.

Sean's father stood abruptly. “I'm glad to see you're in good health, 
Pop,” he said. “Sean, thanks for the ride. I guess I'll see you 
once you've finished your research.” His face was hard, composed. 
“Adele, nice to have met you.”

“Likewise,” Adele said.

“Bye, then,” Sean's father said, and walked with dignified calm to 
the elevator.

“Bye, Dad,” Sean called softly at his retreating back.

He turned back to Grampa, but Grampa's eyes were dull, and he was 
methodically twitching, top-to-bottom.

“Adele,” Sean said, taking her free hand.

“Yes?” she said.

“How would you and Ethan like to come to Universal with me for the 
afternoon?”

“I'd love to,” Ethan said. Sean looked at Ethan, and couldn't 
decide if he was switched off or not.

Whichever, Adele didn't seem to mind.

\section{Afterword}

I probably have some mild ADD, but I think I've turned it into a 
pro-survival adaptation for a life composed of lots of short, intense 
bursts of stimulus and work. But I'm also familiar with the powerful 
urge to switch off and make the boring stuff \emph{go away}.

My grandfather, Avram Doctorow, died of complications from senile 
dementia in a seniors' psych ward at a Jewish hospital in Toronto. It 
was a good place, but it wasn't a happy place, and his last few years 
were very hard. Most of the time, he just wasn't \emph{there}, but 
every now and again, he'd realize where he was, what had happened to 
him, and he'd cry uncontrollably. Those times are the most haunting 
things I've ever seen.

\end{document}
